---
title: "SESAM community composition"
output:
  pdf_document:
    keep_tex: yes
  html_notebook: default
  html_document:
    df_print: paged
---

# Game theoretically inspired derivation of community allocation $\alpha$

Deprecated: assumptions of ESS not fulfilled, because cost depends on alpha.
See sesam_LRP_opt for a derivation from finding optima at zero derivatives.

One of the core ideas of SESAM is the optimization of community composition, in
terms of enzyme allocation.
SESAM 3.3, assumed the proportion of enzyme allocation into enzyme $Z$ to be 
proportional
to the revenue, i.e. return per investment, of this given enzyme.
But, the arguments given, however, were not sufficiently rigorous (Wutzler X, appendix X).
Therefore, a more rigorous mathematical treatment of these assumptions 
is required.

The developments of XXWutzler22 with the new formulation of decomposition based 
on quasi-steady state of
enzymes and the new formulation of revenue with limitation-weighted cost, it
becomes possible to express the revenue as a function of the enzyme allocation.
Here, we make use of this possibility to study community optimization
from a mathematical game-theoretically inspired perspective.

We find that ...

## Methodology outline

With borrowing ideas from determining an evolutionary stable strategy (ESS)
in game theory, we look at a system with an incumbent microbial community playing
enzyme allocation strategy $\vec{\alpha}^*$, allocating share $\alpha^*_Z$ to 
enzyme $Z$. We seek $\vec{\alpha}^*$ to give optimal utility by asking if 
there is any beneficial deviant strategy $\vec{\alpha}$, that could yield a higher
utility.

The competing pure strategies are the different enzymes, 
and the utility is the efficiency of
generating supply flux for microbial growth per flux of enzyme allocation
$u_Z = \frac{{d_Z}_w(\alpha_Z)}{d_E \alpha_Z a_E B}$.
The supply ${d_Z}_w$ is a limitation-weighted mineralization flux or uptake flux of 
nutrients and carbon as detailed below. $a_E B$ is the total carbon flux allocated
to enzyme production proportional to microbial biomass, $B$, 
and $\alpha_Z$ is the share invested into production of
enzyme $Z$. The enzyme investment is a limitation-weighted flux of elemental 
investments into enzyme production.

In contrast to basic ESS theory, in our problem, the utility depends 
on the strategy mix $\alpha$ and it is not straightforward to setup a game situation
where individuals play against each other. This is because the utility depends on 
how large is the amount of
the enzyme population and the return is shared by all microbial biomass independent
of which enzymes they produced.

For example, consider the case of two depolymerizing enzymes, $L$, and $R$. 
The total return generated is 
$$
u(\vec{\alpha}) = d_E a_E B \left(\alpha_L \, u_L(\alpha_L) + \alpha_R \, u_R(\alpha_R)\right)
$$

It is shared by microbial populations both, the ones primarily producing 
labile litter degrading enzymes, and the ones primarily producing residue
degrading enyzmes.

The classic way of finding an ESS based on a mixed Nash equilibrium
is to find the partitioning so that the utility
for the other player is the same for all pure strategies participating in the mix. 
That is, if a pure strategy with higher utility was available to him, he would
not assign this mix, but assign higher share to the pure strategy with higher
utility.

This way does not work in the current case, because the utilities for the
pure strategies depend on the partitioning. Hence, it may happen that 
when all utilities as a function of their share are equal, 
that investing more into one of the pure
strategies may increase the utility more than decrease resulting from investing
less into another strategy.

The way we propose for the current case, is to seek an allocation where 
all the derivatives of total utility to the allocations are equal.

The derivate, i.e. change in generated supply, 
monotonically decreases with its share, because the generated supply saturates at
high enzyme levels. At those high enzyme levels, a increase in allocation only
increases the supply marginally. The maximum utility of an enzyme
is achieved at an arbitrarily small share ${du_Z}_{max} = u_Z(\alpha_Z \to 0)$.
If one of the derivatives is larger than the others, then increasing it at the
expense of decreasing the share to another enzyme will increase total utility.

The algorithm to determine the optimum strategy is as follows.

1) Order the enzymes according to their maximum derivative, ${du_Z}_{max}$, 
index them by $i$, set $i=1$ and start with a mix that includes only the most 
efficient enzyme $\{Z_1\}$.

2) Solve for the allocation
strategy $\alpha_i$ that makes derivative of the utility to allocation 
equal for each enzyme in the current mix, 
$$
\frac{d u(\vec{\alpha_i})}{d \alpha_Z} = du_i \, \forall Z \in \{Z_{1 \ldots i}\}
$$
and allocates nothing to enzymes that are not part of the current mix. 

3) If $du_i > {du_{Z_{i+1}}}_{max}$ stop and report the found optimum 
$\alpha^* = \alpha_i$. Otherwise increase $i$, i.e. include enzyme $Z_{i+1}$ in the
mix and go to step 2.
 
## Two depolymerizing enzymes
After the general way forward is outlined, the following text reports the
specifics of the utilities and the formulas for solving for the strategy
that equalizes the utilities, for the case of two enzymes 
depolymerizing labile OM, $L$, and residue OM, $R$, produced in shares
$\alpha_L = \alpha, \alpha_R = (1-\alpha)$ respectively.

The return is an limitation weighted average of the returns for all elements.

$$
\begin{aligned}
{d_L}_C &= k_L L \frac{\alpha a_E B}{k_{mn} + \alpha a_E B}\\
{d_L}_N &= {d_L}_C / {\beta_N}_L \\
{d_L}_P &= {d_L}_C / {\beta_P}_L \\
{d_L}_w &= w_C {d_L}_C + w_N {d_L}_N + w_P {d_L}_P \\
&= k_L L \frac{\alpha a_E B}{k_{mn} + \alpha a_E B} (w_C + w_N / {\beta_N}_L + w_P / {\beta_P}_L) \\
&= d_L \frac{\alpha a_E B}{k_{mn} + \alpha a_E B}
\end{aligned}
$$
with $d_L = k_L L (w_C + w_N / {\beta_N}_L + w_P / {\beta_P}_L)$.

similar
$$
{d_R}_w = d_R \frac{(1-\alpha) a_E B}{k_{mn} + (1-\alpha) a_E B}
$$
with $d_R = k_R R (w_C + w_N / {\beta_N}_R + w_P / {\beta_P}_R)$.


The carbon investment into enzyme $Z$ is $\operatorname{inv}_{Z,C} = \alpha_Z a B$. 
Similar to
weighing the elemental limitations for the return, SESAM weighs the
investments into enzymes, but uses stoichiometry of enzymes to convert
elemental costs to common carbon-based currency.

$$
\begin{aligned}
{\operatorname{inv}_Z}_w &= \alpha_Z a_E B (w_C + w_N / {\beta_N}_E + w_P / {\beta_P}_E)
\\
&= \alpha_Z a_E B \, d_E
\end{aligned}
$$
with $d_E = w_C + w_N / {\beta_N}_E + w_P / {\beta_P}_E$.

The utility from labile litter degrading enzyme with allocation $\alpha$, i.e. its revenue, is
$$
\begin{aligned}
u_L &= {d_L}_w / {\operatorname{inv}_Z}_w
\\
&=  \frac{d_L \alpha a_E B}{k_{mn} + \alpha a_E B} \frac{1}{d_E \alpha a_E B }
\\
&=  \frac{d_L}{d_E } \frac{1}{k_{mn} + \alpha a_E B} 
\end{aligned}
$$

And similar, the the revenue from residue litter with allocation $\alpha_R = 1 - \alpha$ is
$$
u_R = \frac{d_R}{d_E } \frac{1}{k_{mn} + \alpha_R a_E B} 
$$

For all strategies revenue $u_Z$ decreases with $\alpha_Z$.

Since the denominator is positive for $\alpha_Z \in [0,1]$, 
${u_Z}_{max} = u_Z(\alpha_Z \to 0) = u_Z(\alpha_Z = 0)$

The total return is the utility per enzyme flux multiplied by total
limitation-weighted enzyme flux $d_E a_E B$.
$$
\begin{aligned}
u(\vec{\alpha}) &= d_E a_E B\left(\alpha_L \, u_L(\alpha_L) + \alpha_R \, u_R(\alpha_R)\right)
\\
&= a_E B \left(\frac{\alpha_L d_L}{k_{mn} + \alpha_L a_E B} +  \frac{\alpha_R d_R}{k_{mn} + \alpha_R a_E B} \right)
\end{aligned}
$$

The $\alpha_Z$ occurs only in one term, hence, the derivative of the total
utility equals the derivative of revenue up to a factor that is the same
for all enzymes.
Hence, the constraint of equality of derivatives of total utiltiy to each enzyme 
allocation translates to equality of $\alpha_Z$ times the derivative of the
utility of each enzyme, i.e. its revenue.

We have the following derivatives of the total utility $u$.

$$
\begin{aligned}
\frac{d u}{d \alpha_Z} &= d_E a_E B \frac{d(\alpha_Z u_Z)}{d \alpha_Z}
\\
&= d_E a_E B \frac{d}{d \alpha_Z} \left( \frac{d_Z}{d_E } \frac{\alpha_Z}{k_{mn} + \alpha_Z a_E B} \right)
\\
&= a_E B d_Z \frac{(k_{mn} + \alpha_Z a_E B) - \alpha_Z a_E B}{(k_{mn} + \alpha_Z a_E B)^2} 
\\
&= a_E B k_{mn} \frac{d_Z }{(k_{mn} + \alpha_Z a_E B)^2} 
\\
\end{aligned}
$$

The three possible strategies yield potentially optimal results. Two resemble
the notion of pure Nash equilibrium and one resembles the notion of
a mixed Nash equilibrium.

- $u_L(\alpha_L = 1) > u_R(\alpha_R \to 0)$: only degrading labile litter, i.e. $\alpha = 1$.
- $u_R(\alpha_R = 1) > u_L(\alpha_L \to 0)$: only degrading residue litter, i.e. $\alpha = 0$.
- $u_L(\alpha) = u_R(1-\alpha)$: degrading both litter types.

For the third strategy we can solve for $\frac{d u_L}{d \alpha_L} = \frac{d u_R}{d \alpha_R}$.

$$
\begin{aligned}
\frac{d u}{d \alpha_L} &= \frac{d u}{d \alpha_R}
\\
\frac{d_L }{(k_{mn} + \alpha a_E B)^2}  &= \frac{d_R }{(k_{mn} + (1-\alpha) a_E B)^2}
\end{aligned}
$$
This provides a quadratic equation of $\alpha$, which one can solve.

The second root (TODO why second root) of the following equation 
provides the optimal $\alpha$.

$$
\begin{aligned}
\alpha_{1,2} &=
\frac{B a d_{L} + d_{L} k + d_{R} k \pm \sqrt{d_{L}} \sqrt{d_{R}} \left(B a + 2 k\right) }{B a \left(d_{L} - d_{R}\right)}
\\
\alpha_{1,2} &=
\frac{a_E B d_{L} + k_{mn}(d_{L} + d_{R}) \pm \sqrt{d_{L} d_{R}} \left(a_E B + 2 k_{mn}\right) }{a_E B \left(d_{L} - d_{R}\right)}
\end{aligned}
$$

### Visualize two enzymes 

Lets plot some scenarios of dL = 1 (black) and dL = 0.5 (blue, maroon)
versus dR increasing from 0 to 1.
In the maroon case, the biomass is increased.


```{r echo=FALSE}
calc_alpha <- function(dL, dR, p){
  dLmdR = dL - dR
  #ifelse(cbind(dLmdR,dLmdR) == 0, c(0.5,0.5), {
  ifelse(dLmdR == 0, 0.5, {
    aeB = p$a_E*p$B
    A = aeB * dL + p$k_mn*(dL+dR)
    D = (aeB + 2*p$k_mn)*sqrt(dL*dR)
    d = aeB*dLmdR
    # q = p$k_mn^2 + aeB*(2+aeB)*dL/dLmdR
    # D = sqrt(A^2-q)
    # alpha0 <- ifelse(dL < dR, (A + D)/aeB, (A - D)/aeB)    
    # alpha <- pmin(1,pmax(0,alpha0))
    #alpha <- cbind((A+D)/d, (A-D)/d)
    alpha0 <- (A-D)/d  # only the second root is reasonable
    alpha <- pmin(1,pmax(0,alpha0))
  })
}
p = within(list(
  a_E = 0.1, 
  B=1, 
  e_P = 0,
  tau = 365/30,
  synB = 15
), k_mn <- a_E*B/2)
dR <- seq(0,1,length.out=5); dL <- 0.5
#calc_alpha(dL, seq(0,1,length.out=5), p)
nstep = 200
dRs = seq(0,1, length.out=nstep)
plot(dRs, calc_alpha(1,dRs,p), type="l", ylim=c(0,1), ylab=expression(alpha), xlab=expression(d[R]))
#plot(dRs, calc_alpha(0.5,dRs,p), type="l", ylim=c(0,1), ylab=expression(alpha), xlab=expression(d[R]))
abline(h=0.5, col="lightgray", lty="dotted")
lines(dRs, calc_alpha(0.5,dRs,p), col="blue")
pB <- within(p, B <- 0.5)
lines(dRs, calc_alpha(0.5,dRs,pB), col="maroon")
```

The shape is only marginally modified by a halved biomass (maroon).
The largest different is that the range is increased where allocation to
a single enzyme is best.

```{r eval=FALSE, include=FALSE}
# check symmetry for dR=0.5 changing dL
plot(dRs, calc_alpha(dRs,0.5,p), type="l", ylim=c(0,1), ylab=expression(alpha), xlab=expression(d[L])); abline(h=0.5, col="lightgray", lty="dotted")
```

The following plot for example shown by maroon line above demonstrates 
that the derivatives of the utilities are equal in maroon example above.

```{r echo=FALSE}
calc_du_dalpha <- function(alpha, dS, p, e_P =  p$e_P){
  aeB = p$a_E * p$B
  aeB * p$k_mn * dS / (e_P + p$k_mn + alpha*aeB)^2
}
dL <- 0.5
dRs = seq(0,1, length.out=40)
dE <- 0.2
alphas = calc_alpha(dL, dRs, pB)
duL = calc_du_dalpha(alphas, dL, pB)
duR = calc_du_dalpha(1-alphas, dRs, pB)
plot(alphas, duL, ylab="duL (black) duR (blue)")
lines(alphas, duR, col="blue", lty="dotted")
```

### Checking optimality
Is the utility of a community playing a different strategy $\alpha$
inferior to a a community playing $\alpha^*$? 
We numerically compare the total utility 
$u(\vec{\alpha}) = \alpha_L u_L(\alpha) + \alpha_R u_R(\alpha_R))$ between
$\vec{\alpha^*}$ and a slightly modified $\vec{\alpha}$.

First, decrease alpha slightly, $\alpha = 0.99 \alpha^*$.

```{r include=FALSE}
u_decomp <- function(alpha, dLorR, dE, p){
  dLorR/dE/(p$k_mn + alpha * p$a_E * p$B)
}
u_biomin <- function(alpha, dP, dE, p){
  dP*p$k_mn/dE/((p$k_mn + p$e_P)^2 + alpha * p$a_E * p$B * ((p$k_mn + p$e_P)))
}
u_2depoly <- function(alpha, dL, dR, dE, p){
  (alpha * u_decomp(alpha, dL, dE, p) + (1-alpha)* u_decomp((1-alpha), dR, dE, p))*p$a_E*p$B*dE
}
du_dalphaS <- function(alpha, dS, p, e_P =  p$e_P){
  aeB = p$a_E * p$B
  aeB * p$k_mn * dS / (e_P + p$k_mn + alpha*aeB)^2
}
du_dalpha <- function(alpha3, dL, dR, dP, p){
  if (is.matrix(alpha3)) {
    cbind(
      du_dalphaS(alpha3[,"L", drop=FALSE], dL, p, e_P = 0),
      du_dalphaS(alpha3[,"R", drop=FALSE], dR, p, e_P = 0),
      du_dalphaS(alpha3[,"P", drop=FALSE], dP, p)
    )#,c("L","R","P"))
  } else {
    c(
      du_dalphaS(alpha3["L"], dL, p, e_P = 0),
      du_dalphaS(alpha3["R"], dR, p, e_P = 0),
      du_dalphaS(alpha3["P"], dP, p)
    )
  }
}
```

```{r eval=FALSE, include=FALSE}
.tmp.f <- function(){
  alpha3 <- c(L=0.6,R=0.3,P=0.1)  
  du_dalpha(alpha3, dL, dR, dP, p)
  du_dalpha(t(alpha3), dL, dR, dP, p)
  du_dalpha(rbind(alpha3, alpha3), dL, dR, dP, p)
}
```

```{r echo=FALSE}
is_ess <- alphas < 1 & alphas > 0
aps = 0.99 * alphas
u_ap = u_2depoly(aps, dL, dRs, dE, pB)
u_a =  u_2depoly(alphas, dL, dRs, dE, pB)
#rbind(dRs, u_ap, u_a, u_ap - u_a)[,is_ess]
#plot(dRs[is_ess], u_a[is_ess], type="l")
#lines(dRs[is_ess], u_ap[is_ess], col="blue")
plot(dRs[is_ess], (u_ap - u_a)[is_ess], ylab=expression(u[alpha]-u[alpha^'*']), 
     xlab=expression(d[R]), ylim=c(min((u_ap - u_a)[is_ess]),0))
abline(h=0)
#max((u_ap - u_a)[is_ess], na.rm = TRUE)
```
```{r eval=FALSE, include=FALSE}
# check specific case
dL <- 0.5; dR <- 0.4
alpha <- calc_alpha(dL, dR, pB); alpha
u_2depoly(alpha, dL, dR, dE, pB)
u = c(u_2depoly(alpha-0.01, dL, dR, dE, pB),u_2depoly(alpha, dL, dR, dE, pB))
u[1] - u[2]
```

The utility $u_\alpha$ is always smaller than $u_{\alpha^*}$, 
although the difference becomes small for $d_R \to d_L$.

Now for slightly larger $\alpha = max(1, 1.01 \alpha^*)$

```{r echo=FALSE}
is_ess <- alphas < 1 & alphas > 0
aps = pmax(1,1.01 * alphas)
u_ap = u_2depoly(aps, dL, dRs, dE, pB)
u_a =  u_2depoly(alphas, dL, dRs, dE, pB)
#rbind(dRs, u_ap, u_a, u_ap - u_a)[,is_ess]
#plot(dRs[is_ess], u_a[is_ess], type="l")
#lines(dRs[is_ess], u_ap[is_ess], col="blue")
plot(dRs[is_ess], (u_ap - u_a)[is_ess], ylab=expression(u[alpha]-u[alpha^'*']), 
     xlab=expression(d[R]), ylim=c(min((u_ap - u_a)[is_ess]),0))
abline(h=0)
max((u_ap - u_a)[is_ess], na.rm = TRUE)
```
Again, the utility of the modified strategy is always smaller.
There is no profitable deviation from strategy $\alpha^*$.

## Three enzymes

Consider the case where in addition to the two depolymerizing enzymes, there
is an additional biomineralizing enzyme. 
This biomineralizing enzyme is also produced by plants with rate $e_P$. 
Hence, we need to
calculate the effect of microbe-produced enzyme in addition to plant.

With $d_P = w_P (k_L L / {\beta_P}_L + k_R R / {\beta_P}_R)$
assuming that P is produced from SOM at the same potential rate as decomposition
and assuming that biomineralization enzymes attack both, labile and residue OM
we calculate the return ${d_P}_m$.

$$
\begin{aligned}
{d_P}_m &= d_P \frac{ e_P + \alpha_P a_E B}{k_{mn} + e_P + \alpha_P a_E B} - d_P \frac{ e_P}{k_{mn} + e_P}
\\
 &= d_P \frac{ (e_P + \alpha_P a_E B)(k_{mn} + e_P) - e_P(k_{mn} + e_P + \alpha_P a_E B)}{(k_{mn} + e_P + \alpha_P a_E B)(k_{mn} + e_P)} 
\\
 &= d_P \frac{ e_P k_{mn} + \alpha_P a_E B k_{mn} + e_P^2 + \alpha_P a_E B e_P - (e_P k_{mn} + e_P^2  + \alpha_P a_E B e_P)}
{(k_{mn} + e_P)^2 + \alpha_P a_E B (k_{mn} + e_P)} 
\\
 &= d_P \frac{ \alpha_P a_E B k_{mn}}
{(k_{mn} + e_P)^2 + \alpha_P a_E B (k_{mn} + e_P)} 
\\
 &= d_P \frac{k_{mn}}{e_P + k_{mn}} \frac{ \alpha_P a_E B }
{(e_P + k_{mn}) + \alpha_P a_E B } 
\end{aligned}
$$

The return has the form of a Michaelis-Menten kinetics with decreased
affinity, i.e. increased half-saturation, and the maximum rate multiplied
by another Michaelis-Menten term of $k_{mn}$ with half-saturation $e_P$.

The revenue from biomineralizing enzyme with allocation $\alpha_P$ is
$$
\begin{aligned}
u_P &= {d_P}_m / {\operatorname{inv}_P}_w
\\
&=  \frac{ d_P\alpha_P a_E B k_{mn}}
{(k_{mn} + e_P)^2 + \alpha_P a_E B (k_{mn} + e_P)} 
\frac{1}{d_E \alpha_P a_E B }
\\
&=  \frac{d_P}{d_E } \frac{k_{mn}}{(k_{mn} + e_P)^2 + \alpha_P a_E B (k_{mn} + e_P)}
\\
&=  \frac{d_P}{d_E } \frac{k_{mn}}{e_P + k_{mn}} \frac{ 1 }
{(e_P + k_{mn}) + \alpha_P a_E B }
\end{aligned}
$$

Hence, the derivative $\frac{u}{d \alpha_P}$

$$
\begin{aligned}
\frac{du}{d \alpha_P} &= d_E a_E B \frac{d (\alpha_P u_P)}{d \alpha_P} 
\\
&= a_E B d_P \frac{k_{mn}}{e_P + k_{mn}}  \frac{d }{d \alpha_P}\left(  \frac{ \alpha_P }
{(e_P + k_{mn}) + \alpha_P a_E B } \right)
\\
&= a_E B d_P \frac{k_{mn}}{e_P + k_{mn}}   
\frac{ (e_P + k_{mn}) + \alpha_P a_E B - \alpha_P (a_E B) }
{((e_P + k_{mn}) + \alpha_P a_E B)^2 }
\\
&=  a_E B k_{mn}
\frac{ d_P  }
{((e_P + k_{mn}) + \alpha_P a_E B)^2 }
\end{aligned}
$$
$$
\begin{aligned}
\end{aligned}
$$



For the mixed strategy optimum involving three enzymes, we have

$$
\begin{aligned}
\frac{d(\alpha_L u_L)}{d \alpha_L} 
&= \frac{d(\alpha_R u_R)}{d \alpha_R} 
&= \frac{d(\alpha_P u_P)}{d \alpha_P} 
\\
 \frac{d_L}{(k_{mn} + \alpha a_E B)^2} 
&=  \frac{d_R}{(k_{mn} + (1-\alpha-\alpha_P) a_E B)^2} 
&=  \frac{ d_P }
{((e_P + k_{mn}) + \alpha_P a_E B)^2 }
\end{aligned}
$$

We first compute $\alpha$ given $\alpha_P$ using the first equality.
(see sesam_LRP_deriv_sympy.py)

$$
\begin{aligned}
{\alpha_L}_{12} &= \frac{- a_E B \alpha_{P} d_{L} + a_E B d_{L} \pm \sqrt{d_{L}} \sqrt{d_{R}} \left(a_E B \alpha_{P} - a_E B - 2 k_{mn}\right) + d_{L} k_{mn} + d_{R} k_{mn}}{a_E B \left(d_{L} - d_{R}\right)}
\\
&= \frac{a_E B( 1-  \alpha_{P}) d_{L}  + k_{mn}(d_{L} + d_{R}) \pm \sqrt{d_{L}d_{R}} \left(a_E B (1-\alpha_{P}) + 2 k_{mn}\right) }
{a_E B \left(d_{L} - d_{R}\right)}
\end{aligned}
$$

Next we insert the second root of $\alpha_L(\alpha_P)$ in equating
the first and third utility to solve for $\alpha_P$.
TODO - why second root?


$$
\begin{aligned}
{\alpha_P}_{12} &= (A \pm D)/B
\\
A &= 2 B a d_{L}^{\frac{3}{2}} d_{P} \sqrt{d_{R}} - B a d_{L}^{2} d_{P} - B a d_{L} d_{P} d_{R} + 4 d_{L}^{\frac{3}{2}} d_{P} \sqrt{d_{R}} k - d_{L}^{3} e_{P} - d_{L}^{3} k \\
&- 2 d_{L}^{2} d_{P} k + 2 d_{L}^{2} d_{R} e_{P} + 2 d_{L}^{2} d_{R} k - 2 d_{L} d_{P} d_{R} k - d_{L} d_{R}^{2} e_{P} - d_{L} d_{R}^{2} k 
\\
D &= 
 \sqrt{d_{P}} \left(B a + e_{P} + 3 k\right) \sqrt{- 2 d_{L}^{\frac{9}{2}} \sqrt{d_{R}} + 4 d_{L}^{\frac{7}{2}} d_{R}^{\frac{3}{2}} - 2 d_{L}^{\frac{5}{2}} d_{R}^{\frac{5}{2}} + d_{L}^{5} - d_{L}^{4} d_{R} - d_{L}^{3} d_{R}^{2} + d_{L}^{2} d_{R}^{3}}
\\
B &= B a \left(2 d_{L}^{\frac{3}{2}} d_{P} \sqrt{d_{R}} + d_{L}^{3} - d_{L}^{2} d_{P} - 2 d_{L}^{2} d_{R} - d_{L} d_{P} d_{R} + d_{L} d_{R}^{2}\right)
\end{aligned}
$$

```{r include=FALSE}
calc_alpha3_optL <- function(alphaP, dL, dR, dP, p){
  dLmdR = dL - dR
  #ifelse(cbind(dLmdR,dLmdR) == 0, c(0.5,0.5), {
  ifelse(dLmdR == 0, 0.5, {
    aeB = p$a_E*p$B
    A = aeB*(1-alphaP) * dL + p$k_mn*(dL+dR)
    D = (aeB*(1-alphaP) + 2*p$k_mn)*sqrt(dL*dR)
    d = aeB*dLmdR
    # q = p$k_mn^2 + aeB*(2+aeB)*dL/dLmdR
    # D = sqrt(A^2-q)
    # alpha0 <- ifelse(dL < dR, (A + D)/aeB, (A - D)/aeB)    
    # alpha <- pmin(1,pmax(0,alpha0))
    #alpha <- cbind((A+D)/d, (A-D)/d)
    alpha0 <- (A-D)/d  # only the second root is reasonable
    alpha <- pmin(1,pmax(0,alpha0))
  })
}
calc_alpha3_optLRP <- function(dL,dR, dP, p){
  aeB = p$a_E*p$B
  A = 2*aeB*dL^(3/2)*dP*sqrt(dR) - aeB*dL^2*dP - aeB*dL*dP*dR + 4*dL^(3/2)*dP*sqrt(dR)*p$k_mn - dL^3*p$e_P - dL^3*p$k_mn - 
      2*dL^2*dP*p$k_mn + 2*dL^2*dR*p$e_P + 2*dL^2*dR*p$k_mn - 2*dL*dP*dR*p$k_mn - dL*dR^2*p$e_P - dL*dR^2*p$k_mn 
  D = sqrt(dP)*(aeB + p$e_P + 3*p$k_mn)*sqrt(-2*dL^(9/2)*sqrt(dR) + 4*dL^(7/2)*dR^(3/2) - 2*dL^(5/2)*dR^(5/2) + dL^5 - dL^4*dR - dL^3*dR^2 + dL^2*dR^3)
  B = aeB*(2*dL^(3/2)*dP*sqrt(dR) + dL^3 - dL^2*dP - 2*dL^2*dR - dL*dP*dR + dL*dR^2)
  alphaP <- (A+D)/B
  alphaL <- calc_alpha3_optL(alphaP, dL,dR,dP,p)
  alphaR <- 1-alphaL-alphaP
  alpha=cbind(L=alphaL,R=alphaR,P=alphaP)
  alpha
}
.tmp.f <- function(){
  dL = 0.7; dR = 0.5; dP = 0.3
  p = within(list(
    a_E = 0.1, 
    B=1, 
    e_P = 0,
    tau = 365/30,
    synB = 15
  ), k_mn <- a_E*B/2)
  calc_alpha3_optLRP(dL,dR,dP,p)
  calc_alpha3_optLRP(dL,dR,c(0.3,0.5),p)
  #sum(alpha3)
}
```


### Depolymerizing and biomineralizing enzyme

For the mixed strategy optimum involving 
a depolymerizing and a biomineralizing enzyme we have:

$$
\begin{aligned}
\begin{aligned}
\frac{d(\alpha_L u_L)}{d \alpha_L} 
&= \frac{d(\alpha_P u_P)}{d \alpha_P} 
\\
 \frac{d_L}{(k_{mn} + \alpha_L a_E B)^2} 
&=  \frac{ d_P }
{((e_P + k_{mn}) + (1-\alpha_L) a_E B)^2 }
\end{aligned}\end{aligned}
$$

$$
\begin{aligned}
{\alpha_L}_{12} = 
\frac{(a_E B  + e_{P} + k_{mn}) d_{L} + k_{mn} d_{P}  \pm \sqrt{d_{L}d_{P}} \left(a_E B + e_{P} + 2 k_{mn}\right)}
{a_E B \left(d_{L} - d_{P}\right)}
\end{aligned}
$$
For enzymes R and P just replace $d_L$ by $d_R$.

```{r include=FALSE}
calc_alphaS_optSP <- function(dS, dP, p){
  dLmdR = dS - dP
  #ifelse(cbind(dLmdR,dLmdR) == 0, c(0.5,0.5), {
  alphaL <- ifelse(dLmdR == 0, 0.5, {
    aeB = p$a_E*p$B
    A = (aeB+p$e_P+p$k_mn)*dL + p$k_mn*dP
    D = sqrt(dL*dP)*(aeB + p$e_P + 2*p$k_mn)
    d = aeB*dLmdR
    # q = p$k_mn^2 + aeB*(2+aeB)*dL/dLmdR
    # D = sqrt(A^2-q)
    # alpha0 <- ifelse(dL < dR, (A + D)/aeB, (A - D)/aeB)    
    # alpha <- pmin(1,pmax(0,alpha0))
    #alpha <- cbind((A+D)/d, (A-D)/d)
    alpha0 <- (A-D)/d  # only the second root is reasonable
    alphaL <- pmin(1,pmax(0,alpha0))
  })
  alphaL
}
.tmp.f <- function(){
  dL = 0.7; dR = 0.1; dP = 0.3
  p = within(list(
    a_E = 0.1, 
    B=1, 
    e_P = 0,
    tau = 365/30,
    synB = 15
  ), k_mn <- a_E*B/2)
  calc_alphaS_optSP(dL,seq(0,1,length.out=5),p)
  #sum(alpha3)
}
```

### Visualize three enzymes 

Lets plot some scenarios of dL = 1 (black) and dL = 0.5 (blue, maroon)
versus dR of 0.7 and d_P increasing from 0 to 1 without plant enzyme
production, e_P =0.

In the maroon case eP = aE*B/2.

The utility of a depolymerizing enzmye decreases with $\alpha$ and stronger
with larger biomass (blue line).


```{r echo=FALSE}
# utility of depolymerizing enzyme
p = within(list(
  a_E = 0.1, 
  B=1, 
  e_P = 0
), k_mn <- a_E*B/2)
alpha = seq(0,1,length.out=31)
dL = dLorR= 1
dR = 0.7
dP = seq(0,1,length.out=30)
dE = 0.2
plot(u_decomp(alpha, dL, dE,  within(p, B <-2)) ~ alpha, type = "l", col="blue",
     ylab=expression(u[L]), xlab=expression(alpha))
lines(u_decomp(alpha, dL, dE, p) ~ alpha)
```

The utility is smaller, with plant mineralizing enzymes, but does not
decrease as strongly.

```{r echo=FALSE}
# utility of biomineralizing enzyme
u_e0 <- u_biomin(alpha, dL, dE,  p)
u_e1 <- u_biomin(alpha, dL, dE, within(p, e_P <- a_E*B/5))
plot(u_e0 ~ alpha, type = "l", ylim=range(c(u_e0,u_e1)),
     ylab=expression(u[L]), xlab=expression(alpha))
lines(u_e1 ~ alpha,col="blue")
```

```{r echo=FALSE}
calc_alpha3_opt = function (dL,dR, dP, p, dE=0.2){
  dumax = sort(du_dalpha(c(L=0,R=0,P=0), dL, dR, dP, p), decreasing = TRUE)
  if(names(dumax[1]) == "L") {
    du1 = du_dalphaS(1, dL, p, e_P=0)
    if (du1 > dumax[2]) return(c(L=1,R=0,P=0))
    if (names(dumax[2]) == "R"){
      alphaL = calc_alpha(dL, dR, p)
      du2 = du_dalphaS(alphaL, dL, p, e_P = 0)
      if (du2 > dumax[3]) return(c(L=alphaL,R=(1-alphaL),P=0))
    } else { # umax2 must be P
      alphaL = calc_alphaS_optSP(dL,dP,p)
      du2 = du_dalphaS(alphaL, dL, p, e_P = 0)
      if (du2 > dumax[3]) return(c(L=alphaL,R=0,P=(1-alphaL)))
    }
    alpha3 = calc_alpha3_optLRP(dL,dR,dP,p)
    return(alpha3)
  }
  if(names(dumax[1]) == "R") {
    du1 = du_dalphaS(1, dR, p, e_P = 0)
    if (du1 > dumax[2]) return(c(L=0,R=1,P=0))
    if (names(dumax[2]) == "L"){
      alphaL = calc_alpha(dL, dR, p)
      du2 = du_dalphaS(alphaL, dL, p, e_P = 0)
      if (du2 > dumax[3]) return(c(L=alphaL,R=(1-alphaL),P=0))
    } else { # umax2 must be P
      alphaR = calc_alphaS_optSP(dR,dP,p)
      du2 = du_dalphaS(alphaR, dR, p, e_P = 0)
      if (du2 > dumax[3]) return(c(L=0,R=alphaR,P=(1-alphaR)))
    }
    alpha3 = calc_alpha3_LRP(dL,dR,dP,p)
    return(alpha3)
  }
  if(names(dumax[1]) == "P") {
    du1 = du_dalphaS(1, dP, p)
    if (du1 > dumax[2]) return(c(L=0,R=0,P=1))
    if (names(dumax[2]) == "L"){
      alphaL = calc_alphaS_optSP(dL,dP,p)
      du2 = du_dalphaS(alphaL, dL, p, e_P = 0)
      if (du2 > dumax[3]) return(c(L=alphaL,R=0,P=(1-alphaL)))
    } else if (names(dumax[2]) == "R") {
      alphaR = calc_alphaS_optSP(dR,dP,p)
      du2 = du_dalphaS(alphaR, dR, p, e_P = 0)
      if (du2 > dumax[3]) return(c(L=0,R=alphaR,P=(1-alphaR)))
    }
    alpha3 = calc_alpha3_optLRP(dL,dR,dP,p)
    return(alpha3)
  }
  stop("unhandled umax: ", dumax)
} 
.tmp.f <- function(){
  dL = 0.7; dR = 0.5; dP = 0.6
  dL = 0.7; dR = 0.5; dP = 0.3
  alpha3 = calc_alpha3_opt(dL,dR,dP,p)
  alpha3
  # all derivatives equal?
  # c(u_decomp(alpha3[,"L"], dL, dE, p), u_decomp(alpha3[,"R"], dR, dE, p), u_biomin(alpha3[,"P"], dP, dE, p))
  du_dalpha(alpha3, dL, dR, dP, p)
}
```

```{r include=FALSE}
compute_utilities <- function(alpha3,dL,dR,dP,p,dE=0.2){
  c(
      u_decomp(alpha3["L"], dL, dE, p), 
      u_decomp(alpha3["R"], dR, dE, p), 
      u_biomin(alpha3["P"], dP, dE, p)
    )
}
compute_total_utilities <- function(alpha3,dL,dR,dP,p,dE=0.2){
  u = compute_utilities(alpha3,dL,dR,dP,p,dE)
  sum(alpha3 * u)
}
calc_alpha3_relative = function(dL,dR,dP,p,dE=0.2) {
  u0 = u1 = sapply(list(L=dL, R=dR, P=dP), function(x) x/(dE*p$k_mn))
  # fixpoint iteration
  for (i in 1:5) {
    alpha3 = u1/sum(u1)
    u1 = compute_utilities(alpha3,dL,dR,dP,p,dE)
    #print(c(u1, sum(alpha3*u1)))
  }
  alpha3
}
.tmp.f <- function(){
  calc_alpha3_relative(dL,dR,dP,p)  
}
```

```{r echo=FALSE}
nstep = 200
dL = 0.7
dR = 0.5
dPs = seq(0,1, length.out=nstep)
alpha3 = sapply(dPs, function(dP){calc_alpha3_opt(dL,dR,dP,p)} )
alpha3r = sapply(dPs, function(dP){calc_alpha3_relative(dL,dR,dP,p)} )

plot(dPs, alpha3["P",], type="l", ylim=c(-0.1,1.1), ylab=expression(alpha), xlab=expression(d[P]))
#plot(dPs, alpha3["R",], type="l",  ylab=expression(alpha), xlab=expression(d[P]))
lines(dPs, alpha3["L",], type="l", lty="dashed")
lines(dPs, alpha3["R",], type="l", lty="dotted")
#
lines(dPs, alpha3r["P",], type="l", col="maroon")
lines(dPs, alpha3r["L",], type="l", lty="dashed", col="maroon")
lines(dPs, alpha3r["R",], type="l", lty="dotted", col="maroon")
```

The relative-revenue heuristics overestimtes the allocation to the enzymes
with lowest returns and underestimates allocation to the enzymes
with highest returns.
This is most significant in the case where not allocating to the
enzyme at all is optimal.
Otherwise, it approximates the optimal solution quite well.

```{r echo=FALSE}
usum = sapply(seq_along(dPs), function(i){compute_total_utilities(alpha3[,i],dL,dR,dPs[i],p,dE)})
usum_r = sapply(seq_along(dPs), function(i){compute_total_utilities(alpha3r[,i],dL,dR,dPs[i],p,dE)})
plot(dPs, usum, type="l", ylim=range(c(usum,usum_r)),ylab="total utility", xlab=expression(d[P]))
#plot(dPs, alpha3["R",], type="l",  ylab=expression(alpha), xlab=expression(d[P]))
lines(dPs, usum_r, type="l", col="maroon")
```

The total utility approximated by the relative-revenue heuristics
is quite close to the optimal one, aside from the no-allocation case.

```{r eval=FALSE, include=FALSE}
# inspect specific ase
dP = 0.3
alpha3s = calc_alpha3_opt(dL,dR,dP,p)
alpha3sr = calc_alpha3_relative(dL,dR,dP,p)
cbind(t(alpha3s), alpha3sr)
c(compute_total_utilities(alpha3s[1,],dL,dR,dP,p,dE), compute_total_utilities(alpha3sr,dL,dR,dP,p,dE))
```


