---
title: "SESAM community composition"
output:
  pdf_document: 
    keep_tex: yes
  html_notebook: default
---

# Derivate-based community allocation change 

## Methodology outline

SESAM assumes that microbial community develops in a way to optimize growth
of the entire community. 
We approximate growth by the total limitation-weighted return, i.e. decomposition
and depolymerization flux, for given enzyme allocation.
The revenue of allocation to enzyme $Z$ is
$\operatorname{rev}_Z = \frac{{d_Z}_w(\alpha_Z)}{\alpha_Z \omega_{Enz} a_E B}$.
The return ${d_Z}_w$ is a limitation-weighted mineralization flux 
or uptake flux of nutrients and carbon as detailed below. 
The investment is the share, $\alpha_Z$, invested into production of
enzyme $Z$ multiplied by total limitation-weighted flux, $\omega_{Enz} a_E B$, allocated
to enzyme production.

Although its possible to derive explicit formula for the allocation that 
optimizes total return for simple cases, 
the formulas quickly grow and involve higher-order 
polynomials of $\alpha$ with several solutions outside the reasonable bound 
$\alpha_Z \in [0,1]$.

Here we follow an alternative local approach 
were we assume the rate change of $\alpha_Z$ over time 
to be proportional to the deviation of 
the derivative of change of total return to $\alpha_Z$ to the average across
the derivatives for different enzymes.
The higher the increase in total return for shifting allocation towards a specific
enzyme, the faster the community changes in this direction.

The total return has been derived in Appendix x to be a weighted sum
of enzyme revenues, and derivatives of 
$\frac{d (\alpha_Z \operatorname{rev}_Z)}{d\alpha_Z}$ 
have been derived for depolymerizing and biomineralizing enzymes.

$$
\begin{aligned}
u_T &=  \omega_{Enz} a_E B \sum_Z \alpha_Z \operatorname{rev}_Z(\alpha_Z)
\\
\frac{d u_T}{d \alpha_Z} &= \omega_{Enz} a_E B \sum_Z \frac{d (\alpha_Z \operatorname{rev}_Z)}{d\alpha_Z}
\end{aligned}
$$

Because the different $\alpha_Z$ occur in different terms, it is sufficient
to compute the derivative for the return of each enzyme separately.

We assume that the larger the change in return with increasing allocation, i.e. 
the derivative to allocation coefficient $\alpha_Z$,
the larger is the change in allocation.
In addition to the assumption of proportionaly to the derivative, we assume that
the community changes
at a rate of the same magnitude as synthesis and turnover of microbial biomass.

$$
\begin{aligned}
\frac{d \alpha_Z}{dt} &\propto \frac{d u}{d \alpha_Z} - m_{du}
\\
&= \left( \frac{|\operatorname{syn}_B|}{B} + \tau \right) 
\frac{\frac{d u}{d \alpha_Z} - m_{du}}{m_{du}}
\\
m_{du} &= \operatorname{mean}_\zeta\left(  \frac{d u}{d \alpha_\zeta} \right)
\end{aligned}
$$
where $m_{du}$ is the average across derivatives of return across enzymes 
that are allocated to.
If all changes are the same, i.e. equal to the mean, the allocation 
is optimal, since it does not increase in any direction.

We want the change to be proportional to the change in return compared 
to the average return.
Subtracting this mean ensures that the sum of all the changes in $\alpha$ 
sums to zero so that the sum across $\alpha$ is preserved.
The proportionality factor normalizes the change in return and multiplies
this relative change by the rate of microbial turnover, composed of biomass
synthesis and biomass turnover.


### Exclude enzymes whose negative relative change is larger than its share

Community may not allocate to all enzymes. Hence, $m_{du}$ averages only across 
a subset of enzymes.
An enzyme is not allocated to if its normalized change towards zero is larger than 
than its current share, i.e. is more negative than $-\alpha_Z$.

$$
\begin{aligned}
Z_0 &= \left\{ Z | \frac{\frac{d u}{d \alpha_Z} - m_{due}}{m_{due}} < -\alpha_Z  \right\}
\\
\frac{d \alpha_Z}{dt} &= \left( \frac{|\operatorname{syn}_B|}{B} + \tau \right)
\begin{cases}
-\alpha_Z &\text{ for } Z \in Z_0
\\
 \frac{\frac{d u}{d \alpha_Z} - m_{due}}{m_{due}} &\text{ otherwise}
\end{cases}
\\
&= \left( \frac{|\operatorname{syn}_B|}{B} + \tau \right) 
\max \left( \frac{\frac{d u}{d \alpha_Z} - m_{due}}{m_{due}}, -\alpha_Z \right)
\\
m_{due} &= \frac{\sum_{\zeta \notin Z_0} \frac{d u}{d \alpha_Z} }
{ |\{Z\} \setminus Z_0| + \sum_{\zeta \in Z_0} \alpha_Z}
\end{aligned}
$$
Where $|\{Z\} \setminus Z_0|$ denotes the number of enzymes allocated to, 
i.e. the cardinality of the set of all enzymes without those in $Z_0$
The relative change of those excluded enzymes is set to $-\alpha_Z$, resulting
in negative changes going to zero as $\alpha_Z$ approaches zero.
Hence, the relative change is lower-bounded by $-\alpha_Z$.

$m_{due}$ has to be adjusted, so that $\sum \frac{d \alpha_Z}{dt} = 0$ holds.
$$
\begin{aligned}
\sum_{\zeta \notin Z_0} \frac{\frac{d u}{d \alpha_Z} - m_{due}}{m_{due}} + \sum_{\zeta \in Z_0} -\alpha_Z = 0
\end{aligned}
$$

This definition is recursive, because $m_{due}$ is computed 
across a set that is defined using $m_{due}$
In order to determine $Z_0$ 
one can start with the empty set and add all enzymes that fulfill the condition. 
If enzymes were added then the mean across remaining derivatives increases, 
and the condition has to be checked again. Hence, we repeat checking until 
no more enzymes fulfill the
condition and the mean does not change any more.


## Vizualization

```{r include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(RColorBrewer) # brewer.pal
library(knitr)

mycolors <- brewer.pal(5,"Dark2")
mylinetypes <- c("solid","dashed","dotdash","twodash","dotted","longdash")
scale_colour_discrete <- function(...) scale_colour_manual(...,values=mycolors)

adjust_facets <- function(theme){
  theme +
    theme(panel.border = element_rect(colour = "black", fill = NA)) +
    theme(strip.background = element_blank()) 
}
themePaper <- (theme_classic(base_size = 10) + theme(axis.title = element_text(size = 9))) %>% 
  adjust_facets()
themePres <- theme_classic(base_size = 18) %>% 
  adjust_facets()

#figWidthPage <- 18/2.54; figHeightPage <- figWidthPage/1.6 
# GMD fixes at 12 instead of entire page 18
figWidthPage <- 12/2.54; figHeightPage <- figWidthPage/1.6 
#, fig.height=4.43, fig.width=7.1
#figWidthWide <- 12/2.54; figHeightWide <- figWidthWide/1.6
figWidthWide <- 3.8; figHeightWide <- 5
#, fig.height=2.95, fig.width=4.72
figWidthCol <- 8.3/2.54; figHeightCol <- figWidthCol/1.6
#, fig.height=2.04, fig.width=3.27
#
theme_set(themePres); figWidth <- figWidthWide; figHeight <- figHeightWide 
theme_set(themePaper); figWidth <- figWidthCol; figHeight <- figHeightCol


opts_chunk$set(
   # out.extra='style="display:block; margin: auto"'
    fig.width = figWidth, fig.height = figHeight #goldener Schnitt 1.6 #GI single column
    #, dev.args = list(pointsize = 11)
    , dev = c('pdf','png')
    , echo = FALSE
    )
knit_hooks$set(spar  =  function(before, options, envir) {
    if (before) {
        par( las = 1 )                   #also y axis labels horizontal
        par(mar = c(2.0,3.3,0,0) + 0.3 )  #margins
        par(tck = 0.02 )                          #axe-tick length inside plots             
        par(mgp = c(1.1,0.2,0) )  #positioning of axis title, axis labels, axis
    }
})

.approaches <- c("Optimal","Derivative","Relative")
colScaleApproach <- scale_colour_manual(
  name = "Approach", values = setNames(mycolors[seq_along(.approaches)], .approaches))
linetypeScaleApproach <- scale_linetype_manual(
  name = "Approach",values = setNames(mylinetypes[seq_along(.approaches)], .approaches))

```

### Revenues

```{r echo=FALSE}
# utility of depolymerizing enzyme
revenue_decomp <- function(alpha, dLorR, o_enz, p){
  dLorR/o_enz/(p$k_mn + alpha * p$a_E * p$B)
}
p = within(list(
  a_E = 0.1, 
  B=1, 
  e_P = 0,
  tau = 365/30,
  synB = 15
), k_mn <- a_E*B/2)
nrow <- 31
ds <- data.frame(
  alpha = seq(0,1,length.out=nrow),
  dL = 1,
  dR = 0.7,
  dP = seq(0,1,length.out=nrow),
  o_enz = 0.2
)
ds2 <- bind_rows(
  mutate(ds, Biomass="High", u_L = revenue_decomp(alpha, dL, o_enz,  within(p, B <-2))),
  mutate(ds, Biomass="Low", u_L = revenue_decomp(alpha, dL, o_enz,  within(p, B <-1)))
)
ds2 %>% ggplot(aes(alpha, u_L, color=Biomass, linetype=Biomass)) + geom_line() +
  xlab(bquote(~alpha[L])) +
  ylab(bquote('Revenue '*rev[L]*'('*g~ m^-2~s^-1*')')) +
  theme(legend.position = c(0.95,0.95), legend.justification = c(1,1)) +
  c()
```
```{r echo=FALSE}
# utility of biomineralizing enzyme
revenue_biomin <- function(alpha, dP, o_enz, p){
  dP*p$k_mn/o_enz/((p$k_mn + p$e_P)^2 + alpha * p$a_E * p$B * ((p$k_mn + p$e_P)))
}
ds2 <- bind_rows(
  mutate(ds, PlantProduction="none", u=revenue_biomin(alpha, dL, o_enz,  p)),
  mutate(ds, PlantProduction="same magnitude", u=revenue_biomin(alpha, dL, o_enz, within(p, e_P <- k_mn/2))),
)
ds2 %>% ggplot(aes(alpha, u, color=PlantProduction, linetype=PlantProduction)) + geom_line() +
  xlab(bquote(~alpha[P])) +
  ylab(bquote('Revenue '*rev[P]*'('*g~ m^-2~s^-1*')')) +
  theme(legend.position = c(0.95,0.95), legend.justification = c(1,1)) +
  c()
```

The utility of a depolymerizing enzmye, i.e. its return per enzyme, 
decreases with $\alpha$ and stronger
with larger microbial biomass.
The revenue of the biomineralizing enzyme is smaller when plant also produces 
enzymes but does not decrease as strongly.

### Derivatives of $\alpha$ times Revenues
```{r include=FALSE}
library(rootSolve)
# this is du/dalpha, i.e. o_enz*ae*B * d(alpha * u_Z)/dalpha
calc_du_dalpha <- function(alpha, dS, p, e_P =  p$e_P){
  aeB = p$a_E * p$B
  aeB * p$k_mn * dS / (e_P + p$k_mn + alpha*aeB)^2
}
fdu  <- function(alpha3, dL, dR, dP, p) c(
  L = calc_du_dalpha(alpha3[["L"]], dL, p, e_P = 0),
  R = calc_du_dalpha(alpha3[["R"]], dR, p, e_P = 0),
  P = calc_du_dalpha(alpha3[["P"]], dP, p)
  )
fdalpha = function(alpha3, dL, dR, dP, p){
  du = fdu(alpha3, dL, dR, dP, p)
  # only change alpha if its larger than zero or if the change is positive
  mean_du_prev <- -Inf; mean_du <- mean(du)
  is <- rep(TRUE, length(alpha3))
  while (mean_du_prev != mean_du) {
    is <- is & (alpha3 > 0 | du > mean(du))
    mean_du_prev <- mean_du
    mean_du <- mean(du[is])
  }
  du[!is] <- mean_du
  dud = (du - mean_du)/mean_du
  #u_total = compute_total_utilities(alpha3, dL, dR, dP, p)
  dalpha = (p$tau + abs(p$synB)/p$B) * dud
  dalpha
}
calc_alpha3_proptoderiv <- function(dL,dR, dP, p){
  alpha0 = alpha3 = c(L=1,R=1,P=1)/3 #compute_alpha3_relative(dL,dR,dP,p)
  #steady(alpha0, 0, function(t,alpha3,p){list(fdalpha(alpha3))})
  res_steady <- runsteady(alpha0, c(0,Inf), function(t,alpha3,p){
    list(fdalpha(alpha3,dL,dR,dP,p))}, p)
  alpha3 = setNames(pmax(0,res_steady$y), names(alpha0))
  alpha3 = alpha3/sum(alpha3)
  alpha3
}
```



```{r echo=FALSE}
# d alpha * revenue of depolymerizing enzyme
nrow <- 31
ds <- data.frame(
  alpha = seq(0,1,length.out=nrow),
  dL = 1
)
ds2 <- bind_rows(
  mutate(ds, Biomass="High", u_L = calc_du_dalpha(alpha, dL, within(p, B <-2), e_P = 0)),
  mutate(ds, Biomass="Low", u_L = calc_du_dalpha(alpha, dL, within(p, B <-1), e_P = 0))
)
ds2 %>% ggplot(aes(alpha, u_L, color=Biomass, linetype=Biomass)) + geom_line() +
  xlab(bquote(~alpha[L])) +
  #ylab(bquote(d(alpha[L]*rev[L])/d*alpha[L]*' ('*gC~ m^-2~s^-1*')')) +
  ylab(bquote(d~return/d*alpha[L]*' ('*g~ m^-2~s^-1*')')) +
  theme(legend.position = c(0.95,0.95), legend.justification = c(1,1)) +
  c()
```
```{r echo=FALSE}
# d alpha * revenue of biomineralizing enzyme
nrow <- 31
ds <- data.frame(
  alpha = seq(0,1,length.out=nrow),
  dP = 1
)
ds2 <- bind_rows(
  mutate(ds, PlantProduction="none", u_P = calc_du_dalpha(alpha, dP, p, e_P = 0)),
  mutate(ds, PlantProduction="same magnitude", u_P = calc_du_dalpha(alpha, dP, p, e_P = p$k_mn/2))
)
ds2 %>% ggplot(aes(alpha, u_P, color=PlantProduction, linetype=PlantProduction)) + geom_line() +
  xlab(bquote(~alpha[P])) +
  #ylab(bquote(d(alpha[P]*rev[P])/d*alpha[P]*' ('*g~ m^-2~s^-1*')')) +
  ylab(bquote(d~return/d*alpha[P]*' ('*g~ m^-2~s^-1*')')) +
  theme(legend.position = c(0.95,0.95), legend.justification = c(1,1)) +
  c()
```

Similar to the Revenues, the derivatives of (revenue times allocation), 
which are proportional to the derivatives of total return, 
decreases with increasing allocation.
They decrease stronger with higher biomass and less with plant enzyme
production.

```{r echo=FALSE}
# d alpha * revenue of two depolymerizing enzymes
nrow <- 31
ds <- data.frame(
  alpha = seq(0,1,length.out=nrow),
  dL = 2,
  dR = 1
)
ds2 <- bind_rows(
  mutate(ds, Enzyme="Labile", du_Z = calc_du_dalpha(alpha, dL, p, e_P = 0)),
  mutate(ds, Enzyme="Residue", du_Z = calc_du_dalpha(alpha, dR, p, e_P = 0))
)
du_R_max = max(filter(ds2,Enzyme=="Residue")$du_Z)
alpha3 = calc_alpha3_proptoderiv(ds$dL[1], ds$dR[1], 0, p)
du_equal = calc_du_dalpha(alpha3[["L"]], ds$dL[1], p, e_P = 0)
p1 <- ds2 %>% ggplot(aes(alpha, du_Z, color=Enzyme, linetype=Enzyme)) + 
  geom_hline(yintercept=du_R_max, linetype="dotted", color="gray") +
  geom_hline(yintercept=du_equal, linetype="dashed", color="gray") +
  geom_line() +
  xlab(bquote(~alpha[Z])) +
  #ylab(bquote(d(alpha[Z]*rev[Z])/d*alpha[Z]*' ('*gC~ m^-2~s^-1*')')) +
  #ylab(bquote(d~return/d*alpha[Z]*' ('*g~ m^-2~s^-1*')')) +
  ylab(bquote(d*u[T]/d*alpha[Z]*' ('*g~ m^-2~s^-1*')')) +
  theme(legend.position = c(0.95,0.95), legend.justification = c(1,1)) +
  c()
print(p1)
```
```{r eval=FALSE, include=FALSE}
ggsave("sesamalloc_derivative.pdf", p1, width = figWidthCol, height = figHeightCol)
```

Although potential return is higher for labile litter degrading enzyme 
($d_L = 2 gm^{-2}yr^{-1}$) 
than residue degrading enzyme ($d_R = 1 gm^{-2}yr^{-1}$) 
optimal allocation includes both enzymes in the mix.
When increasing $\alpha_L > 0.25$, is is not benefitial to invest
only into the labile enzyme, because the increase in return with 
further increasing $\alpha_L$ is lower than with increasing $\alpha_R$.
Optimal allocation is attained when both derivatives are equal and allocation
shares add up to one, which happens here at allocation about 1/3 to residue
depolymerizing enzyme and 2/3 to labile depolymerizing enzyme.


### Allocation

On prescribed potential returns ($d_L = 0.7, d_R = 0.5, d_P = \{0\ldots1\}$),
The inferred optimum allocation differs across approaches.

```{r include=FALSE}
compute_revenues <- function(alpha3,dL,dR,dP,p,o_enz=0.2){
  c(
      revenue_decomp(alpha3["L"], dL, o_enz, p), 
      revenue_decomp(alpha3["R"], dR, o_enz, p), 
      revenue_biomin(alpha3["P"], dP, o_enz, p)
    )
}
compute_total_utilities <- function(alpha3,dL,dR,dP,p,o_enz=0.2){
  u = compute_revenues(alpha3,dL,dR,dP,p,o_enz)
  o_enz*p$a_E*p$B*sum(alpha3 * u)
}
calc_alpha3_relative = function(dL,dR,dP,p,o_enz=0.2) {
  u0 = u1 = sapply(c(L=dL, R=dR, P=dP), function(x) x/(o_enz*p$k_mn))
  # fixpoint iteration
  for (i in 1:5) {
    alpha3 = u1/sum(u1)
    u1 = compute_revenues(alpha3,dL,dR,dP,p,o_enz)
    #print(c(u1, sum(alpha3*u1)))
  }
  alpha3
}
dL = 0.7; dR = 0.5; dP = 0.1
calc_alpha3_relative(dL,dR,dP,p)

dL = 0.7; dR = 0.5; dP = 0.3
#dL = 0.7; dR = 0.5; dP = 0.01
#alpha3 <- c(L=0.5652477, R=0.4572952, P=-0.02254293)
dalpha <- fdalpha(c(L=1,R=1,P=1)/3, dL, dR, dP, p)
alpha3 = calc_alpha3_proptoderiv(dL, dR, dP, p)
alpha3
fdu(alpha3, dL, dR, dP, p)  # all equal
alpha3r = calc_alpha3_relative(dL, dR, dP, p)
alpha0 = c(L=1,R=1,P=1)/3 #compute_alpha3_relative(dL,dR,dP,p)
c(alpha0 = compute_total_utilities(alpha0, dL, dR, dP, p), 
  alpha3r = compute_total_utilities(alpha3r, dL, dR, dP, p), 
  alpha3 = compute_total_utilities(alpha3, dL, dR, dP, p)
  )
```

```{r eval=FALSE, include=FALSE}
# Inspect specific case 
dL = 0.7; dR = 0.1; dP = 0.3; pB <- within(p, B <- 1)
dL = 0.5; dR = 0.4; dP = 0.0; pB <- within(p, B <- 0.5)
dL = 0.7; dR = 0.5; dP = 0.3; pB <- within(p, B <- 1)


dalpha <- fdalpha(c(L=1,R=1,P=1)/3, dL, dR, dP, pB)
alpha3 = calc_alpha3_proptoderiv(dL, dR, dP, pB)
alpha3
fdu(alpha3, dL, dR, dP, pB)  # they are all equal
compute_total_utilities(alpha3, dL, dR, dP, pB)
# deviation has smaller utility
alpha3p <- alpha3 + c(-0.01, +0.01, 0)
u = c(compute_total_utilities(alpha3p, dL, dR, dP, pB), compute_total_utilities(alpha3, dL, dR, dP, pB) )
u[1] - u[2]
```

```{r include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(forcats)
```

```{r echo=FALSE, fig.width=figWidthPage, fig.height=figHeightPage}
dL = 0.7; dR = 0.5; dP = 0.3
unit = 1000
unit = 1
#dL = 0.7; dR = 0.5; dP = 0
calc_alpha3_proptoderiv(dL, dR, dP, p)
calc_alpha3_relative(dL*unit, dR*unit, dP*unit, p, o_enz=1)
dPs = seq(0,1,length.out=51)[-1]
dsAlpha_prop = cbind(dP=dPs, bind_rows(lapply(dPs, function(dP){calc_alpha3_proptoderiv(dL, dR, dP, p)})))
dsAlpha_prop_highB = cbind(dP=dPs, bind_rows(lapply(dPs, function(dP){calc_alpha3_proptoderiv(dL, dR, dP, within(p, B <- 5))})))
dsAlpha_rel = cbind(dP=dPs, bind_rows(lapply(dPs, function(dP){calc_alpha3_relative(dL, dR, dP, p)})))
dsAlpha = bind_rows(
  cbind(Approach = "Derivative", Biomass="Moderate", dsAlpha_prop),
  #cbind(Approach = "Deriv+high biomass", Biomass="High", dsAlpha_prop_highB),
  cbind(Approach = "Derivative", Biomass="High", dsAlpha_prop_highB),
  cbind(Approach = "Relative", Biomass="Moderate", dsAlpha_rel)
) %>% 
  mutate(Approach = fct_relevel(Approach, "Derivative"))  
p2 <- pivot_longer(dsAlpha, !c(dP,Approach,Biomass), names_to="Enzyme") %>% 
  mutate(Enzyme = fct_relevel(Enzyme, "L","R")) %>% 
  #ggplot(aes(dP, value, color=Enzyme, linetype=Approach)) + geom_line() +
  ggplot(aes(dP, value, color=Enzyme, linetype=Approach, size=Biomass)) + geom_line() +
  xlab(bquote('Potential return from biomineralization ('*g~ m^-2~s^-1*')')) +
  ylab(bquote('Allocation '*alpha)) +
  #theme(axis.title.y = element_blank()) +
  c()
.values <- c(
  setNames(mylinetypes[seq_along(.approaches)], .approaches), 
  "Deriv+high biomass" = mylinetypes[which(.approaches == "Derivative")]
)
p2 <- p2 + scale_linetype_manual("Approach", values = .values) +
  scale_size_manual("Biomass", values = c(High=1.2,Moderate=0.8))
print(p2)  
```

```{r eval=FALSE, include=FALSE}
ggsave("sesamalloc_prescribed_returns.pdf", p2, width = figWidthPage, height = figHeightPage*1.1)
```

The Realtive approach allocates larger shares to less proficient enzymes
at low biomass. The Optimal approach only does this, if microbial biomass
levels are high.

```{r echo=FALSE}
#trace(compute_total_utilities, recover) # untrace(compute_total_utilities)
dsU <- dsAlpha %>%
  rowwise() %>% 
  mutate(u = compute_total_utilities(c(L=L, R=R, P=P), dL, dR, dP, p)) %>% 
  select(Approach, Biomass, dP, u)
ggplot(dsU, aes(dP,u, linetype=Approach, color=Biomass)) + geom_line() +
  scale_size_manual("Biomass", values = c(High=1.2,Moderate=0.8))
```

While allocations differ, the total returns are similar across approaches.
The differences are most pronounced near situations where 
it is not profitable to invest into an enzyme at all.


```{r eval=FALSE, include=FALSE}
# plot the difference - does not add to plotting both utilities
library(purrr)
calc_diff = function(.data){
  filter(.data, Approach=="Derivative")$u[1] - filter(.data, Approach=="Relative")$u[1]
}
dsUdiff <- dsU %>% group_by(dP) %>% nest() %>% mutate(udiff = map_dbl(data, calc_diff))
ggplot(dsUdiff, aes(dP, udiff)) + geom_line() + ylim(0,NA)
```


```{r eval=FALSE, include=FALSE}
# The time scale of adaptation
# is about one month, as expected.
# This is shown by inspect the time course of the solved system. 
library(deSolve)
res_ode <- lsode(alpha0, seq(0,0.5,length.out=201), function(t,alpha3,p){list(fdalpha(alpha3,dL,dR,dP,p))}, p)
plot(res_ode)
```

