In order to understand the coupling of carbon (C) and nitrogen (N) cycles, it is
necessary to understand C and N-use efficiencies of microbial soil organic
matter (SOM) decomposition. While important controls of those efficiencies by
microbial community adaptations have been shown at the scale of a soil pore, an
abstract simplified representation of community adaptations is needed at
ecosystem scale.

Therefore we developed the soil enzyme allocation model (SEAM), which takes a
holistic, partly optimality based approach to describe C and N dynamics at the
spatial scale of an ecosystem and time-scales of years and longer. We explicitly
modelled community adaptation strategies of resource allocation to extracellular
enzymes and enzyme limitations on SOM decomposition. Using SEAM, we explored
whether alternative strategy-hypotheses can have strong effects on SOM and
inorganic N cycling.

Results from prototypical simulations and a calibration to observations of an
intensive pasture site showed that the so-called revenue enzyme allocation
strategy was most viable. This strategy accounts for microbial adaptations to
both, stoichiometry and amount of different SOM resources, and supported the
largest microbial biomass under a wide range of conditions. Predictions of the
holistic SEAM model were qualitatively similar to precitions of the SYMPHONY
model, which explicitly represents competing microbial guilds. With adaptive enzyme
allocation under conditions of high C/N ratio of litter inputs, N that was
formerly locked in slowly degrading SOM pools was made accessible, whereas with
high N inputs, N was sequestered in SOM and protected from leaching.

The findings imply that it is important for ecosystem scale models to account
for adaptation of C and N use efficiencies in order to represent C-N couplings.
The combination of stoichiometry and optimality principles is a promising route
to yield simple formulations of such adaptations at community level suitable for
incorporation into land surface models.
