---
title: "SESAM Soil enzyme steady allocoation model"
output:
  pdf_document:
    keep_tex: true
  html_notebook: default
---

Start from Sesam3Overview.

## Comparison using the FACE scenario
```{r, echo=FALSE, results='hide', message=FALSE}
library(ggplot2)
library(grid)   #unit
library(RColorBrewer) # brewer.pal
library(tidyr)
library(dplyr)

myColors <- brewer.pal(5,"Dark2")	
.simpleCap <- function(s) {
    paste(toupper(substring(s, 1, 1)), substring(s, 2),  sep = "")
}
colScale <- scale_colour_manual(name = "Variant",values = myColors)
.treatments <- structure(c(1,0.5), names = c("Litter input pulse","No Litter input"))
sizeScale <- scale_size_manual(name = "Treatment",values = .treatments)

themeDefault <- theme_classic()  
themePaper <- theme_classic(base_size = 10)  
```

### Setup parameters and drivers
Mass units are $gC/m^2$ or $gN/m^2$. Time unit is $yr$.
```{r}
require(sesam, quietly = TRUE)

parmsModel <- list(
  cnB = 11
  ,cnE = 3.1     # Sterner02: Protein (Fig. 2.2.), high N investment (low P) need 
    # to have low C/N for CO2 increase scenario
  ,kN = 60       ##<< /yr enzyme turnover 60 times a year, each 6 days -> fast 
    # priming pulses
  ,km = 0.05     ##<< enzyme half-saturation constant, in magnitude of enzymes, 
    # determined by kN
  ,kNB = 0.8     ##<< amount of recycling enzyme turnover by biomass (
    # added to assimilable, i.e. uptake instead of R)
  #,kR = 1/(10)        ##<< 1/(x years) # to demonstrate changes on short time scale
  ,kR = 1/(20)        ##<< 1/(x years) # to demonstrate changes on short time scale
  ,kL = 5           ##<< 1/(x years)   # formerly 1 year
  ,aE = 0.001*365   ##<< C biomass allocated to enzymes 1/day /microbial biomass
  ,m = 0.016*365    ##<< maintenance respiration rate   1/day /microbial biomass, 
  ## Bogedom Fig. 1
  ,tau = 1/60*365  ##<< biomass turnover rate (12 days)
  ,eps = 0.68      ##<< carbon use efficiency for growth respiration
  #,eps = 0.5      ##<< carbon use efficiency for growth respiration
  #,epsTvr = 0.3   ##<< carbon use efficiency of microbial tvr (part by predators 
  ,epsTvr = 0.45   ##<< carbon use efficiency of microbial tvr (part by predators 
  ## which respire and corresponding amount of N must be mineralized)
  #,iB = 0.38 * 10.57 #0.0110068*365   ##<< potential immobilization flux rate 
  ,iB = 0.4 ##<< potential immobilization flux rate 
  #,iB = 1/12*(24*365) # tvr of I of 12 hours
  ## (immoPot = iB I)
  ,l = 0.96       #0.00262647*365     ##<< leaching rate of mineralN l I
  ,nu = 0.9     # microbial N use efficiency accounting for apparent 
  ## minceralization of N during uptake in heterogeneous soils
  , isEnzymeMassFlux = TRUE  ##<< compute enzyme mass fluxes
  , kIPlant = 10.57 #0.0289652*365          ##<< plant uptake iP I
)

parmsModel <- within(parmsModel, {
    kmN <- km*kN     # Sesam2 only uses a lumped parameter instead of km and kN
    kmR <- kmL <- km  # seam2 can accept parameters differing by substrate
    cnER <- cnEL <- cnE 
    kNR <- kNL <- kN    
    #isFixedI <- TRUE  # in this demonstration keep inorganic pool fixed
})
```

For simplicity model drivers currently are prescibed as constant parameters.
```{r}
drivers <- list(
        iR = 0          ##<< input of residue litter
        ,iL = 400         # g/m2 input per year (half NPP)
        ,cnIR = 7       ##<< C/N ratio of the residue litter input
        ,cnIL = 30      ##<< C/N ratio of the labile litter input: 
        ,iI = 0         ##<< input of mineral N
)
parms0 <- within( c(parmsModel, drivers), {
  iI <- iL/cnIL/200    # assume a fractionof 1/10 of annual litter production
  kIPlant <- 1/5*365 # tvr of 5 days
  plantNUpAbs <- iL/cnIL + iI # plants take maximum out as much as returned by litter
  # plus inorganic input, difference is grazed
  # keep leaching as otherwise it accumulates during priming
  #l <- 0 # no leaching
})
parms0Noenz <- within(parms0, isEnzymeMassFlux <- FALSE)
```

The initial pools must be specified. The Steady models do not have the enzyme pools as state variables.
```{r}
x0 <- x0Explicit <- c( #aE = 0.001*365
        B = 17                     ##<< microbial biomass 
        ,ER  = 2*parms0$km                  ##<< total enzyme pool
        ,EL  = 4*parms0$km                   ##<< total enzyme pool
        ,R = 1100                ##<< N rich substrate
        ,RN = 1100/parms0$cnIR   ##<< N rich substrate N pool
        ,L = 100                 ##<< N poor substrate
        ,LN = 100/parms0$cnIL    ##<< N poor substrate N pool
        ,I =  0.04 #150 #0.04                ##<< inorganic pool gN/m2
        , alpha = 0.5            ##<< microbial community partitioning [0,1]
)
x0Steady <- x0Explicit[c("B","R","RN","L","LN","I")]
x0SteadyA <- c(x0Steady,alpha = 0.5)
x0B <- x0SteadyA[setdiff(names(x0SteadyA),"B")]
```

We store the information that differs between scenarios in a tibble.
```{r}
scenarios <- tibble::tibble(
  scenario = c("Explicit","Sesam3a","NeglectEnzFlux","Sesam3B")
  ,mod = c(derivSeam3a, derivSesam3a, derivSesam3a, derivSesam3B)
  ,x0  = list(x0Explicit, x0SteadyA, x0SteadyA, x0B)
  ,parms = list(parms0, parms0, parms0Noenz, parms0)
)
tmp <- structure(myColors[1:nrow(scenarios)], names = scenarios$scenario)
colScale <- scale_colour_manual(name = "Model variant",values = tmp )
```

Next the simulation runs are performed.
```{r}
runFaceSimulation <- function(
  scen, t1S =20, t2I = 40, t3S = 70, parms = parms0
){
  # spinup to steady state
  times <- seq(0,1000, length.out = 2)
  #times <- seq(0,5, length.out = 500)
  #setwd("../..");loadPkg()
  . <- scen
  res <- res1 <- as.data.frame(lsoda( .$x0, times, .$mod, parms = .$parms))
  #ggplot(res, aes(time, I)) + geom_line()
  xE <- unlist(tail(res,1))
  #xE["I"] <- 0.04  # decrease I pool # does not help filled up by mircobial tvr
  #
  # t1S years of steady state recorded
  #derivSesam3a(0, xE, .$parms)
  res <- res1S <- as.data.frame(lsoda( 
    xE[1:length(.$x0) + 1], 1:t1S, .$mod, parms = .$parms))
  #
  # t2I years of increased input
  fInputInc = 1.2
  parmsC2 <- within(.$parms, {# more C in litter, but not more N 
    iL <- iL*fInputInc
    cnIL <- cnIL*fInputInc
  }) 
  derivSesam3a(0, xE, parmsC2)
  times <- seq(0,t2I, length.out = t2I*6 + 1)# record bymonthly
  res <- res2I <- as.data.frame(lsoda( 
    xE[1:length(.$x0) + 1], times, .$mod, parms = parmsC2))
  res2I$time <- res2I$time + t1S 
  xE2 <- unlist(tail(res2I,1))
  #
  # t3S years of initial input
  times <- seq(0,t3S, length.out = t3S*6 + 1) # record bymonthly
  res <- res3S <- as.data.frame(lsoda( 
    xE2[1:length(.$x0) + 1], times, .$mod, parms = .$parms))
  res3S$time <- res3S$time + t1S + t2I
  xE3 <- tail(res3S,1)
  resc <- rbind( res1S, res2I[-1,], res3S[-1,])
  resc$scen <- scen$scenario
  resc
}
.tmp.f <- function(){
  scen <- unlist(filter(scenarios, scenario == "Explicit"), recursive = FALSE)
  scen <- unlist(filter(scenarios, scenario == "Sesam3a"), recursive = FALSE)
  scen <- unlist(filter(scenarios, scenario == "Sesam3B"), recursive = FALSE)
  resScen <- runFaceSimulation( scen )
}
resScen <- scenarios %>% 
  #filter(scenario == "Sesam3a") %>% 
  rowwise() %>% do( runFaceSimulation(.) )
```


```{r}
vars <- c("L","R","alpha","RL") 
vars <- c("L","R","alpha","B","RL","I") 
vars <- c("L","R","alpha","B","RL","I","plantNUp","PhiB","PhiTotal") 
#vars <- c("alpha","B", "cnL", "PhiB","PhiTotal","immoPot", "CsynBC","CsynBN") 
dsp <- resScen %>% 
  filter(scen != "NeglectEnzFlux") %>% 
  #filter(scen == "Sesam3a") %>% 
  #filter(time >= 20, time <= 40) %>% 
  mutate(RL = R + L, cnRL = (R + L)/(RN + LN)) %>% 
  #select_(~time, ~scen, ~L, ~R, ~RL) %>% gather_("Pool","value", c("L","R","RL"))
  select(one_of(c("time","scen",vars))) %>% gather("Pool","value", vars)
p2f <- ggplot( dsp, aes(x = time, y = value, col = scen, linetype = scen)) + geom_line(size = 1) +
  facet_grid(Pool ~ .,scales = "free") + 
  xlab("Time (yr)") + ylab(expression(Carbon~stock~(gCm^{-2}))) + #labs(linetype = "Substrate pool") +
  themeDefault +
  #scale_colour_discrete(drop = TRUE,limits = levels(dsp$Allocation)) +
  theme(legend.position = c(0.95,1.04), legend.justification = c(1,1)) +
  theme(panel.border = element_rect(colour = "black", fill = NA)) +
  theme(strip.background = element_blank()) +			
  theme(panel.grid.major.x = element_line(colour = "grey75")) +
  c()
#print(p2f + colScale)
print(p2f)
```
Why is Sesam3B differnt at the end of FACE?
```{r}
dst <- filter(resScen, scen %in% c("Sesam3a","Sesam3B"), time %in% c(20,60))
#total N balance
dst <- dst %>% mutate(NSys = B/parms0$cnB + LN + RN + I)
select(dst, scen, time, NSys, I)

dst <- filter(resScen, scen %in% c("Sesam3a","Sesam3B"), time >= 20, time <= 60)
ggplot(dst, aes(time, I, color = scen)) + geom_line()
```
There is a slight increase of N in the systme due to change
in N in microbial biomass is not accounted in N balance.

Why do I need such a low iB?
The potential immobilization must be low to shift the community to N limitation.
If its increased (observed I turnovertime of within a day) microbes never
become N limited. Is this reasonable?

## Result

Model Sesam3a matches the dynamics of the explicit SEAM model at this time scale.
When neglecting enzyme mass fluxes, the dynamcis is similar, but the steady state 
, expecially for R, is slightly lower.

The model with steady state microbial biomass, Sesam3B, captures the same dynamics.

```{r simIncreasedC, fig.height=3, fig.width=3.27, fig.height=3.2, fig.width=4.72}
# produce figure for paper without alpha, ... 
vars <- c("L","R","SOM") 
#vars <- c("L","R","SOM","cnRL") 
#vars <- c("L","R","alpha","B","RL","cnRL") 
#vars <- c("alpha","B", "cnL", "PhiB","PhiTotal","immoPot", "CsynBC","CsynBN") 
dsp <- ungroup(resScen) %>% mutate(SOM = R + L, cnRL = (R + L)/(RN + LN)) %>% 
  #select_(~time, ~scen, ~L, ~R, ~RL) %>% gather_("Pool","value", c("L","R","RL")) %>% 
  mutate(Model = recode_factor(
    as.factor(scen)
    , Explicit = "Seam3"
    , Sesam3a = "Sesam3"
    , Sesam3B = "Sesam3_SteadyB"
    , NeglectEnzFlux = "Sesam3_NoEnzFlux"
    )
    # delete CNRL of EnzFlux because at other scale
    ,cnRL = ifelse(scen == "NeglectEnzFlux", NA, cnRL)
    ) %>% 
  #recode_factor(vars, Explicit = "Seam3") %>% 
  select(one_of(c("time","Model",vars))) %>% 
  gather("Pool","value", vars) 
p2f <- ggplot( dsp, aes(x = time, y = value, col = Model, linetype = Model)) + geom_line(size = 1) +
  facet_grid(Pool ~ .,scales = "free") + 
  xlab("Time (yr)") + ylab(expression(Carbon~stock~(gCm^{-2}))) + #labs(linetype = "Substrate pool") +
  themePaper +
  #scale_colour_discrete(drop = TRUE,limits = levels(dsp$Allocation)) +
  #theme(legend.position = c(0.8,0.9), legend.justification = c(1,1)) +
  theme(legend.position = c(0.98,0.02), legend.justification = c(1,0)) +
  theme(panel.border = element_rect(colour = "black", fill = NA)) +
  theme(strip.background = element_blank()) +			
  theme(panel.grid.major.x = element_line(colour = "grey75")) +
  c()
#print(p2f + colScale)
print(p2f)


```
```{r}

```


