---
title: "SESAM CUE in substrate pulse"
output: 
  pdf_document:
    keep_tex: true
  html_notebook: default
---

## Purpose
Kaiser14 showed emergent CUE stabilized during substrate pulse by
microbial competition.
Here we simulate a similar scenario using Sesam3a

```{r, echo=FALSE, results='hide', message=FALSE}
library(ggplot2)
library(grid)   #unit
library(reshape)  # melt
library(RColorBrewer) # brewer.pal
library(tidyr)
library(dplyr)
library(knitr)

myColors <- brewer.pal(5,"Dark2")	
.simpleCap <- function(s) {
    paste(toupper(substring(s, 1, 1)), substring(s, 2),  sep = "")
}
colScale <- scale_colour_manual(name = "Variant",values = myColors)
.treatments <- structure(c(1,0.5), names = c("Litter input pulse","No Litter input"))
sizeScale <- scale_size_manual(name = "Treatment",values = .treatments)

themeTw <- theme_bw(base_size = 10) + theme(axis.title = element_text(size = 9))
#themeTw <- theme_bw(base_size = 11) + theme(axis.title = element_text(size = 10))
figWidthPage <- 18/2.54; figHeightPage <- figWidthPage/1.6 
#, fig.height=4.43, fig.width=7.1
figWidthWide <- 12/2.54; figHeightWide <- figWidthWide/1.6
#, fig.height=2.95, fig.width=4.72
figWidthCol <- 8.3/2.54; figHeightCol <- figWidthCol/1.6
#, fig.height=2.04, fig.width=3.27
opts_chunk$set(
   # out.extra='style="display:block; margin: auto"'
    #, fig.align="center"
    #, fig.width=4.6, fig.height=3.2
    #fig.width = 6, fig.height = 3.75 #goldener Schnitt 1.6
    #fig.width = figWidthWide, fig.height = figHeightWide #goldener Schnitt 1.6 #GI single column
    fig.width = figWidthPage, fig.height = figHeightPage #goldener Schnitt 1.6 #GI single column
    #, dev.args = list(pointsize = 11)
    , dev = c('pdf','png')
    , echo = FALSE
    )
knit_hooks$set(spar  =  function(before, options, envir) {
    if (before) {
        par( las = 1 )                   #also y axis labels horizontal
        par(mar = c(2.0,3.3,0,0) + 0.3 )  #margins
        par(tck = 0.02 )                          #axe-tick length inside plots             
        par(mgp = c(1.1,0.2,0) )  #positioning of axis title, axis labels, axis
    }
})

themeDefault <- themeTw#theme_classic()  
themePres <- theme_classic()
```


### Setup parameters and drivers
Mass units are $gC/m^2$ or $gN/m^2$. Time unit is $yr$.
```{r}
require(sesam, quietly = TRUE)

parmsModel <- list(
  cnB = 11
  ,cnE = 3.1     # Sterner02: Protein (Fig. 2.2.), high N investment (low P) need 
    # to have low C/N for CO2 increase scenario
  ,kN = 60       ##<< /yr enzyme turnover 60 times a year, each 6 days -> fast 
    # priming pulses
  ,km = 0.05     ##<< enzyme half-saturation constant, in magnitude of enzymes, 
    # determined by kN
  ,kNB = 0.8     ##<< amount of recycling enzyme turnover by biomass (
    # added to assimilable, i.e. uptake instead of R)
  #,kR = 1/(10)        ##<< 1/(x years) # to demonstrate changes on short time scale
  ,kR = 1/(20)        ##<< 1/(x years) # to demonstrate changes on short time scale
  ,kL = 5           ##<< 1/(x years)   # formerly 1 year
  ,aE = 0.001*365   ##<< C biomass allocated to enzymes 1/day /microbial biomass
  ,m = 0.005*365    ##<< maintenance respiration rate   1/day /microbial biomass, 
  ## Bogedom Fig. 1
  ,tau = 1/60*365  ##<< biomass turnover rate (12 days)
  ,eps = 0.5      ##<< carbon use efficiency for growth respiration
  #,epsTvr = 0.3   ##<< carbon use efficiency of microbial tvr (part by predators 
  ,epsTvr = 0.45   ##<< carbon use efficiency of microbial tvr (part by predators 
  ## which respire and corresponding amount of N must be mineralized)
  #,iB = 0.38 * 10.57 #0.0110068*365   ##<< potential immobilization flux rate 
  ,iB = 0.4 ##<< potential immobilization flux rate 
  ## (immoPot = iB I)
  ,l = 0.96       #0.00262647*365     ##<< leaching rate of mineralN l I
  ,nu = 0.9     # microbial N use efficiency accounting for apparent 
  ## minceralization of N during uptake in heterogeneous soils
  , isEnzymeMassFlux = TRUE  ##<< compute enzyme mass fluxes
  ,kIPlant = 10.57 #0.0289652*365     ##<< plant uptake rate first order of I
)

parmsModel <- within(parmsModel, {
  kmN <- km*kN     # Sesam2 only uses a lumped parameter instead of km and kN
  kmR <- kmL <- km  # seam2 can accept parameters differing by substrate
  cnER <- cnEL <- cnE 
  kNR <- kNL <- kN    
  #isFixedI <- TRUE  # in this demonstration keep inorganic pool fixed
})
```

For simplicity model drivers currently are prescibed as constant parameters.
```{r}
drivers <- list(
  iR = 0          ##<< input of residue litter
  ,iL = 40 #400         # g/m2 input per year (half NPP)
  ,cnIR = 7       ##<< C/N ratio of the residue litter input
  ,cnIL = 30      ##<< C/N ratio of the labile litter input: 
  ## N poor substrate, here no inorganic N supply, 
  ,plantNUp = 0   ##<< pland uptake of organic N
  ## currently as absolute flux
  ,iI = 0         ##<< input of mineral N
)
parms0 <- within( c(parmsModel, drivers), {
  # change inorganic pool only by mobilization/immobilization flux
  kIPlant <- 0
  plantNUpAbs <- 0#iL/cnIL
  l <- 0 
  #isFixedI <- TRUE
})
parms0Noenz <- within(parms0, isEnzymeMassFlux <- FALSE)

parmsFixedAlpha <- within(parms0, isFixedAlpha <- TRUE)
```

The initial pools must be specified. The Steady models do not have the enzyme pools as state variables.
```{r}
x0 <- x0Explicit <- c( #aE = 0.001*365
        B = 17                     ##<< microbial biomass 
        ,ER  = 2*parms0$km                  ##<< total enzyme pool
        ,EL  = 4*parms0$km                   ##<< total enzyme pool
        ,R = 1100                ##<< N rich substrate
        ,RN = 1100/parms0$cnIR   ##<< N rich substrate N pool
        ,L = 10#100                 ##<< N poor substrate
        ,LN = 10/parms0$cnIL #100/parms0$cnIL    ##<< N poor substrate N pool
        ,I =  0.04                ##<< inorganic pool gN/m2
        , alpha = 0.5            ##<< microbial community partitioning [0,1]
)
x0Steady <- x0Explicit[c("B","R","RN","L","LN","I")]
x0SteadyA <- c(x0Steady,alpha = 0.5)
x0B <- x0SteadyA[setdiff(names(x0SteadyA),"B")]
```

Simulate steady state with low L inputs and then L pulse by changed L state.
```{r}
simLPulse <- function(x0, parms, cn0=75){
  # # spinup to steady state at low substrate inputs
  # times <- seq(0,2000, length.out = 2)
  # #times <- c(0,exp(seq(log(0.02),log(2000), length.out=100)))
  # #setwd("../..");loadPkg()
  # res <- res1 <- as.data.frame(lsoda( x0, times, derivSesam3a, parms = parms))
  # #plot(R~time, res)
  # xE <- unlist(tail(res,1))
  # #xE["I"] <- 0.04  # decrease I pool # does not help filled up by mircobial tvr
  #
  # increase L, no L input any more
  parms$iL <- 0
  x0P = x0 #xE[1:length(x0) + 1]
  x0P["B"] <- .3
  x0P["L"] <- 100
  x0P["LN"] <- 100/cn0
  x0P["R"] <- 1200
  x0P["RN"] <- 1200/parms$cnB
  x0P["I"] <- 1
  timesP <- c(0,exp(seq(log(0.002),log(10), length.out = 200)))
  res <- resP <- as.data.frame(lsoda( x0P, timesP, derivSesam3a, parms = parms))
}
resVarAlpha <- simLPulse(x0SteadyA, parms0, cn0 = 50)
resVarAlphaCN60 <- simLPulse(x0SteadyA, parms0, cn0 = 70)
resVarAlphaCN90 <- simLPulse(x0SteadyA, parms0, cn0 = 90)
x0FixedAlpha <- x0SteadyA; x0FixedAlpha["alpha"] <- 0.3
resAlpha30  <- simLPulse(x0FixedAlpha, parmsFixedAlpha, cn0 = 50)
resAlpha30CN60  <- simLPulse(x0FixedAlpha, parmsFixedAlpha, cn0 = 70)
resAlpha30CN90  <- simLPulse(x0FixedAlpha, parmsFixedAlpha, cn0 = 90)
# x0Alpha10 <- x0SteadyA; x0Alpha10["alpha"] <- 0.1
# resAlpha10  <- simLPulse(x0Alpha10, parmsFixedAlpha)
# x0Alpha90 <- x0SteadyA; x0Alpha90["alpha"] <- 0.6
# resAlpha90  <- simLPulse(x0Alpha90, parmsFixedAlpha)
resP <- rbind(
  cbind(scen = "adaptable", cn0 = 50, resVarAlpha)
  ,cbind(scen = "fixed community", cn0 = 50, resAlpha30)
  ,cbind(scen = "adaptable", cn0 = 70, resVarAlphaCN60)
  ,cbind(scen = "fixed community", cn0 = 70, resAlpha30CN60)
  ,cbind(scen = "adaptable", cn0 = 90, resVarAlphaCN90)
  ,cbind(scen = "fixed community", cn0 = 90, resAlpha30CN90)
  #,cbind(scen = "alpha=0.1", resAlpha10)
  #,cbind(scen = "alpha=0.9", resAlpha90)
)
```


```{r}
vars <- c("L","R","alpha","B", "cueDef", "cueSyn") 
#vars <- c("L","R","alpha","B", "I","cueDef", "cueSyn") 
#vars <- c("L","R","alpha","B","RL","cnRL") 
#vars <- c("alpha","B", "cnL", "PhiB","PhiTotal","immoPot", "CsynBC","CsynBN") 
dsp0 <- resP %>% #filter(.data$time < 5) %>%  
  rename(uptakeC = uC, respTotal = resp) %>% mutate(tvrBPred = 0.0) %>% 
  cueSesam() %>% 
  mutate(cn0 = as.factor(cn0)) %>% 
  mutate(S = R + L, cnS = (R + L)/(RN + LN)) %>% 
  mutate(relS = S/S[1], relL = L/L[1], fracDecomp = 1 - relL) %>% 
  #select_(~time, ~scen, ~L, ~R, ~RL) %>% gather_("Pool","value", c("L","R","RL"))
  select(one_of(c("time","relS","relL","fracDecomp","scen","cn0", vars))) 
dsp <- dsp0 %>% gather("Pool","value", vars)
p2f <- ggplot( dsp, aes(x = time, y = value, col = cn0, linetype = cn0)) + geom_line(size = 1) +
  facet_grid(Pool ~ scen,scales = "free") + 
  xlab("Time (yr)") + 
  #xlab("Substrate (proportion of initial)") + scale_x_reverse() +
  #xlab("Fraction C decomposed") +
  #ylab(expression(Carbon~stock~(gCm^{-2}))) + #labs(linetype = "Substrate pool") +
  themeDefault +
  #scale_colour_discrete(drop = TRUE,limits = levels(dsp$Allocation)) +
  theme(legend.position = c(0.95,1.04), legend.justification = c(1,1)) +
  theme(panel.border = element_rect(colour = "black", fill = NA)) +
  theme(strip.background = element_blank()) +			
  theme(panel.grid.major.x = element_line(colour = "grey75")) +
  c()
#print(p2f + colScale)
print(p2f)
```

```{r}
#vars2 <- c("cueDef", "cueSyn") 
vars2 <- c("cueDef") 
dsp2 <- dsp0 %>% select(one_of(c("time","relS","relL","fracDecomp","scen","cn0", vars2))) %>% gather("Pool","value", vars2)
p2 <- ggplot( dsp2, aes(x = fracDecomp, y = value, col = cn0, linetype = cn0)) + 
#p2 <- ggplot( dsp2, aes(x = fracDecomp, y = value, col = cn0, linetype = cn0)) + 
  geom_line(size = 1) +
  facet_grid(~ scen) + #,scales = "free") + 
  #xlab("Time (yr)") + 
  #xlab("Substrate (proportion of initial)") + scale_x_reverse() +
  xlab("Fraction C decomposed") +
  ylab("Carbon use efficiency") +
  expand_limits(y = 0.1) +
  #themeDefault +
  theme_classic(base_size = 18) + # presentation
  #scale_colour_discrete(drop = TRUE,limits = levels(dsp$Allocation)) +
  #theme(legend.position = c(0.95,1.04), legend.justification = c(1,1)) +
  theme(legend.position = c(0.95,0.05), legend.justification = c(1,0)) +
  theme(panel.border = element_rect(colour = "black", fill = NA)) +
  theme(strip.background = element_blank()) +			
  #theme(panel.grid.major.x = element_line(colour = "grey75")) +
  c()
#print(p2f + colScale)
print(p2)
```

```{r}
#vars2 <- c("cueDef", "cueSyn") 
vars2 <- c("cueDef") 
dsp2 <- dsp0 %>% select(one_of(c("time","relS","relL","fracDecomp","scen","cn0", vars2))) %>% gather("Pool","value", vars2)
p2 <- dsp2 %>% filter(time <= 3) %>% 
  ggplot( aes(x = time, y = value, col = cn0, linetype = cn0)) + 
#p2 <- ggplot( dsp2, aes(x = fracDecomp, y = value, col = cn0, linetype = cn0)) + 
  geom_line(size = 1) +
  facet_grid(~ scen) + #,scales = "free") + 
  #xlab("Time (yr)") + 
  #xlab("Substrate (proportion of initial)") + scale_x_reverse() +
  xlab("Time (yr)") +
  ylab("Carbon use efficiency") +
  #themeDefault +
  theme_classic(base_size = 18) + # presentation
  #scale_colour_discrete(drop = TRUE,limits = levels(dsp$Allocation)) +
  #theme(legend.position = c(0.95,1.04), legend.justification = c(1,1)) +
  theme(legend.position = c(0.95,0.05), legend.justification = c(1,0)) +
  theme(panel.border = element_rect(colour = "black", fill = NA)) +
  theme(strip.background = element_blank()) +			
  #theme(panel.grid.major.x = element_line(colour = "grey75")) +
  c()
#print(p2f + colScale)
print(p2 + expand_limits(y = 0.1))
```


```{r simPulse, fig.height=2, fig.width=3.27}
p2p <- p2 + 
  expand_limits(y = 0.1) +
  theme_classic(base_size = 10) +
  theme(legend.position = c(0.98,0.02), legend.justification = c(1,0)) +
  theme(panel.border = element_rect(colour = "black", fill = NA)) +
  theme(strip.background = element_blank()) +	
  theme(legend.title = element_blank()) +
  c()  
print(p2p)
```

