---
title: "Simulating decomposition of a litter bag"
output: html_notebook
---

## Summary

## Model versions
- SESAM4a: soil enzyme steady allocation model with microbial turnover rate by
predation increases linearly with microbial biomass.

```{r, echo=FALSE, results='hide', message=FALSE}
library(ggplot2)
library(grid)   #unit
library(reshape)  # melt
library(RColorBrewer) # brewer.pal
library(tidyr)
library(dplyr)
library(purrr)

myColors <- brewer.pal(5,"Dark2")	
.simpleCap <- function(s) {
    paste(toupper(substring(s, 1, 1)), substring(s, 2),  sep = "")
}
colScale <- scale_colour_manual(name = "Variant",values = myColors)
.treatments <- structure(c(1,0.5), names = c("Litter input pulse","No Litter input"))
sizeScale <- scale_size_manual(name = "Treatment",values = .treatments)

themeDefault <- theme_classic()  
```

### Setup parameters and drivers
Mass units are $gC/m^2$, $gN/m^2$ or $gP/m^2$. Time unit is $yr$.

```{r echo=TRUE, message=FALSE, warning=FALSE}
require(sesam)

parms0 <- list(
  cnB = 8.3       ##<< Cleveland07, grassland
  , cnBW = 8.4    ##<< C/N ratio of cell material that go to RC (cell walls) NRSOM Tipping 16
  , cnE = 3.2     # Sterner02: Protein (Fig. 2.2.), high N investment (low P)
  , cnIR = 4.5     ##<< between micr and enzyme signal
  , cnIL = 30      ##<< N litter OM
  , kmN = 0.3*0.01*365  ##<< /yr enzyme turnover 1% turning over each day
  , kNB = 0.8      ##<< amount of recycling enzyme turnover by biomass (added to uptake instead of RC)
  , kR = 1/(50)    ##<< 1/(x years)
  #, kL = 1/(1)     ##<< 1/(x years)
  , kL = 1/(5/365)     ##<< 1/(5 days) # for the priming study
  #, kL = 1/(2/12)  ##<< two months # for the litter bag decomposition
  , aE = 0.001*365 ##<< C biomass allocated to enzymes gC/day /microbial biomass
  #, km = 0.3       ##<< enzyme half-saturation constant
  , m = 0.02*365   ##<< maintenance respiration rate   gC/day /microbial biomass
  , tau = 1/60*365/5 ##<< biomass turnover rate (12 days), devided by 5 without predation
  , eps = 0.5      ##<< carbon use efficiency
  , epsPred = 0.3  ##<< carbon use efficiency of microbial tvr (predators respire)
  , iR = 0         ##<< input modelled explicitely
  , iL = 300       ##<< g/m2 input per year (half NPP)
  , plantNUp = 0   ##<< plant N uptake balancing N inputs
  , useFixedAlloc = FALSE    ##<< set to true to use fixed enzyme allocation (alpha = 0.5)
  , kIPlant = 10.57 #0.0289652*365         ##<< plant uptake iP I
  , iB = 0.38 * 10.57 #0.0110068*365   ##<< immobilization flux iB I
  , iI = 0         ##<< input of mineral N
  , l = 0.96 #0.00262647*365       ##<< leaching rate of mineralN l I
  , nu = 0.9       ##<< microbial N use efficiency
  , isEnzymeMassFlux = TRUE  ##<< steady state BC solution accounts for enyzme mass fluxes
  , nuP = 0.3      ##<< microbial uptake of depolymerized P, (1-nuP) is mineralized
  , cpE = 100     ##<< Sterner02, protein with no P
  , cpB = 47.3    ##<< Cleveland07, grassland
  , cpBW = 61     ##<< NRSOM Tipping16
  , cpIR = 61
  , cpIL = 919    ##<< NPSOM Tipping 16
  , iBP = 0.38 * 10.57 # start with same as N
  , cW = 0.8 # proportion of cell wall within microbial biomass
  , B0 = 0  # minimal biomass, below which predation rate is zero
  #, tauP = 0.2 # slope of predation rate with biomass
  , tauP = 0.5 # slope of predation rate with biomass
  , nuP = 0.01 # DOP almost all mineralized before taken up
)
parms0 <- within(parms0,{
  kIPlant <- plantNUpAbs <- 0			# no plant uptake
  lP <- l       # leaching rate of inorganic P equals that of N
  kIPPlant <- kIPlant  # plant uptake rate of P equals that of N
  iIP <- l      # assume no P inputs compensate for leaching
})
# assure no P-limitation
parms0 <- within(parms0,{
  lP <- kIPPlant <- 0   # no leading of P
  iBP <- 1e6            # no limit of P immobilization flux
})
# assure no limitaiton by abiotic N
parms0 <- within(parms0,{
  lI <- kIPlant <- 0    # no leading of P
  iB <- 1e6            # no limit of P immobilization flux
})
# neglect influence of inorganic pools (keep them at fixed initial values)
parms0$isFixedI <- TRUE
```

For simplicity model drivers are currently prescibed as constant parameters.
```{r}
drivers <- list(
  iR = 0           ##<< input of residue litter
  #, iL = 0         ##<< input of labile litter
  , iL = 5         ##<< input of labile litter assume some DOC to prevent biomass depletion
  , cnIR = 7       ##<< C/N ratio of the residue litter input
  , cnIL = 30      ##<< C/N ratio of labile litter input
  , cpIR = 40      ##<< C/P ratio of the residue litter input
  , cpIL = 919     ##<< C/P ratio of the labile litter input, NPSOM Tipping16
  , iI = 0         ##<< input of mineral N
  , iIP = 0        ##<< input of labile mineral P
)
parms0 <- c(parms0[setdiff(names(parms0),names(drivers))], drivers)
```

The initial pools must be specified. 
```{r}
x0 <- c( #aE = 0.001*365
  BC = 85                    ##<< microbial biomass in C units
  , RC = 1                   ##<< C residue OM
  , RN = 1/parms0$cnIR      ##<< N residue OM 
  , RP = 1/parms0$cpIR      ##<< P residue OM 
  , LC = 1000                ##<< C litter OM
  , LN = 1000/parms0$cnIL   ##<< N litter OM 
  , LP = 1000/parms0$cpIL   ##<< P litter OM 
  , I =  0.01               ##<< inorganic pool 
  , IP =  0.01              ##<< inorganic pool 
  , alpha = 0.0             ##<< microbial community partitioning [0,1]
)
```

# Run decomposition simulation
```{r}
times <- c(seq(0,2/12, length.out = 101),seq(2/12,2,length.out = 101)[-1],seq(2,20,length.out = 31)[-1])
res <- res1 <- as.data.frame(lsoda( x0, times, derivSesam4a, parms = parms0))
xE <-  unlist(tail(res,1)) # final state
x0S <- xE[1 + seq_along(x0)] 
```

Plot time course.
```{r echo=FALSE}
resC <- res %>% filter(time <= 1/12 * (1/(5/365))/parms0$kL )  %>% cueSesam %>% # constrained period
  mutate(timeDay = time*365)
ggplot( resC, aes(timeDay, LC)) + geom_line()
#ggplot( resC, aes(timeDay, RC)) + geom_line()
# microbial turnover
resAgg <- resC %>% select(timeDay, tvrB, tvrBPred) %>% gather(key = "variable", "value", -timeDay)
ggplot( resAgg, aes(timeDay, value, color = variable, linetype = variable)) + geom_line()
# organic-inorganic N fluxes
resAgg <- resC %>% select(timeDay, PhiTotal, PhiU, PhiB, PhiTvr) %>% gather(key = "variable", "value", -timeDay)
ggplot( resAgg, aes(timeDay, value, color = variable, linetype = variable)) + geom_line()
# N decomposition fluxes
# resAgg <- resC %>% select(timeDay, decNLR, decNE, decNB) %>% gather(key = "variable", "value", -timeDay)
# ggplot( resAgg, aes(timeDay, value, color = variable, linetype = variable)) + geom_line()
# synthesis and respiration
resAgg <- resC %>% select(timeDay, uptakeC, synB, respB, respO, respTvr) %>% gather(key = "variable", "value", -timeDay)
ggplot( resAgg, aes(timeDay, value, color = variable, linetype = variable)) + geom_line()

ggplot( resC, aes(timeDay, BC)) + geom_line()
```

## Inspecting C/N ratios and CUE
Microbial community changes take longer.
```{r echo=FALSE}
resC <- res1 %>% filter(time <= 1*(1/(5/365))/parms0$kL)  %>% cueSesam %>% # constrained period
  mutate(timeDay = time*365)
# CUE
resAgg <- resC %>% select(timeDay, cueDef, cueSyn) %>% gather(key = "variable", "value", -timeDay)
ggplot( resAgg, aes(timeDay, value, color = variable, linetype = variable)) + geom_line()
# elemental limitations
# resAgg <- resC %>% select(timeDay, limC, limN, limP) %>% gather(key = "variable", "value", -timeDay)
# ggplot( resAgg, aes(timeDay, value, color = variable, linetype = variable)) + geom_line()
# resAgg <- resC %>% select(timeDay, CsynBC, CsynBN) %>% gather(key = "variable", "value", -timeDay)
# ggplot( resAgg, aes(timeDay, value, color = variable, linetype = variable)) + geom_line()
resAgg <- resC %>% select(timeDay, alpha, alphaTarget, alphaC, alphaN) %>% 
  gather(key = "variable", "value", -timeDay)
ggplot( resAgg, aes(timeDay, value, color = variable, linetype = variable)) + geom_line()
```

course of C:N relative to initial values
```{r echo=FALSE}
 resAgg <- cueSesam(res1) %>% 
  select(time, cueDef, cueSyn) %>% gather(key = "variable", "value", -time) %>% 
  filter( time <= 5 * (1/(5/365))/parms0$kL)
 ggplot( resAgg, aes(time, value, color = variable, linetype = variable)) + geom_line()
# CN
resCM <- res1 %>% 
  mutate(betaL = LN/LC, betaR = RN/RC, OM = LC + RC + BC
         , OMN = LN + RN + BC/parms0$cnB, betaOM = OMN/OM
         , x = OM/OM[1], y = OMN/OMN[1]         )
ggplot( resCM, aes(x, y)) + geom_point()
#
resAgg <- resCM %>% filter(time <= 0.5*(1/(5/365))/parms0$kL)  %>% 
  mutate(timeDay = time*365) %>% 
  select(time, betaL, betaR, betaOM) %>% 
  gather(key = "variable", "value", -time)
ggplot( resAgg, aes(time, value, color = variable, linetype = variable)) + geom_line()
#
resAgg <- cueSesam(resCM) %>% 
  select(x, cueDef, cueSyn) %>% 
  gather(key = "variable", "value", -x)
ggplot( resAgg, aes(x, value, color = variable, linetype = variable)) + geom_line()
```
The initial decomposition state is determined by microbial growth and associated CUE.
For a declining biomass, CUE declines, on starvation to negative values.

For inferring CUE from litterbags, C0 should be only the first removal of a litter bag,
instead of the initial litter concentrations.

For the studied litter decomposition, the system stabilized after about 1 yr, but then
it consists almost solely of residue pool RC, and assumed trickling inputs into the 
fresh pool.

The initial missing N accumulation which is ascribed to lower CUE in Manzoni17,
is probably due to limitation of immobilization flux.
But what can we learn from it?

TODO: show a simulation with limited immobilization flux.


## Oscillations
There are curious oscillarions.
```{r echo=FALSE}
resC <- res1 %>% filter(time >= 2/12 * (1/(5/365))/parms0$kL & time <= 5 * (1/(5/365))/parms0$kL)  %>% cueSesam # constrained period
ggplot( resC, aes(time, BC)) + geom_line()
# CUE
resAgg <- resC %>% select(time, cueDef, cueSyn) %>% gather(key = "variable", "value", -time)
ggplot( resAgg, aes(time, value, color = variable, linetype = variable)) + geom_line()
# microbial community
resAgg <- resC %>% select(time, alpha, alphaTarget) %>% gather(key = "variable", "value", -time)
ggplot( resAgg, aes(time, value, color = variable, linetype = variable)) + geom_line()
```
They arise by microbial community changes, that are slow when
biomass is low and turnover is also low. They are stronger if labile pool
has a slower turnover rate.
Here it was set to a very small value for the priming study.

The LC is not depleted yet, but cannot deliver enough carbon to sustain biomass.

# Priming spike 
Simulate a spike of fresh litter on exhausted biomass.
```{r}
x02 <- x0S
LC <- 200
x02["BC"] <- 1e-8
x02["LC"] <- LC
x02["LN"] <- LC/parms0$cnIL
x02["LP"] <- LC/parms0$cpIL
  times1 <- c(seq(0,1/12, length.out = 51), seq(1/12,1, length.out = 51)[-1])
  res <- res2 <- as.data.frame(lsoda( x02, times1, derivSesam4a, parms = parms0))
  xE2 <-  unlist(tail(res,1)) # final state
  x0S <- xE[1 + seq_along(x0)] 
```

Plot time course.
```{r echo=FALSE}
resC1 <- res %>% mutate(timeDay = time*365) %>% 
  filter(timeDay <= 365)  %>%  # constrained period
  cueSesam 
resC <- resC1 %>% filter(timeDay <= 40)
ggplot( resC, aes(timeDay, LC)) + geom_line()
#ggplot( resC, aes(timeDay, RC)) + geom_line()
# microbial turnover
resAgg <- resC %>% select(timeDay, tvrB, tvrBPred) %>% gather(key = "variable", "value", -timeDay)
ggplot( resAgg, aes(timeDay, value, color = variable)) + geom_line()
# organic-inorganic N fluxes
resAgg <- resC %>% select(timeDay, PhiTotal, PhiU, PhiB, PhiTvr) %>% gather(key = "variable", "value", -timeDay)
ggplot( resAgg, aes(timeDay, value, color = variable, linetype = variable)) + geom_line()
# N decomposition fluxes
# resAgg <- resC %>% select(timeDay, decNLR, decNE, decNB) %>% gather(key = "variable", "value", -timeDay)
# ggplot( resAgg, aes(timeDay, value, color = variable, linetype = variable)) + geom_line()
# synthesis and respiration
resAgg <- resC %>% select(timeDay, uptakeC, synB, respB, respO, respTvr) %>% 
  gather(key = "variable", "value", -timeDay)
ggplot( resAgg, aes(timeDay, value, color = variable, linetype = variable)) + geom_line()

ggplot( resC, aes(timeDay, BC)) + geom_line()
```

```{r}
# CUE
resAgg <- resC1 %>% select(timeDay, cueDef, cueSyn) %>% gather(key = "variable", "value", -timeDay)
ggplot( resAgg, aes(timeDay, value, color = variable, linetype = variable)) + geom_line()
# elemental limitations
# resAgg <- resC %>% select(timeDay, limC, limN, limP) %>% gather(key = "variable", "value", -timeDay)
# ggplot( resAgg, aes(timeDay, value, color = variable, linetype = variable)) + geom_line()
# resAgg <- resC %>% select(timeDay, CsynBC, CsynBN) %>% gather(key = "variable", "value", -timeDay)
# ggplot( resAgg, aes(timeDay, value, color = variable, linetype = variable)) + geom_line()
resAgg <- resC1 %>% select(timeDay, alpha, alphaTarget, alphaC, alphaN) %>% 
  gather(key = "variable", "value", -timeDay)
ggplot( resAgg, aes(timeDay, value, color = variable, linetype = variable)) + geom_line()

```

```{r}

```

