---
title: "SESAM Soil enzyme steady allocoation model"
output:
  pdf_document:
    keep_tex: true
  html_notebook: default
---

Simplified from 19GMD_paper/CompareModels.Rmd

## Simulate a FACE scenario using the Sesam3b model
```{r, echo=FALSE, results='hide', message=FALSE}
library(ggplot2)
library(grid)   #unit
library(RColorBrewer) # brewer.pal
library(tidyr)
library(dplyr)

myColors <- brewer.pal(5,"Dark2")	
.simpleCap <- function(s) {
    paste(toupper(substring(s, 1, 1)), substring(s, 2),  sep = "")
}
colScale <- scale_colour_manual(name = "Variant",values = myColors)
.treatments <- structure(c(1,0.5), names = c("Litter input pulse","No Litter input"))
sizeScale <- scale_size_manual(name = "Treatment",values = .treatments)

themeDefault <- theme_classic()  
themePaper <- theme_classic(base_size = 10)  
```

### Setup parameters and drivers
Mass units are $gC/m^2$ or $gN/m^2$. Time unit is $yr$.
```{r}
require(sesam, quietly = TRUE)

parmsModel <- list(
  cnB = 11
  ,cnE = 3.1     # Sterner02: Protein (Fig. 2.2.), high N investment (low P) need 
    # to have low C/N for CO2 increase scenario
  ,kN = 60       ##<< /yr enzyme turnover 60 times a year, each 6 days -> fast 
    # priming pulses
  ,km = 0.05     ##<< enzyme half-saturation constant, in magnitude of enzymes, 
    # determined by kN
  ,kNB = 0.8     ##<< amount of recycling enzyme turnover by biomass (
    # added to assimilable, i.e. uptake instead of R)
  #,kR = 1/(10)        ##<< 1/(x years) # to demonstrate changes on short time scale
  ,kR = 1/(20)        ##<< 1/(x years) # to demonstrate changes on short time scale
  ,kL = 5           ##<< 1/(x years)   # formerly 1 year
  ,aE = 0.001*365   ##<< C biomass allocated to enzymes 1/day /microbial biomass
  ,m = 0.016*365    ##<< maintenance respiration rate   1/day /microbial biomass, 
  ## Bogedom Fig. 1
  ,tau = 1/60*365  ##<< biomass turnover rate (12 days)
  ,eps = 0.68      ##<< carbon use efficiency for growth respiration
  #,eps = 0.5      ##<< carbon use efficiency for growth respiration
  #,epsTvr = 0.3   ##<< carbon use efficiency of microbial tvr (part by predators 
  #,epsTvr = 0.45   ##<< carbon use efficiency of microbial tvr (part by predators 
  ,epsPred = 3     ##<< carbon use efficiency of microbial tvr (part by predators 
  ## which respire and corresponding amount of N must be mineralized)
  #,iBN = 0.38 * 10.57 #0.0110068*365   ##<< potential immobilization flux rate 
  ,iBN = 0.4 ##<< potential immobilization flux rate of inorganic N
  #,iBN = 1/12*(24*365) # tvr of IN of 12 hours
  ## (immoNPot = iBN IN)
  ,lN = 0.96       #0.00262647*365     ##<< leaching rate of mineral N: lN IN
  ,nuN = 0.9     # microbial N use efficiency accounting for apparent 
  ## minceralization of N during uptake in heterogeneous soils
  , isEnzymeMassFlux = TRUE  ##<< compute enzyme mass fluxes
  , kINPlant = 10.57 #0.0289652*365          ##<< plant uptake iP IN
  #
  #
  ,lP = 0.001  # very small to not run into P limitation  
  , cpE = 1e5
  , cpB = 1e8 # for not running into P limitation, here assume no P required in biomass
  , iBP = 0.38 * 10.57 # start with same as N
  , kIPPlant = 0.0 #kINPlant  # plant uptake rate of P equals that of N
  #
  # additional parameters with version 4: detailed microbial turnover
  #, cnBW = 10    ##<< C/N ratio of cell walls (that go to R, )
  # , cnBW = cnB    ##<< do not partition
  # , cpBW = 1e8   # not running into P limitation
  # #, cW = 0.5 # proportion of cell wall within microbial biomass
  # , cW = 1 # proportion of cell wall within microbial biomass
  # , B0 = 0  # minimal biomass, below which predation rate is zero
  # #, tauP = 0.1 # slope of predation rate with biomass
  # , tauP = 0 # slope of predation rate with biomass
  # , kSP = 1/30 # 1/x years
  # , pESP = 0.01  # /year production of phospatase by plants, small to check balance
  # , nuPP = 0.05 # most of the phosphats to to mineral pool
)

parmsModel <- within(parmsModel, {
    kmN <- km*kN     # Sesam2 only uses a lumped parameter instead of km and kN
    kmR <- kmL <- km  # seam2 can accept parameters differing by substrate
    cnER <- cnEL <- cnE 
    kNR <- kNL <- kN    
    #isFixedI <- TRUE  # in this demonstration keep inorganic pool fixed
})
```

For simplicity model drivers currently are prescribed as constant parameters.
```{r}
drivers <- list(
  iR = 0          ##<< input of residue litter
  ,iL = 400         # g/m2 input per year (half NPP)
  ,cnIR = 7       ##<< C/N ratio of the residue litter input
  ,cnIL = 30      ##<< C/N ratio of the labile litter input: 
  ,iIN = 0.0         ##<< input of mineral N
  ,iIP = 0.0 # no inorganic P inputs
  , cpIR = 40
  , cpIL = 40*3
)
parms0 <- within( c(parmsModel, drivers), {
  iIN <- iL/cnIL/200    # assume a fraction of 1/10 of annual litter production
  kINPlant <- 1/5*365 # tvr of 5 days
  plantNUpAbs <- iL/cnIL + iIN # plants take maximum out as much as returned by litter
  # plus inorganic input, difference is grazed
  # keep leaching as otherwise it accumulates during priming
  #lN <- 0 # no leaching
  kLP <- kRP <- kSP
  pELP <- pERP <- pESP
})
```

The initial pools must be specified. The Steady models do not have the enzyme pools as state variables.
```{r}
x0 <- x0Explicit <- c( #aE = 0.001*365
        BC = 17                     ##<< microbial biomass 
        ,RC = 1100                ##<< N rich substrate
        ,RN = 1100/parms0$cnIR   ##<< N rich substrate N pool
        ,RP = 1100/parms0$cpIR
        ,LC = 100                 ##<< N poor substrate
        ,LN = 100/parms0$cnIL    ##<< N poor substrate N pool
        ,LP = 100/parms0$cpIL    ##<< N poor substrate N pool
        ,IN =  0.04 #150 #0.04                ##<< inorganic pool gN/m2
        #,IP =  0.04 #150 #0.04                ##<< inorganic pool gN/m2
        ,IP =  150 # prevent P limitation          ##<< inorganic pool gN/m2
        #, alpha = 0.5            ##<< microbial community partitioning [0,1]
        , alphaR = 0.5, alphaL=0.4, alphaRP=0.1    ##<< microbial community partitioning [0,1]
)
parms = parms0
```

Next the simulation runs are performed.
```{r}
  t1S =20; t2I = 40; t3S = 70; parms = parms0
  tmp = derivSesam4b(0, x0, parms)

  # spinup to steady state
  times <- seq(0,1000, length.out = 500)
  #times <- seq(0,20, length.out = 50)
  res <- res1 <- as.data.frame(lsoda( x0, times, derivSesam4b, parms = parms))
  #ggplot(res, aes(time, IN)) + geom_line()
  #ggplot(res, aes(time, alphaR)) + geom_line()
  #ggplot(res, aes(time, BC)) + geom_line()
  #ggplot(res, aes(time, RC)) + geom_line()
  #ggplot(res, aes(time, limP)) + geom_line()
  xE <- unlist(tail(res,1))
  .interactive_trythis <- function(){
    # one call to the derivative function, 
    #    first element returns the derivative
    #    second element is a vector of additonal fluxes/outputs
    #      e.g. tmp[[2]]["uC"] is the carbon uptake of microbial biomass
    tmp = derivSesam4b(0, xE[1+(1:length(x0))], parms); tmp
    # task: compute CUE from these outputs
  }
  #
  # t1S years of steady state recorded
  # task: get familiar with lsoda from package deSolve
  res <- res1S <- as.data.frame(lsoda( 
    xE[1:length(x0) + 1], 1:t1S, derivSesam4b, parms = parms))
  #
  # t2I years of increased input
  fInputInc = 1.2
  parmsC2 <- within(parms, {# more C in litter, but not more N 
    iL <- iL*fInputInc
    cnIL <- cnIL*fInputInc
  }) 
  #derivSesam4b(0, xE, parmsC2)
  times <- seq(0,t2I, length.out = t2I*6 + 1)# record by-monthly
  res <- res2I <- as.data.frame(lsoda( 
    xE[1:length(x0) + 1], times, derivSesam4b, parms = parmsC2))
  res2I$time <- res2I$time + t1S 
  xE2 <- unlist(tail(res2I,1))
  #
  # t3S years of initial input
  times <- seq(0,t3S, length.out = t3S*6 + 1) # record bymonthly
  res <- res3S <- as.data.frame(lsoda( 
    xE2[1:length(x0) + 1], times, derivSesam4b, parms = parms))
  res3S$time <- res3S$time + t1S + t2I
  xE3 <- tail(res3S,1)
  resc <- rbind( res1S, res2I[-1,], res3S[-1,])
  head(resc)
```

Plot some results.

```{r}
vars <- c("L","R","alphaR","RL") 
vars <- c("L","R","alphaR","B","RL","IN") 
vars <- c("L","R","alphaR","B","RL","IN","plantNUp","PhiNB","PhiNTotal") 
#vars <- c("alpha","B", "cnL", "PhiNB","PhiNTotal","immoNPot", "CsynBC","CsynBN") 
dsp <- resc %>% 
  #filter(time >= 20, time <= 40) %>% 
  mutate(RL = R + L, cnRL = (R + L)/(RN + LN)) %>% # compute additional quantities
  #select_(~time, ~scen, ~L, ~R, ~RL) %>% gather_("Pool","value", c("L","R","RL"))
  select(any_of(c("time",vars))) %>% gather("Pool","value", vars)
p2f <- ggplot( dsp, aes(x = time, y = value)) + geom_line(linewidth = 1) +
  facet_grid(Pool ~ .,scales = "free") + 
  xlab("Time (yr)") + ylab(expression(Carbon~stock~(gCm^{-2}))) + #labs(linetype = "Substrate pool") +
  themeDefault +
  #scale_colour_discrete(drop = TRUE,limits = levels(dsp$Allocation)) +
  theme(legend.position = c(0.95,1.04), legend.justification = c(1,1)) +
  theme(panel.border = element_rect(colour = "black", fill = NA)) +
  theme(strip.background = element_blank()) +			
  theme(panel.grid.major.x = element_line(colour = "grey75")) +
  c()
#print(p2f + colScale)
print(p2f)
```
```{r}
# task: compute and plot CUE

```

Task: play with some of the parameters.
Look at simulated steady-state after spin-up and at limitation weights.


```{r}
.tmp.f <- function(){
  
}
```

