---
title: "Inspecting CUE effect on steady state of a simple one pool model"
output: html_notebook
---

## Summary
Markus drew a up a very simple SOM(S) and Microbial biomass (B) model,
computed Steady state and searched the CUE for which S is at minimu.


```{r, echo=FALSE, results='hide', message=FALSE}
library(ggplot2)
library(grid)   #unit
library(reshape)  # melt
library(RColorBrewer) # brewer.pal
library(tidyr)
library(dplyr)
library(purrr)

myColors <- brewer.pal(5,"Dark2")	
.simpleCap <- function(s) {
    paste(toupper(substring(s, 1, 1)), substring(s, 2),  sep = "")
}
colScale <- scale_colour_manual(name = "Variant",values = myColors)
.treatments <- structure(c(1,0.5), names = c("Litter input pulse","No Litter input"))
sizeScale <- scale_size_manual(name = "Treatment",values = .treatments)

themeDefault <- theme_classic()  
```

## The model
$$
\begin{align}
dS/dt &= I - f(S)g(B) + \tau B\\
dB/dt &= \epsilon f(S)g(B) - \tau B
\end{align}
$$

## Steady states
Steady state of S by setting derivatives to zero and adding the two equations:
$$
\begin{align}
0 &= I + (\epsilon - 1) f(S^*)g(B^*) \\
f(S^*) &= \frac{I}{(1-\epsilon) g(B^*)} \\
S^* &= f^{-1}\left( \frac{I}{(1-\epsilon) g(B^*)} \right) \\
\frac{dS}{dt} &= (f^{-1})' \left( \frac{I}{(1-\epsilon) g(B^*)} \right) \frac{I}{(1-\epsilon)^2 g(B^*)}\\
\end{align}
$$

With:
$$
\begin{align}
y &= f(S) = kS \\
f^{-1}(y) &= y/k \\
(f^{-1})'(y) &= 1/k \\
S^* &=  \frac{I}{k (1-\epsilon) g(B^*)}  \\
\end{align}
$$
Steady state stocks increase with decreasing steady state biomass if $g$ is 
monotonically increasing.
(This contradicts findings of Markus Lange at the Jena Experiment)

Steady state for Biomass B:
$$
\begin{align}
0 &= \epsilon \, k S^* g(B^*) - \tau B^* \\
0 &= \epsilon \, k \frac{I}{k (1-\epsilon) g(B^*)} g(B^*) - \tau B^* \\
0 &= \epsilon \, \frac{I}{(1-\epsilon) }  - \tau B^* \\
B^* &= \frac{\epsilon}{(1-\epsilon)} \frac{I}{\tau} \\
\end{align}
$$

$B^*$ does neither depend on $k$ (as time is not important for staedy state) nor
on $g(B)$.

With $g(B) = B/(k_M + B)$:
$$
\begin{align}
S^* &=  \frac{I}{k (1-\epsilon) g(B^*)}  \\
    &=  \frac{I}{k (1-\epsilon)}  \frac{k_M+B^*}{B^*}\\
    &=  \frac{I}{k (1-\epsilon)}  \frac{k_M+\frac{\epsilon}{(1-\epsilon)}
        \frac{I}{\tau}}{\frac{\epsilon}{(1-\epsilon)} \frac{I}{\tau}}\\
    &= \frac{k_M+ \frac{\epsilon}{(1-\epsilon)} \frac{I}{\tau}}{k \epsilon / \tau} \\
    &= \frac{\tau k_M+ \frac{\epsilon}{(1-\epsilon)} I}{k \epsilon} \\
    &= \frac{1}{k }\left( \frac{\tau k_M}{\epsilon }+ \frac{I}{(1-\epsilon)} \right)\\
\end{align}
$$
On increasing $\epsilon$, the first biomass related part decreases
but the second input related part increases. For $\epsilon \to 0$, $B^* \to 0$ and
steady state stocks increase because of no decomposition. 
For $\epsilon \to 1$, $B^* \to \infty$ but steady state stocks also
increase. Between there is a minimum obtained when

$$
\begin{align}
f(\epsilon) &= \frac{\tau k_M}{\epsilon }+ \frac{I}{(1-\epsilon)} \to min\\
f'(\epsilon) &= -\frac{\tau k_M}{\epsilon^2 }+ \frac{I}{(1-\epsilon)^2} \to min\\
0 &= (\epsilon -1)^2 \tau k_M +  \epsilon^2 I \\
0 &= (I + \tau k_M) \epsilon^2 -2 \tau k_M \epsilon + \tau k_M\\
\end{align}
$$

```{r include=FALSE}
.solveSquare <- function(
  ### provides the solutions of eq. 0 = a x^2 + b x + c
  a,b,c   ##<< coefficients of the square equation
){
  p2 <- b/a/2
  q <- c/a
  D <- sqrt(p2*p2 - q)
  x0 <- -p2 + c(+1,-1)*D
recover()  
  ### complex vector of length 2: roots
  x0
}
```


```{r echo=FALSE}
I <- 100
tau <- 12
k <- 1/5
B <- eps/(1 - eps) * I/tau
kM <- quantile(B, 0.8)
tkm <- tau*kM
#epsMin <- .solveSquare(I + tkm, -2*tkm, tkm )
S <- S1 <- I/(k*(1 - eps))*(kM + B)/B
#S <- S2 <- 1/k*(tau*kM/eps + I/(1 - eps))
ds <- data.frame(eps,B,S)
ggplot(ds, aes(eps,B)) + geom_line() + geom_point() +
  themeDefault
ggplot(ds, aes(eps,S)) + geom_line() + geom_point() +
  #geom_vline(aes(xintercept = epsMin), linetype = "dashed", color = "grey") +
  themeDefault
```

For CUE in a reasonable range, steady state stocks decrease with increasing CUE.
```{r}

```

