---
title: "CUE effect on SOM steady state with neglecting litter stock changes"
output: html_notebook
---

## Summary

Extending CUEInspectMarkus2.Rmd for microbial turnover increasing with
microbial biomass (density effect): $tvr_B = \tau B^\delta$.

```{r, echo=FALSE, results='hide', message=FALSE}
library(ggplot2)
library(grid)   #unit
library(reshape)  # melt
library(RColorBrewer) # brewer.pal
library(tidyr)
library(dplyr)
library(purrr)

myColors <- brewer.pal(5,"Dark2")	
.simpleCap <- function(s) {
    paste(toupper(substring(s, 1, 1)), substring(s, 2),  sep = "")
}
colScale <- scale_colour_manual(name = "Variant",values = myColors)
.treatments <- structure(c(1,0.5), names = c("Litter input pulse","No Litter input"))
sizeScale <- scale_size_manual(name = "Treatment",values = .treatments)

themeDefault <- theme_classic()  
```

## The model
$$
\begin{align}
dL/dt &= IN -f_L(L)g(B) \\
dB/dt &= \epsilon f_L(L)g(B) + \epsilon f_R(R)g(B) - \tau B^\delta \\
dR/dt &= -f_R(R)g(B) + \tau B^\delta\\
\end{align}
$$
where decomposition is a multiplicative function of stocks and microbial biomass,
and a part $(1-\epsilon)$ is respired.

And microbial turnover increases with microbial biomass, i.e. $\delta > 1$.

## Steady states
At steady state:
$$
\begin{align}
f_L(L)g(B) &= IN \\
dB/dt &= \epsilon IN + \epsilon f_R(R)g(B) - \tau B^\delta \\
\end{align}
$$

Steady state of $R$ by setting derivatives to zero and adding the equations of $dB$ and $dR$:
$$
\begin{align}
0 &= \epsilon IN + (\epsilon - 1) f_R(R^*)g(B^*) \\
f_R(R^*) &= \frac{\epsilon IN}{(1-\epsilon) g(B^*)} \\
R^* &= f^{-1}\left( \frac{\epsilon IN}{(1-\epsilon) g(B^*)} \right) \\
\frac{dR}{dt} &= (f^{-1})' \left( \frac{\epsilon IN}{(1-\epsilon) g(B^*)} \right) \frac{\epsilon IN}{(1-\epsilon)^2 g(B^*)}\\
\end{align}
$$

With:
$$
\begin{align}
y &= f_R(R) = k_R R \\
f^{-1}(y) &= y/k_R \\
(f^{-1})'(y) &= 1/k_R \\
R^* &=  \frac{\epsilon IN}{k_R (1-\epsilon) g(B^*)}  \\
\end{align}
$$
Steady state for Biomass B:
$$
\begin{align}
0 &= \epsilon IN +\epsilon \, k_R R^* g(B^*) - \tau {B^*}^\delta \\
0 &= \epsilon IN + \epsilon \, k_R \frac{\epsilon IN}{k_R (1-\epsilon) g(B^*)} g(B^*) - \tau {B^*}^\delta \\
0 &= \epsilon IN + \epsilon \, \frac{\epsilon IN}{(1-\epsilon) }  - \tau {B^*}^\delta \\
B^* &= \left( \left( 1 + \frac{\epsilon}{(1-\epsilon)} \right) \frac{\epsilon IN}{\tau} \right)^{1/\delta}
     = \left( \frac{\epsilon IN}{\tau (1-\epsilon)} \right)^{1/\delta} \\
\end{align}
$$

$B^*$ does neither depend on dynamics of the litter pool nor on $k_R$ (as time is not important for staedy state) nor
on $g(B)$.

With $g(B) = B/(k_M + B)$:
$$
\begin{align}
R^* &=  \frac{\epsilon IN}{k_R (1-\epsilon) g(B^*)}  
    =  \frac{\epsilon IN}{k_R (1-\epsilon)}  \frac{k_M+B^*}{B^*}
    =  \frac{\epsilon IN}{k_R (1-\epsilon)}  \left( k_M {B^*}^{-1} + 1\right) \\
    &= \frac{\epsilon IN}{k_R (1-\epsilon)}  \left( k_M \left( \frac{\epsilon IN}{\tau (1-\epsilon)} \right)^{-1/\delta} + 1\right) \\
    &= \frac{1}{k_R}  \left( k_M \tau^{1/\delta} \left( \frac{\epsilon IN}{1-\epsilon} \right)^{1-1/\delta} + \frac{\epsilon IN}{1-\epsilon}\right) \\
\end{align}
$$

And with $f_L(L) = k_L L$:
$$
\begin{align}
L^* &= f_L^{-1}\left( IN / g(B^*) \right)    
    = \frac{IN}{k_L}  \frac{k_M+B^*}{B^*}
    = \frac{IN}{k_L}  ( k_M {B^*}^{-1} +1 ) \\
    &= \frac{IN}{k_L}  \left( k_M \left( \frac{\epsilon IN}{\tau (1-\epsilon)} \right)^{-1/\delta} +1 \right) \\
    &= \frac{IN}{k_L}  \left( k_M \left( \frac{\tau (1-\epsilon)}{\epsilon IN} \right)^{1/\delta} +1 \right) \\
\end{align}
$$
With $1 \leq \delta \leq 2$ and $x > 1$, the term $x^{1-1/\delta}$ is usually smaller than $x^1$. 
Hence the $R^*$ dynamics at high litter inputs is similar to the model without density 
dependent turnover.

On increasing $\epsilon$ the second input related term of $R^*$ increases. 
For $\epsilon \to 0$, $B^* \to 0$ and
residue steady state stocks decrease towards a minimumm, but the 
litter stock will accumulate unlimited.
For $\epsilon \to 1$, $B^* \to \infty$ but steady state stocks also
increase unlimited, as there is no loss of carbon from the system.

Hence with sufficient litter input residue stocks always decrease with decreasing CUE.
The balance of total SOM stocks will depend on the balance of increasing 
litter stock and decreasing residue stock. However, residue stocks and their
changes are usually larger than changes in litter stocks. Moreover the balance
between litter and residue is  also be strongly controlled by their stoichiometry, 
e.g. their nitrogen contents.

## Figures and numerical solution

```{r include=FALSE}
eps <- seq(0.1, 0.8, by = 0.05)
IN <- 100
tau <- 12
kR <- 1/5
#delta <- 2; kL <- kR
#delta <- 1.5; kL <- kR*5
delta <- 1.2; kL <- 5
B <- (eps*IN/(tau*(1 - eps)))^(1/delta)
kM <- as.numeric(quantile(B, 0.8))
tkm <- tau*kM
R <- R1 <- eps*IN/(kR*(1 - eps))*(kM + B)/B
R <- R2 <- 1/kR*(kM*tau^(1/delta)*(eps*IN/(1 - eps))^(1 - 1/delta) + eps*IN/(1 - eps))
L <- L1 <- IN/kL*(kM + B)/B
L <- L2 <- IN/kL*(kM*(tau*(1 - eps)/(eps*IN))^(1/delta) + 1)
ds <- data.frame(eps,L,B,R)
ggplot(ds, aes(eps,B)) + geom_line() + geom_point() +
  themeDefault
ggplot(ds, aes(eps,R)) + geom_line() + geom_point() +
  themeDefault
```


The following plots show the relatinship between residue steady states and CUE.
The red square denotes the state of a simulation of the model over 100 years.
The gray dashed line denotes the CUE for which SOM stocks are minimal.

```{r echo=FALSE}
p <- list(IN = IN, tau = tau, kR = kR, kL = kL, kM = kM, delta = delta, eps = 0.4)
unlist(p)
```

```{r include=FALSE}
require(deSolve)
dMod2 <- function(t,x,p){
  gB <- x["B"]/(p$kM + x["B"])
  decR <- p$kR*x["R"] * gB
  decL <- p$kL*x["L"] * gB
  tvrB <- p$tau*x["B"]^p$delta
  dL <- p$IN - decL
  dB <- p$eps*decL + p$eps*decR - tvrB
  dR <- -decR + tvrB
  list(setNames( c(dL, dB, dR), c("L","B","R")))
}
x0 <- c( L = 500, B = 14, R = 1800)
ans0 <- dMod2(0, x0, p)
times <- seq(0,500, length.out = 101)
res <- as.data.frame(lsoda(x0, times, dMod2, p))
xE <- unlist(tail(res,1)[-1])
ggplot(res, aes(time,R)) + geom_line()
```

```{r echo=FALSE}
dsSim <- cbind(eps = p$eps, as.data.frame(t(xE)))
ggplot(ds, aes(eps,B)) + geom_line() + geom_point() +
  geom_point(data = dsSim, col = "red", shape = 22) +
  themeDefault
ggplot(ds, aes(eps,R)) + geom_line() + geom_point() +
  geom_point(data = dsSim, col = "red", shape = 22) +
  themeDefault
ggplot(ds, aes(eps,L)) + geom_line() + geom_point() +
  geom_point(data = dsSim, col = "red", shape = 22) +
  themeDefault
ggplot(ds, aes(eps,L + B + R)) + geom_line() + geom_point() +
  geom_point(data = dsSim, col = "red", shape = 22) +
  themeDefault
```

The density-dependent turnover diminishes the effect of litter accumulation
with low CUE.


```{r}

```

