---
title: "SESAM predictions of emergent CUE"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document:
    keep_tex: true
---

Based on Sesam3P version, with parameterization that avoids P limitation.
Plant uptake of inorganic N varies to match N input by litter, to balance the 
system.

## Emergent CUE of SESAM predictions
```{r, echo=FALSE, results='hide', message=FALSE}
library(ggplot2)
library(grid)   #unit
library(RColorBrewer) # brewer.pal
library(tidyr)
library(dplyr)
library(purrr)
library(forcats)
library(knitr)

mycolors <- brewer.pal(6,"Dark2")
mylinetypes <- c("solid","dashed","dotdash","twodash","dotted","longdash")
scale_colour_discrete <- function(...) scale_colour_manual(...,values=mycolors)

adjust_facets <- function(theme){
  theme +
    theme(panel.border = element_rect(colour = "black", fill = NA)) +
    theme(strip.background = element_blank()) 
}
themePaper <- (theme_classic(base_size = 10) + theme(axis.title = element_text(size = 9))) %>% 
  adjust_facets()
themePres <- theme_classic(base_size = 18) %>% 
  adjust_facets()

#figWidthPage <- 18/2.54; figHeightPage <- figWidthPage/1.6 
# GMD fixes at 12 instead of entire page 18
figWidthPage <- 12/2.54; figHeightPage <- figWidthPage/1.6 
#, fig.height=4.43, fig.width=7.1
#figWidthWide <- 12/2.54; figHeightWide <- figWidthWide/1.6
#figWidthWide <- 3.8; figHeightWide <- 5
#, fig.height=2.95, fig.width=4.72
figWidthWide <- 7; figHeightWide <- figWidthWide/1.6 # ppt wide 2 figures on one page
figWidthCol <- 8.3/2.54; figHeightCol <- figWidthCol/1.6
#, fig.height=2.04, fig.width=3.27
#
theme_set(themePaper); figWidth <- figWidthCol; figHeight <- figHeightCol
theme_set(themePres); figWidth <- figWidthWide; figHeight <- figHeightWide 
#theme_set(themePres); figWidth <- figWidthSAB; figHeight <- figHeightSAB 


opts_chunk$set(
   # out.extra='style="display:block; margin: auto"'
    fig.width = figWidth, fig.height = figHeight #goldener Schnitt 1.6 #GI single column
    #, dev.args = list(pointsize = 11)
    , dev = c('pdf','png')
    , echo = FALSE
    )
knit_hooks$set(spar  =  function(before, options, envir) {
    if (before) {
        par( las = 1 )                   #also y axis labels horizontal
        par(mar = c(2.0,3.3,0,0) + 0.3 )  #margins
        par(tck = 0.02 )                          #axe-tick length inside plots             
        par(mgp = c(1.1,0.2,0) )  #positioning of axis title, axis labels, axis
    }
})

.simpleCap <- function(s) {
    paste(toupper(substring(s, 1, 1)), substring(s, 2),  sep = "")
}
.approaches <- c("Optimal","Derivative","Relative","Fixed")
.approaches_ext <- c(.approaches, c("Relative_kR","Relative_kmNR"))
colScaleApproach <- scale_colour_manual(
  name = "Approach", values = c(
    setNames(mycolors[seq_along(.approaches)], .approaches)
  ,c(Relative_kR = mycolors[3], Relative_kmNR=mycolors[3])))
linetypeScaleApproach <- scale_linetype_manual(
  name = "Approach",values = setNames(mylinetypes[seq_along(.approaches_ext)], .approaches_ext))
.treatments <- structure(c(1,0.5), names = c("Litter input pulse","No Litter input"))
sizeScale <- scale_size_manual(name = "Treatment",values = .treatments)
```

```{r}
library(devEMF)
figWidthSAB <- 8.1/2.54; figHeightSAB <- figWidthCol/1.6
save_emf <- function(filename, p=last_plot(), width = figWidthSAB, height = figWidthCol/1.6) {
  fname = paste0(basename(filename),".emf")
  emf(file = fname, emfPlus = FALSE, width = width, height = height)
  print(p)
  dev.off()  
  message("saved plot to ",fname)
}
```

### Setup parameters and drivers
Mass units are $gC/m^2$ or $gN/m^2$. Time unit is $yr$.
```{r}
require(sesam, quietly = TRUE)

parmsModel <- list(
  cnB = 11
  ,cnE = 3.1     # Sterner02: Protein (Fig. 2.2.), high N investment (low P) need 
    # to have low C/N for CO2 increase scenario
  ,kN = 60       ##<< /yr enzyme turnover 60 times a year, each 6 days -> fast 
    # priming pulses
  ,km = 0.05     ##<< enzyme half-saturation constant, in magnitude of enzymes, 
    # determined by kN
  ,kNB = 0.8     ##<< amount of recycling enzyme turnover by biomass (
    # added to assimilable, i.e. uptake instead of R)
  #,kR = 1/(10)        ##<< 1/(x years) # to demonstrate changes on short time scale
  ,kR = 1/31.40989 # matching initial R pools of about 3200 g/m2
  #,kR = 1/(20)        ##<< 1/(x years) # to demonstrate changes on short time scale
  ,kL = 5           ##<< 1/(x years)   # formerly 1 year
  ,aE = 0.001*365   ##<< C biomass allocated to enzymes 1/day /microbial biomass
  ,m = 0.016*365    ##<< maintenance respiration rate   1/day /microbial biomass, 
  ## Bogedom Fig. 1
  ,tau = 1/60*365  ##<< biomass turnover rate (12 days)
  ,eps = 0.68      ##<< carbon use efficiency for growth respiration
  #,eps = 0.5      ##<< carbon use efficiency for growth respiration
  ,epsTvr = 0.3   ##<< carbon use efficiency of microbial tvr (part by predators 
  #,epsTvr = 0.45   ##<< carbon use efficiency of microbial tvr (part by predators 
  #,epsPred = 3     ##<< carbon use efficiency of microbial tvr (part by predators 
  ## which respire and corresponding amount of N must be mineralized)
  #,iBN = 0.38 * 10.57 #0.0110068*365   ##<< potential immobilization flux rate 
  ,iBN = 0.4 ##<< potential immobilization flux rate of inorganic N
  #,iBN = 1/12*(24*365) # tvr of IN of 12 hours
  ## (immoNPot = iBN IN)
  ,lN = 0.96       #0.00262647*365     ##<< leaching rate of mineral N: lN IN
  ,nuN = 0.9     # microbial N use efficiency accounting for apparent 
  ## minceralization of N during uptake in heterogeneous soils
  , isEnzymeMassFlux = TRUE  ##<< compute enzyme mass fluxes
  , kINPlant = 10.57 #0.0289652*365          ##<< plant uptake iP IN
  #
  #
  ,lP = 0.001  # very small to not run into P limitation  
  ,nuP = 0.0     # microbial P use efficiency: no direct P uptake
  , cpE = 50
  , cpB = 40 # for not running into P limitation, here assume no P required in biomass
  , iBP = 100 # arbitrarily high to not run into P limitation
  , kIPPlant = 0.0 #kINPlant  # plant uptake rate of P equals that of N
  , e_P = 0.0 # no plant P enzyme production
  #
  # additional parameters with version 4: detailed microbial turnover
  #, cnBW = 10    ##<< C/N ratio of cell walls (that go to R, )
  # , cnBW = cnB    ##<< do not partition
  # , cpBW = 1e8   # not running into P limitation
  # #, cW = 0.5 # proportion of cell wall within microbial biomass
  # , cW = 1 # proportion of cell wall within microbial biomass
  # , B0 = 0  # minimal biomass, below which predation rate is zero
  # #, tauP = 0.1 # slope of predation rate with biomass
  # , tauP = 0 # slope of predation rate with biomass
  #, kSP = 1/30 # 1/x years
  # , pESP = 0.01  # /year production of phospatase by plants, small to check balance
  # , nuPP = 0.05 # most of the phosphats to to mineral pool
)

parmsModel <- within(parmsModel, {
    kmN <- km*kN     # Sesam2 only uses a lumped parameter instead of km and kN
    kmR <- kmL <- km  # seam2 can accept parameters differing by substrate
    cnER <- cnEL <- cnE 
    kNR <- kNL <- kN    
})
```

```{r}
# For simplicity model drivers currently are prescribed as constant parameters.
drivers <- list(
  iR = 0           ##<< input of residue litter
  ,iL = 400        # g/m2 input per year (half NPP)
  ,cnIR = 7        ##<< C/N ratio of the residue litter input
#  ,cnIL = 30      ##<< C/N ratio of the labile litter input: 
  ,cnIL = 28       ##<< C/N ratio of the labile litter input: 
  ,iIN = 0.0       ##<< input of mineral N
  ,iIP = 0.0       # no inorganic P inputs
  , cpIR = 40
  , cpIL = 40*3
)
parms0 <- parms <- within( c(parmsModel, drivers), {
  iIN <- 0.0714 #for cnIL = 28 # iL/cnIL/200    # assume a fraction of 1/10 of annual litter production
  kINPlant <- 1/5*365 # tvr of 5 days
  plantNUpAbs <- iL/cnIL + iIN # plants take maximum out as much as returned by litter
  # plus inorganic input, difference is grazed
  # keep leaching as otherwise it accumulates during priming
  #lN <- 0 # no leaching
  kLP <- kL # same efficiency as depolymerizing enzyme 
  kRP <- kR
  cpm <- 500
})
parms0F <- within(parms0,isFixedAlpha<-TRUE)
```

```{r}
x0 <- x0Explicit <- c( #aE = 0.001*365
        B = 17                     ##<< microbial biomass 
        ,R = 1100                ##<< N rich substrate
        ,RN = 3000/parms0$cnIR   ##<< N rich substrate N pool
        ,RP = 3000/parms0$cpIR
        ,L = 100                 ##<< N poor substrate
        ,LN = 100/parms0$cnIL    ##<< N poor substrate N pool
        ,LP = 100/parms0$cpIL    ##<< N poor substrate N pool
        ,IN =  0.04 #150 #0.04                ##<< inorganic pool gN/m2
        #,IP =  0.04 #150 #0.04                ##<< inorganic pool gN/m2
        ,IP =  150 # prevent P limitation          ##<< inorganic pool gN/m2
        #, alpha = 0.5            ##<< microbial community partitioning [0,1]
        #, alphaL=0.4, alphaR = 0.5, alphaRP=0.1    ##<< microbial community partitioning [0,1]
        , alphaL=0.7, alphaR = 0.3    ##<< microbial community partitioning [0,1]
)
times <- seq(0,10000, length.out = 500)
```

```{r}
computeCNSOM <- function(xEs, parms=parms0){
  xEs$SOC <- xEs$L + xEs$R + xEs$B
  xEs$SON <- xEs$LN + xEs$RN + xEs$B/parms$cnB
  xEs$CN <- xEs$SOC / xEs$SON
  xEs
}
```

### Gradient of C/N of litter inputs

Simulate steady state stocks with a gradient of C/N of litter inputs
yields the expected pattern of a decreasing CUE when C/N is higher than 
a threshold elemental ratio.

```{r}
cnILs <- seq(20,40,by = 2)
cnILi <- cnILs[1]
#cnILi <- cnILs[length(cnILs)]
xEs <- as.data.frame(t(sapply(cnILs, function(cnILi){
  parms_i <- within(parms0, {
    cnIL <- cnILi
    plantNUpAbs <- iL/cnIL + iIN # plants take maximum out as much 
    })
  res <- resA <- as.data.frame(lsoda( x0, times, derivSesam3P, parms = parms_i))
  xE <- cueSesam(tail(res,1))
  xE$cnIL <- cnILi
  unlist(xE)
})))
xEs$variant = "Adaptive"
xEs <- computeCNSOM(xEs)
xEFs <- as.data.frame(t(sapply(cnILs, function(cnILi){
  parms_i <- within(parms0F, {
    cnIL <- cnILi
    plantNUpAbs <- iL/cnIL + iIN # plants take maximum out as much 
    })
  res <- resF <- as.data.frame(lsoda( x0, times, derivSesam3P, parms = parms_i))
  xE <- cueSesam(tail(res,1))
  #derivSesam3P(0,unlist(xE[2:12]), parms_i)
  xE$cnIL <- cnILi
  unlist(xE)
})))
xEFs$variant = "Fixed"
xEFs <- computeCNSOM(xEFs)
ggplot(rbind(xEs,xEFs), aes(cnIL, cueSyn, color=variant, shape=variant)) + geom_point() + 
  scale_y_continuous('CUE') +
  scale_x_continuous('C/N litter input') +
  theme()
```

```{r}
ggplot(xEs, aes(cnIL, cueSyn)) + geom_point() + 
  #themePres +
  scale_y_continuous('CUE') +
  scale_x_continuous('C/N litter input') +
  theme()
save_emf("simCN_CNLitter_CUE")
```

There is no difference between adaptive and fixed community enzyme allocation.
At steady state, the decomposition matches the inputs to the system. 

```{r}
ggplot(rbind(xEs,xEFs), aes(cnIL, SOC, color=variant, shape=variant)) + geom_point() + 
  #themePres +
  scale_y_continuous('SOC (g/m2)') +
  scale_x_continuous('C/N litter input') +
  theme()
```
With non-adaptive fixed allocation, the stocks decrease more strongly with
N limitation.
With C limitation (low C/N), the adaptive variant results in higher C stocks, 
whereas with N limitation, the adaptive variant results in lower C stocks.

```{r}
ggplot(xEs, aes(cnIL, SOC)) + geom_point() + 
  scale_x_continuous('C/N litter input') +
  scale_y_continuous('SOC (g/m2)') +
  theme()
save_emf("simCN_CNLitter_stocks")
```



```{r eval=FALSE, include=FALSE}
ggplot(rbind(xEs,xEFs), aes(cnIL, SON, color=variant, shape=variant)) + geom_point() + 
  #themePres +
  scale_y_continuous('SON (g/m2)') +
  scale_x_continuous('C/N litter input') +
  theme()
```

```{r}
ggplot(rbind(xEs,xEFs), aes(cnIL, alphaR, color=variant, shape=variant )) + geom_point() + 
  scale_y_continuous('allocation to R degrading enzymes') +
  scale_x_continuous('C/N litter input') +
  theme()
```

However, this CUE pattern with litter input C/N leads to only slight variation in
total SOM C/N.

```{r}
ggplot(rbind(xEs,xEFs), aes(CN, cueSyn, color=variant, shape=variant)) + geom_point() + 
  theme(axis.text.x = element_text(angle = 30, vjust = 0.5)) +
  scale_x_continuous('C/N SOM stocks') +
  scale_y_continuous('CUE') +
  theme()
```
```{r}
ggplot(xEs, aes(CN, cueSyn)) + geom_point() + 
  theme(axis.text.x = element_text(angle = 30, vjust = 0.5)) +
  scale_x_continuous('C/N SOM stocks') +
  scale_y_continuous('CUE') +
  theme()
save_emf("simCN_CNStocks_CUE")
```


SESAM predicts a nonlinear increase of long-term stocks with CUE.
This corresponds to an initially steep decrease of stocks with increasing 
N limitation.

```{r}
ggplot(rbind(xEs,xEFs), aes(cueSyn, SOC, color=variant, shape=variant)) + geom_point() + 
  scale_y_continuous('SOC') +
  scale_x_continuous('CUE') +
  theme()
```

```{r}
ggplot(xEs, aes(cueSyn, SOC)) + geom_point() + 
  scale_x_continuous('CUE') +
  scale_y_continuous('SOC') +
  theme()
save_emf("simCN_CUE_stocks")

```

### Gradient of residue turnover rate

A stronger stabilization by interaction with minerals can roughly be simulated
by a decreasing decomposition rate of the residue pool.
Here, we vary the rate to decrease from 1/20yr to 1/200 yrs.

```{r}
tauRs <- exp(seq(log(20),log(500),length.out = 10))
tauRi <- tauRs[length(tauRs)]
cnILi <- parms0$cnIL #27
xER <- as.data.frame(t(sapply(tauRs, function(tauRi){
  parms_i <- within(parms0, {
    kR <- kRP <- 1/tauRi
    cnIL <- cnILi
  })
  res <- as.data.frame(lsoda( x0, times, derivSesam3P, parms = parms_i))
  xE <- cueSesam(tail(res,1))
  xE$tauR <- tauRi
  unlist(xE)
})))
xER <- computeCNSOM(xER)
ggplot(xER, aes(tauR, cueSyn)) + geom_point() + 
  #themePres +
  scale_x_continuous('Residue turnover time (yr)') +
  scale_y_continuous('CUE') +
  theme()
```

CUE, in SESAM does not vary with residue turnover time. The slight variation
at very large turnover times result from approaching higher steady-state
stocks very slowly.
Again its the steady state stocks that change with parameters, but result in
the same fluxes that at steady state are determined by the inputs.

### Temperature

```{r}
compute_temp_fac = function(temp, q10=2, T0=20) {
  q10^((temp-T0)/10)
}
temp = seq(5,30, by=5)
mT = compute_temp_fac(temp)
ggplot(data.frame(temp,mT), aes(temp,mT)) + geom_point()
```
```{r}
cnILi <- parms0$cnIL #27
xET <- as.data.frame(t(sapply(temp, function(tempi){
  mTi <- compute_temp_fac(tempi)
  parms_i <- within(parms0, {
    kR <- kRP <- parms0$kR * mTi
    kL <- kLP <- parms0$kL * mTi
    cnIL <- cnILi
  })
  res <- as.data.frame(lsoda( x0, times, derivSesam3P, parms = parms_i))
  xE <- cueSesam(tail(res,1))
  xE$T <- tempi
  xE$mT <- mTi
  unlist(xE)
})))
xET <- computeCNSOM(xET)
ggplot(xET, aes(temp, cueSyn)) + geom_point() + 
  #themePres +
  scale_x_continuous('Temperature (°C)') +
  scale_y_continuous('CUE') +
  theme()

```
```{r}
ggplot(xET, aes(temp, SOC)) + geom_point() + 
  #themePres +
  scale_x_continuous('Temperature (°C)') +
  scale_y_continuous('SOC') +
  theme()
```

Although SOC stocks decrease with increasing temperature, CUE predicted by
SESAM does not change.

### Amounts of litter input

```{r}
#cnILi <- parms0$cnIL 
cnILi <- 32 # assume high C/N ratio as typical
iLs <- seq(50,600, by=50)
xEL <- as.data.frame(t(sapply(iLs, function(iLi){
  parms_i <- within(parms0, {
    iL <- iLi
    plantNUpAbs <- iL/cnIL + iIN # plants take maximum out as much 
    cnIL <- cnILi
  })
  res <- as.data.frame(lsoda( x0, times, derivSesam3P, parms = parms_i))
  xE <- cueSesam(tail(res,1))
  xE$iL <- iLi
  unlist(xE)
})))
xEL <- computeCNSOM(xEL)
xEL$variant = "Adaptive"
xELF <- as.data.frame(t(sapply(iLs, function(iLi){
  parms_i <- within(parms0F, {
    iL <- iLi
    plantNUpAbs <- iL/cnIL + iIN # plants take maximum out as much 
    cnIL <- cnILi
  })
  res <- as.data.frame(lsoda( x0, times, derivSesam3P, parms = parms_i))
  xE <- cueSesam(tail(res,1))
  xE$iL <- iLi
  unlist(xE)
})))
xELF <- computeCNSOM(xELF)
xELF$variant = "Fixed"
ggplot(rbind(xEL, xELF), aes(iL, cueSyn, color=variant, shape=variant)) + geom_point() + 
  #themePres +
  expand_limits(y = c(0.2, 0.3)) +
  scale_x_continuous('Litter input (gC/m^2)') +
  scale_y_continuous('CUE') +
  theme()
```


```{r}
select(rbind(xEL, xELF), variant, iL, limC, limN, SOC, alphaR) %>% 
  arrange(iL, desc(variant))
```

Even with different amount of litter input, there is no change in predicted CUE 
between the adaptive and fixed model variant.
Its only the stocks that change.

```{r}
ggplot(rbind(xEL, xELF), aes(iL, SOC, color=variant, shape=variant)) + geom_point() + 
  #themePres +
  scale_x_continuous('Litter input (gC/m^2)') +
  scale_y_continuous('SOC (g/m2)') +
  theme()
```

```{r}
ggplot(rbind(xEL, xELF), aes(iL, alphaR, color=variant, shape=variant)) + geom_point() + 
  #themePres +
  scale_x_continuous('Litter input (gC/m^2)') +
  scale_y_continuous('alphaR') +
  theme()
```

With lower litter input, the differences between the variants 
(increase at lower C/N of inputs, but)
decrease at higher C/N of inputs.
Hence, we expect differences between fixed and adaptive variant 
to decrease in subsoil, because both C/N of litter inputs and amount
of litter inputs decrease.

### Inspecting single runs
```{r eval=FALSE, include=FALSE}
times <- seq(0,1000, length.out = 500)
times <- seq(0,10000, length.out = 500)
ggplot(res, aes(time, B)) + geom_point()
ggplot(res, aes(time, limC)) + geom_point()
ggplot(res, aes(time, R)) + geom_point()
ggplot(cueSesam(res), aes(time, cueSyn)) + geom_point()


```

