---
title: "Inspecting C:N effects on CUE and steady states"
output: html_notebook
---

## Summary
What is the effect of changing C:N of litter input on carbon use efficiency CUE 
and on predicted steady states?

Here, we vary the C:N ratio of inputs and plot the resulting steady state of stocks
and other variables.

## Model versions
- SESAM4a: soil enzyme steady allocation model with microbial turnover rate by
predation increases linearly with microbial biomass.

```{r, echo=FALSE, results='hide', message=FALSE}
library(ggplot2)
library(grid)   #unit
library(reshape)  # melt
library(RColorBrewer) # brewer.pal
library(tidyr)
library(dplyr)
library(purrr)

myColors <- brewer.pal(5,"Dark2")	
.simpleCap <- function(s) {
    paste(toupper(substring(s, 1, 1)), substring(s, 2),  sep = "")
}
colScale <- scale_colour_manual(name = "Variant",values = myColors)
.treatments <- structure(c(1,0.5), names = c("Litter input pulse","No Litter input"))
sizeScale <- scale_size_manual(name = "Treatment",values = .treatments)

themeDefault <- theme_classic()  
```

### Setup parameters and drivers
Mass units are $gC/m^2$, $gN/m^2$ or $gP/m^2$. Time unit is $yr$.

```{r echo=TRUE, message=FALSE, warning=FALSE}
require(sesam)

parms0 <- list(
  cnB = 8.3       ##<< Cleveland07, grassland
  , cnBW = 8.4    ##<< C/N ratio of cell material that go to R (cell walls) NRSOM Tipping 16
  , cnE = 3.2     # Sterner02: Protein (Fig. 2.2.), high N investment (low P)
  , cnIR = 4.5     ##<< between micr and enzyme signal
  , cnIL = 30      ##<< N litter OM
  , kN = 0.01*365  ##<< /yr enzyme turnover 1% turning over each day
  , kNB = 0.8      ##<< amount of recycling enzyme turnover by biomass (added to uptake instead of R)
  , kR = 1/(50)    ##<< 1/(x years)
  , kL = 1/(1)     ##<< 1/(x years)
  , aE = 0.001*365 ##<< C biomass allocated to enzymes gC/day /microbial biomass
  , km = 0.3       ##<< enzyme half-saturation constant
  , m = 0.02*365   ##<< maintenance respiration rate   gC/day /microbial biomass
  , tau = 1/60*365 ##<< biomass turnover rate (12 days)
  , eps = 0.5      ##<< carbon use efficiency
  , epsPred = 0.3  ##<< carbon use efficiency of microbial tvr (predators respire)
  , iR = 0         ##<< input modelled explicitely
  , iL = 300       ##<< g/m2 input per year (half NPP)
  , plantNUp = 0   ##<< plant N uptake balancing N inputs
  , useFixedAlloc = FALSE    ##<< set to true to use fixed enzyme allocation (alpha = 0.5)
  , kIPlant = 10.57 #0.0289652*365         ##<< plant uptake iP I
  , iB = 0.38 * 10.57 #0.0110068*365   ##<< immobilization flux iB I
  , iI = 0         ##<< input of mineral N
  , l = 0.96 #0.00262647*365       ##<< leaching rate of mineralN l I
  , nu = 0.9       ##<< microbial N use efficiency
  , isEnzymeMassFlux = TRUE  ##<< steady state B solution accounts for enyzme mass fluxes
  , nuP = 0.3      ##<< microbial uptake of depolymerized P, (1-nuP) is mineralized
  , cpE = 100     ##<< Sterner02, protein with no P
  , cpB = 47.3    ##<< Cleveland07, grassland
  , cpBW = 61     ##<< NRSOM Tipping16
  , cpIR = 61
  , cpIL = 919    ##<< NPSOM Tipping 16
  , iBP = 0.38 * 10.57 # start with same as N
  , cW = 0.8 # proportion of cell wall within microbial biomass
  , B0 = 0  # minimal biomass, below which predation rate is zero
  , tauP = 0.1 # slope of predation rate with biomass
  , nuP = 0.01 # DOP almost all mineralized before taken up
)
parms0 <- within(parms0,{
  kmR <- kmL <- km
  eps1 <- eps2 <- eps
  cnER <- cnEL <- cnE
  kNR <- kNL <- kN
  kIPlant <- iL / cnIL	# same litter input as plant uptake
  kIPlant <- 0			# no plant uptake
  lP <- l       # leaching rate of inorganic P equals that of N
  kIPPlant <- kIPlant  # plant uptake rate of P equals that of N
  iIP <- l      # assume no P inputs compensate for leaching
})
# assure no P-limitation
parms0 <- within(parms0,{
  lP <- kIPlant <- 0    # no leading of P
  iBP <- 1e6            # no limit of P immobilization flux
})
# neglect influence of inorganic pools (keep them at fixed initial values)
parms0$isFixedI <- TRUE
```

For simplicity model drivers are currently prescibed as constant parameters.
```{r}
drivers <- list(
  iR = 0           ##<< input of residue litter
  , iL = 400       ##<< input of labile litter
  , cnIR = 7       ##<< C/N ratio of the residue litter input
  , cnIL = 30      ##<< C/N ratio of labile litter input
  , cpIR = 40      ##<< C/P ratio of the residue litter input
  , cpIL = 919     ##<< C/P ratio of the labile litter input, NPSOM Tipping16
  , iI = 0         ##<< input of mineral N
  , iIP = 0        ##<< input of labile mineral P
)
parms0 <- c(parms0[setdiff(names(parms0),names(drivers))], drivers)
```

The initial pools must be specified. 
```{r}
x0 <- c( #aE = 0.001*365
  B = 17                    ##<< microbial biomass in C units
  , R = 1100                ##<< C residue OM
  , RN = 1100/parms0$cnIR   ##<< N residue OM 
  , RP = 1100/parms0$cpIR   ##<< P residue OM 
  , L = 100                 ##<< C litter OM
  , LN = 100/parms0$cnIL    ##<< N litter OM 
  , LP = 100/parms0$cpIL    ##<< P litter OM 
  , I =  0.1               ##<< inorganic N pool 
  , IP =  0.01              ##<< inorganic P pool 
  , alpha = 0.5             ##<< microbial community partitioning [0,1]
)
```

# Run to steady state
The first simulation runs to steady state for given parameters.
```{r}
  times <- c(seq(0,500, length.out = 101), 2000)
  res <- res1 <- as.data.frame(lsoda( x0, times, derivSesam4a, parms = parms0))
  xE <-  unlist(tail(res,1)) # final state
  x0S <- xE[1 + seq_along(x0)] 
```

Plot approaching the steady state.
```{r echo=FALSE}
resC <- res %>% filter(time <= 300)  %>% cueSesam # constrained period
ggplot( resC, aes(time, L)) + geom_line()
ggplot( resC, aes(time, R)) + geom_line()
# microbial turnover
resAgg <- resC %>% select(time, tvrB, tvrBPred) %>% gather(key = "variable", "value", -time)
ggplot( resAgg, aes(time, value, color = variable)) + geom_line()
# elemental limitations
resAgg <- resC %>% select(time, limC, limN, limP) %>% gather(key = "variable", "value", -time)
ggplot( resAgg, aes(time, value, color = variable, linetype = variable)) + geom_line()
resAgg <- resC %>% select(time, CsynBC, CsynBN) %>% gather(key = "variable", "value", -time)
ggplot( resAgg, aes(time, value, color = variable, linetype = variable)) + geom_line()
resAgg <- resC %>% select(time, alpha, alphaC, alphaN) %>% gather(key = "variable", "value", -time)
ggplot( resAgg, aes(time, value, color = variable, linetype = variable)) + geom_line()
# organic-inorganic N fluxes
resAgg <- resC %>% select(time, PhiTotal, PhiU, PhiB, PhiTvr) %>% gather(key = "variable", "value", -time)
ggplot( resAgg, aes(time, value, color = variable, linetype = variable)) + geom_line()
# N decomposition fluxes
resAgg <- resC %>% select(time, decNL, decNR, decNE, decNB) %>% gather(key = "variable", "value", -time)
ggplot( resAgg, aes(time, value, color = variable, linetype = variable)) + geom_line()
# synthesis and respiration
resAgg <- resC %>% select(time, uptakeC, synB, tvrB, tvrBPred, respB, respO, respTvr) %>% gather(key = "variable", "value", -time)
ggplot( resAgg, aes(time, value, color = variable, linetype = variable)) + geom_line()
# CUE
resAgg <- resC %>% select(time, cueDef, cueSyn) %>% gather(key = "variable", "value", -time)
ggplot( resAgg, aes(time, value, color = variable, linetype = variable)) + geom_line()

ggplot( resC, aes(time, B)) + geom_line()
ggplot( resC, aes(time, L + R + B)) + geom_line()
```

# CN scenarios

Starting from this steady state, several scenarios of changing C:N ratio
of litter inputs are simulated. This means for same C input, varying the N inputs.

While the original run was C-limited, perform a run that is N-limited by increasing the 
C:N ratio of litter input.
```{r}
parms <- within(parms0, cnIL <- 80)
times <- c(seq(0,500, length.out = 101), 2000)
res <- as.data.frame(lsoda( x0S, times, derivSesam4a, parms = parms))
xE <- unlist(tail(res,1)) # final state
```

Next, inspect the steady state obtained after a range of C:N litter inputs.
```{r}
cnILs <- seq(parms0$cnIL, 70, length.out = 21)
times <- c(0, 2000)
xEs <- map_dfr(cnILs, function(cnILi){
  parms <- within(parms0, cnIL <- cnILi)
  res <- as.data.frame(lsoda( x0S, times, derivSesam4a, parms = parms))
  xE <- unlist(tail(res,1)) # final state
  cbind(as.data.frame(t(xE)), cnIL = cnILi)
})
#xEs
```

Plotting steady state against C:N of litter input.
```{r echo=FALSE}
ggplot(xEs, aes(cnIL, L)) + geom_point()
ggplot(xEs, aes(cnIL, R)) + geom_point()
ggplot(xEs, aes(cnIL, B)) + geom_point()
# elemental limitations
resAgg <- xEs %>% select(cnIL, limC, limN, limP) %>% gather(key = "variable", "value", -cnIL)
ggplot( resAgg, aes(cnIL, value, color = variable, linetype = variable)) + geom_line()
resAgg <- xEs %>% select(cnIL, CsynBC, CsynBN) %>% gather(key = "variable", "value", -cnIL)
ggplot( resAgg, aes(cnIL, value, color = variable, linetype = variable)) + geom_line()
resAgg <- xEs %>% select(cnIL, alpha, alphaC, alphaN) %>% gather(key = "variable", "value", -cnIL)
ggplot( resAgg, aes(cnIL, value, color = variable, linetype = variable)) + geom_line()
# organic-inorganic N fluxes
resAgg <- xEs %>% select(cnIL, PhiTotal, PhiU, PhiB, PhiTvr) %>% gather(key = "variable", "value", -cnIL)
ggplot( resAgg, aes(cnIL, value, color = variable, linetype = variable)) + geom_line()
# microbial turnover
resAgg <- xEs %>% mutate(tvrBTotal = tvrB + tvrBPred) %>% select(cnIL, tvrBTotal, tvrB, tvrBPred) %>% gather(key = "variable", "value", -cnIL)
ggplot( resAgg, aes(cnIL, value, color = variable, linetype = variable)) + geom_line()

ggplot(xEs, aes(cnIL, L + R + B)) + geom_point()
```

```{r echo=FALSE}
# CUE
resAgg <- cueSesam(xEs) %>% select(cnIL, cueDef, cueSyn) %>% gather(key = "variable", "value", -cnIL)
ggplot( resAgg, aes(cnIL, value, color = variable, linetype = variable)) + geom_line()
```

In steady state, microbes are C -limited up to litter C:N ratios of about 55.
This seems counter-intuitive, but is actually reasonable.

In steady state, all the organic N inputs to the soil will be mineralized, i.e. 
$\Phi_{Total} = \beta_{iL} i_L$. Immobilization only occurs if the other N
mineralization fluxes (during uptake and during turnover) exceed the N litter 
inputs.

Up to this point where N actually limits microbial growth, stocks do not change.
Afterwards they decrease together with microbial biomass stocks and CUE.

Note, also that apparent CUE is lower than the intrinsic catabolic CUE, $\epsilon = 0.5$,
due to investments into enzymes and maintenance that decrease the C available for
biomass synthesis. 
Basing CUE on adding microbial biomass synthesis and measured respiration 
(including respiration by predation), $cue_{Syn}$,
yields similar estimates as those based on modeled microbial uptake, $cue_{Def}$.

## DOM and synthesis flux composition

Despite the high C/N ratios of litter, the substrate for microbial synthesis
has more nutrients, due to decomposition of N-rich residue, and turnover 
of microbial biomass and enzymes.

```{r echo=FALSE}
resAgg <- xEs %>% mutate(cnDOM = uptakeC/uptakeNOrg, cnSyn = synC/synN) %>% 
  select(cnIL, cnDOM, cnSyn) %>% gather(key = "variable", "value", -cnIL)
ggplot( resAgg, aes(cnIL, value, color = variable, linetype = variable)) + geom_line()
```

cnDOM is the C:N ratio of the disolved organic matter (DOM), 
and cnSyn is the C:N ratio of the matter flux used for biosynthesis and mineralization. 
The latter includes immobilization of nutrients and usage of starved biomass (cryptic)
growth.

Unitl a threshold elemental ratio, TER, of about 22, there is mineralization of 
microbial N uptake from DOM. At DOM C/N ratios of slightly above this value,
immobilization of inorganic N kepps the synthesis flux at the co-limitation.

Decomposition of residue OM, $dec_{NR}$ is the largest contribution to synthesis N flux.
Direct flux of microbial turnover to DOM, $dec_{NB}$ also contributes substantially, and in the 
used parameterization also the flux by enzyme production and turnover, $dec_{NE}$.
At steady state there is no microbial starvation, $starv_N$, and immobilization 
of inorganic N, $immo_N$, contributes a small share when organic N is limiting.

```{r echo=FALSE}
resAgg <- xEs %>% mutate(immoN = pmax(0, -PhiB), starvN = recycB/parms0$cnB) %>% 
  select(cnIL, decNR, decNL, decNB, decNE, immoN, starvN) %>% gather(key = "variable", "value", -cnIL)
ggplot( resAgg, aes(cnIL, value, color = variable, linetype = variable)) + geom_line()
```

