---
title: "SESAM Soil enzyme steady allocoation model"
output: html_notebook
---

The SEAM model explicitely tracks enzyme levels. In order to use the model at plot scale or in COMISSION, derive a simpler version, where enzyme levels are assumed to be in fast steady state with production and turnover.

Derivation of the steady state equations assuming revenue allocation strategy is described in file SteadyEnzyme2

## Quasi steady state equations
The results is an equation for partitioning coefficient $\alpha$ and equations for decomposition based on this coefficient instead of enzyme levels.

$$
\alpha_C = \frac{d_R}{d_R + d_L \frac{k_{mkN} + \alpha_C a_E B}{k_{mkN}+ (1 -\alpha_C) a_E B} }
$$
, where $d_R$ and $d_L$ are potential decomposition fluxes with enzymes at saturating level: $d_R = k_R R$. Parameters $k_{m}$ and $k_N$ are the enzyme half-saturation constant in decomposition and  enzyme turnover rate, respectively. Parameter $a_E$ ist the proportion of microbial biomass invested into enzymes per time and $B$ is microbial biomass carbon. This equation can be solved for $\alpha_C$.

$$
\operatorname{dec}_R = k_R R \frac{\alpha a_e B}{k_{m} k_N+\alpha a_e B} \\
\operatorname{dec}_L = k_L L \frac{(1-\alpha) a_e B}{k_{m} k_N +(1-\alpha) a_e B}
$$
Parameters $k_m$ and $k_N$, here, are occuring always as a product and can be lumped into a single parameter. The lumping, however, precludes computing the required investments of resources into enzyme production.

## Model versions
- SEAM2: enzyme explicit (soil enzyme allocation model)
- SESAM2EnzTvr: assuming enzyme levels to be in quasi steady state with other variables but still computing fluxes for enzyme production and accounting for the required N-rich ressources investments before comptuing growth. Enzyme turnover flux feeds back in part to the assimilable pool and to the residue pool and is set equal to enzyme production.
- SESAM2: neglecting mass fluxes into and from enzyme pool

## Comparison using the FACE scenario
```{r, echo=FALSE, results='hide', message=FALSE}
library(ggplot2)
library(grid)   #unit
library(reshape)  # melt
library(RColorBrewer) # brewer.pal
library(tidyr)
library(dplyr)

myColors <- brewer.pal(5,"Dark2")	
.simpleCap <- function(s) {
    paste(toupper(substring(s, 1, 1)), substring(s, 2),  sep = "")
}
colScale <- scale_colour_manual(name = "Variant",values = myColors)
.treatments <- structure(c(1,0.5), names = c("Litter input pulse","No Litter input"))
sizeScale <- scale_size_manual(name = "Treatment",values = .treatments)

themeDefault <- theme_classic()  
```

```{r echo=FALSE, eval=FALSE, results='hide', message=FALSE, warning=FALSE}
library(twDev)
oldWd <- setwd("../.."); loadPkg()
```


### Setup parameters and drivers
Mass units are $gC/m^2$ or $gN/m^2$. Time unit is $yr$.
```{r}
require(sesam)

parmsModel <- list(
  cnB = 11
  ,cnE = 3.1     # Sterner02: Protein (Fig. 2.2.), high N investment (low P) need 
    # to have low C/N for CO2 increase scenario
  ,kN = 60       ##<< /yr enzyme turnover 60 times a year, each 6 days -> fast 
    # priming pulses
  ,km = 0.05     ##<< enzyme half-saturation constant, in magnitude of enzymes, 
    # determined by kN
  ,kNB = 0.8     ##<< amount of recycling enzyme turnover by biomass (
    # added to assimilable, i.e. uptake instead of R)
  ,kR = 1/(10)        ##<< 1/(x years)       # to demonstrate changes on short time scale
  ,kL = 5        ##<< 1/(x years)     # formerly 1 year
  ,aE = 0.001*365   ##<< C biomass allocated to enzymes 1/day /microbial biomass
  ,m = 0.005*365    ##<< maintenance respiration rate   1/day /microbial biomass, Bogedom Fig. 1
  ,tau = 1/60*365  ##<< biomass turnover rate (12 days)
  ,eps = 0.5      ##<< carbon use efficiency for growth respiration
  ,epsTvr = 0.3   ##<< carbon use efficiency of microbial tvr (part by predators which respire and corresponding amount of N must be mineralized)
  ,iBN = 0.38 * 10.57 #0.0110068*365   ##<< potential immobilization flux rate (immoNPot = iBN IN)
  ,lN = 0.96       #0.00262647*365       ##<< leaching rate of mineralN lN IN
  ,nuN = 0.9     # microbial N use efficiency accounting for apparent minceralization of N during uptake in heterogeneous soils
)

parmsModel <- within(parmsModel, {
    kmkN <- km*kN     # Sesam2 only uses a lumped parameter
    kmR <- kmL <- km  # seam2 can accept parameters differing by substrate
    epsR <- epsL <- eps
    cnER <- cnEL <- cnE 
    kNR <- kNL <- kN    
    isFixedI <- TRUE  # in this demonstration keep inorganic pool fixed
})

```

For simplicity model drivers currently are prescibed as constant parameters.
```{r}
drivers <- list(
        iR = 0          ##<< input of residue litter
        ,iL = 400         # g/m2 input per year (half NPP)
        ,cnIR = 7       ##<< C/N ratio of the residue litter input
        ,cnIL = 30      ##<< C/N ratio of the labile litter input: N poor substrate, here no inorganic N supply, 
        ,plantNUp = 0   ##<< pland uptake of organic N
        ,iP = 10.57 #0.0289652*365          ##<< plant uptake of inorganic N (IN), currently as absolute flux
        ,iIN = 0         ##<< input of mineral N
        ,kINPlant = 10.57 #0.0289652*365          ##<< plant uptake iP IN
        ,iIN = 0         ##<< input of mineral N
)
parms0 <- c(parmsModel, drivers)
```

The initial pools must be specified. The Steady models do not have the enzyme pools as state variables.
```{r}
x0 <- x0Explicit <- c( #aE = 0.001*365
        B = 17                     ##<< microbial biomass 
        ,ER  = 2*parms0$km                  ##<< total enzyme pool
        ,EL  = 4*parms0$km                   ##<< total enzyme pool
        ,R = 1100                ##<< N rich substrate
        ,RN = 1100/parms0$cnIR   ##<< N rich substrate N pool
        ,L = 100                 ##<< N poor substrate
        ,LN = 100/parms0$cnIL    ##<< N poor substrate N pool
        ,IN =  0.4                ##<< inorganic pool gN/m2
)
x0Steady <- x0Explicit[c("B","R","RN","L","LN","IN")]
```

We store the information that differs between scenarios in a tibble.
```{r}
#setwd("../..");loadPkg()  # reload function, may have inserted recover
scenarios <- tibble::tibble(
  scenario = c("Explicit","SteadyEnz","Sesam2")
  ,mod = c(derivSeam2, derivSesam2EnzTvr, derivSesam2)
  ,x0  = list(x0Explicit, x0Steady, x0Steady)
)
tmp <- structure(myColors[1:nrow(scenarios)], names = scenarios$scenario)
colScale <- scale_colour_manual(name = "Model variant",values = tmp )
```

Next the simulation runs are performed.
```{r}
runFaceSimulation <- function(scen, t1S =10, t2I = 50, t3S = 50, parms = parms0){
  # spinup to steady state
  times <- seq(0,500, length.out = 101)
  #setwd("../..");loadPkg()
  . <- scen
  res <- res1 <- as.data.frame(lsoda( .$x0, times, .$mod, parms = parms))
  xE <- unlist(tail(res,1))
  #
  # t1S years of steady state recorded
  res <- res1S <- as.data.frame(lsoda( xE[1:length(.$x0) + 1], 1:t1S, .$mod, parms = parms))
  #
  # t2I years of increased input
  fInputInc = 1.2
  parmsC2 <- within(parms, {# more C in litter, but not more N 
    iL <- iL*fInputInc
    cnIL <- cnIL*fInputInc
  }) 
  times <- seq(0,t2I, length.out = 2)
  res <- res2I <- as.data.frame(lsoda( xE[1:length(.$x0) + 1], times, .$mod, parms = parmsC2))
  res2I$time <- res2I$time + t1S 
  xE2 <- unlist(tail(res2I,1))
  #
  # t3S years of initial input
  times <- seq(0,t3S, length.out = 101)
  res <- res3S <- as.data.frame(lsoda( xE2[1:length(.$x0) + 1], times, .$mod, parms = parms))
  res3S$time <- res3S$time + t1S + t2I
  xE3 <- tail(res3S,1)
  resc <- rbind( res1S, res2I[-1,], res3S[-1,])
  resc$scen <- scen$scenario
  resc
}
.tmp.f <- function(){
  scen <- unlist(filter(scenarios, scenario == "Explicit"), recursive = FALSE)
  scen <- unlist(filter(scenarios, scenario == "SteadyEnz"), recursive = FALSE)
  scen <- unlist(filter(scenarios, scenario == "Sesam2"), recursive = FALSE)
  resScen <- runFaceSimulation( scen )
}
resScen <- scenarios %>% rowwise() %>% do( runFaceSimulation(.) )
```


```{r}
dsp <- resScen %>% mutate_(RL = ~(R+L)) %>% select_(~time, ~scen, ~L, ~R, ~RL) %>% gather_("Pool","value", c("L","R","RL"))
p2f <- ggplot( dsp, aes(x = time, y = value, col = scen, linetype = scen)) + geom_line(size = 1) +
  facet_grid(Pool ~ .,scales = "free") + 
  xlab("Time (yr)") + ylab(expression(Carbon~stock~(gCm^{-2}))) + #labs(linetype = "Substrate pool") +
  themeDefault +
  #scale_colour_discrete(drop = TRUE,limits = levels(dsp$Allocation)) +
  theme(legend.position = c(0.95,1.04), legend.justification = c(1,1)) +
  theme(panel.border = element_rect(colour = "black", fill = NA)) +
  theme(strip.background = element_blank()) +			
  theme(panel.grid.major.x = element_line(colour = "grey75")) +
  c()
#print(p2f + colScale)
print(p2f)
```

The SteadyEnz version does not recover stocks at the end of increased C-input.
There are problems with several steady states, try tracking alpha as a state 
variable.

## TODO
- compare to versions of allocation coefficient
    + based on potential decomposition fluxes only
