---
title: "SESAM CUE in substrate pulse"
output: 
  pdf_document:
    keep_tex: true
  html_notebook: default
---

## Purpose
Kaiser14 showed emergent CUE stabilized during substrate pulse by
microbial competition.
Here we simulate a similar scenario using Sesam3a

```{r, echo=FALSE, results='hide', message=FALSE}
library(ggplot2)
library(grid)   #unit
#library(reshape)  # melt
library(RColorBrewer) # brewer.pal
library(tidyr)
library(dplyr)
library(knitr)
library(forcats)

myColors <- brewer.pal(5,"Dark2")	
.simpleCap <- function(s) {
    paste(toupper(substring(s, 1, 1)), substring(s, 2),  sep = "")
}
colScale <- scale_colour_manual(name = "Variant",values = myColors)
.treatments <- structure(c(1,0.5), names = c("Litter input pulse","No Litter input"))
sizeScale <- scale_size_manual(name = "Treatment",values = .treatments)

themeTw <- theme_bw(base_size = 10) + theme(axis.title = element_text(size = 9))
#themeTw <- theme_bw(base_size = 11) + theme(axis.title = element_text(size = 10))
figWidthPage <- 18/2.54; figHeightPage <- figWidthPage/1.6 
#, fig.height=4.43, fig.width=7.1
figWidthWide <- 12/2.54; figHeightWide <- figWidthWide/1.6
#, fig.height=2.95, fig.width=4.72
figWidthCol <- 8.3/2.54; figHeightCol <- figWidthCol/1.6
#, fig.height=2.04, fig.width=3.27
opts_chunk$set(
   # out.extra='style="display:block; margin: auto"'
    #, fig.align="center"
    #, fig.width=4.6, fig.height=3.2
    #fig.width = 6, fig.height = 3.75 #goldener Schnitt 1.6
    #fig.width = figWidthWide, fig.height = figHeightWide #goldener Schnitt 1.6 #GI single column
    fig.width = figWidthPage, fig.height = figHeightPage #goldener Schnitt 1.6 #GI single column
    #, dev.args = list(pointsize = 11)
    , dev = c('pdf','png')
    , echo = FALSE
    )
knit_hooks$set(spar  =  function(before, options, envir) {
    if (before) {
        par( las = 1 )                   #also y axis labels horizontal
        par(mar = c(2.0,3.3,0,0) + 0.3 )  #margins
        par(tck = 0.02 )                          #axe-tick length inside plots             
        par(mgp = c(1.1,0.2,0) )  #positioning of axis title, axis labels, axis
    }
})

themeDefault <- themeTw#theme_classic()  
themePres <- theme_classic()
```


### Setup parameters and drivers
Mass units are $gC/m^2$ or $gN/m^2$. Time unit is $yr$.
```{r}
require(sesam, quietly = TRUE)

parmsModelN <- list(
  cnB = 11
  ,cnE = 3.1     # Sterner02: Protein (Fig. 2.2.), high N investment (low P) need 
    # to have low C/N for CO2 increase scenario
  ,kN = 60       ##<< /yr enzyme turnover 60 times a year, each 6 days -> fast 
    # priming pulses
  ,km = 0.05     ##<< enzyme half-saturation constant, in magnitude of enzymes, 
    # determined by kN
  ,kNB = 0.8     ##<< amount of recycling enzyme turnover by biomass (
    # added to assimilable, i.e. uptake instead of R)
  #,kR = 1/(10)        ##<< 1/(x years) # to demonstrate changes on short time scale
  ,kR = 1/(20)        ##<< 1/(x years) # to demonstrate changes on short time scale
  ,kL = 5           ##<< 1/(x years)   # formerly 1 year
  ,aE = 0.001*365   ##<< C biomass allocated to enzymes 1/day /microbial biomass
  ,m = 0.005*365    ##<< maintenance respiration rate   1/day /microbial biomass, 
  ## Bogedom Fig. 1
  ,tau = 1/60*365  ##<< biomass turnover rate (12 days)
  ,eps = 0.5      ##<< carbon use efficiency for growth respiration
  #,epsTvr = 0.3   ##<< carbon use efficiency of microbial tvr (part by predators 
  ,epsTvr = 0.45   ##<< carbon use efficiency of microbial tvr (part by predators 
  ## which respire and corresponding amount of N must be mineralized)
  #,iBN = 0.38 * 10.57 #0.0110068*365   ##<< potential immobilization flux rate 
  ,iBN = 0.4 ##<< potential immobilization flux rate 
  ## (immoNPot = iBN IN)
  ,lN = 0.96       #0.00262647*365     ##<< leaching rate of mineralN lN IN
  ,nuN = 0.9     # microbial N use efficiency accounting for apparent 
  ## minceralization of N during uptake in heterogeneous soils
  , isEnzymeMassFlux = TRUE  ##<< compute enzyme mass fluxes
  ,kINPlant = 10.57 #0.0289652*365     ##<< plant uptake rate first order of IN
)

parmsModelN <- within(parmsModelN, {
  kmN <- km*kN     # Sesam2 only uses a lumped parameter instead of km and kN
  kmR <- kmL <- km  # seam2 can accept parameters differing by substrate
  cnER <- cnEL <- cnE 
  kNR <- kNL <- kN    
  #isFixedI <- TRUE  # in this demonstration keep inorganic pool fixed
})
```


For simplicity model drivers currently are prescribed as constant parameters.
```{r}
driversN <- list(
  iR = 0          ##<< input of residue litter
  ,iL = 40 #400         # g/m2 input per year (half NPP)
  ,cnIR = 7       ##<< C/N ratio of the residue litter input
  ,cnIL = 30      ##<< C/N ratio of the labile litter input: 
  ## N poor substrate, here no inorganic N supply, 
  ,plantNUp = 0   ##<< pland uptake of organic N
  ## currently as absolute flux
  ,iIN = 0         ##<< input of mineral N
)
```


The initial pools must be specified. 
```{r}
x0N <- c( #aE = 0.001*365
        B = 17                     ##<< microbial biomass 
        #,ER  = 2*parms0$km                  ##<< total enzyme pool
        #,EL  = 4*parms0$km                   ##<< total enzyme pool
        ,R = 1100                ##<< N rich substrate
        ,RN = 1100/driversN$cnIR   ##<< N rich substrate N pool
        ,L = 10#100                 ##<< N poor substrate
        ,LN = 10/driversN$cnIL #100/parms0$cnIL    ##<< N poor substrate N pool
        ,IN =  0.04                ##<< inorganic pool gN/m2
        #, alpha = 0.5            ##<< microbial community partitioning to R [0,1]
        , alpha = 0.3            ##<< microbial community partitioning to R [0,1]
)
```

```{r}
extend_CNP_setting <- function(x0, parms, drivers) {
  driversP <- within(drivers,{
    cpIR <- 40
    cpIL <- 40*3
    iIP <- 0.001  # very small
  })
  parmsP <- within(parms,{
    lP <- 0.001  # very small to not run into P limitation  
    nuP <- 0.0     # microbial P use efficiency: no direct P uptake
     cpE <- 1e5
     cpB <- 1e8 # for not running into P limitation, here assume no P required in biomass
     iBP <- 0.38 * 10.57 # start with same as N
     kIPPlant <- 0.0 #kINPlant  # plant uptake rate of P equals that of N
     e_P <- 0.0 # no plant P enzyme production
     kLP <- kL
     kRP <- kR
  })
  x0P <- c( 
        x0[c("B","R","RN")]
        ,RP = unname(x0["R"]/driversP$cpIR)
        ,x0[c("L","LN")]
        ,LP = unname(x0["L"]/driversP$cpIL)    ##<< N poor substrate N pool
        ,x0[c("IN")]
        ,IP =  150 # prevent P limitation          ##<< inorganic pool gN/m2
        ,alphaL=1- unname(x0["alpha"])
        ,alphaR=unname(x0["alpha"])
  )
  list(x0=x0P, parms=parmsP, drivers=driversP)
}
```

```{r}
xpd <- extend_CNP_setting(x0N, parmsModelN, driversN)
x0 <- xpd$x0
parms0 <- c(xpd$parms, xpd$drivers)

parms <- parms0 <- within( parms0, {
  # change inorganic pool only by mobilization/immobilization flux
  kINPlant <- 0
  plantNUpAbs <- 0#iL/cnIL
  lN <- 0 
  lP <- 0
  iIN <- 0
  iIP <- 0
  #isFixedI <- TRUE
})

x0FixedAlpha <- x0
x0FixedAlpha["alphaR"] <- 0.3
x0FixedAlpha["alphaL"] <- 1-x0FixedAlpha["alphaR"]
parmsFixedAlpha <- within(parms0, isFixedAlpha <- TRUE)

parmsRelative <- within(parms0,isRelativeAlpha<-TRUE)
parmsOptimal <- within(parms0,isOptimalAlpha<-TRUE)
```


Simulate steady state with low L inputs and then L pulse by changed L state.
```{r}
simLPulse <- function(x0, parms, cn0=75){
  # # spinup to steady state at low substrate inputs
  # times <- seq(0,2000, length.out = 2)
  # #times <- c(0,exp(seq(log(0.02),log(2000), length.out=100)))
  # #setwd("../..");loadPkg()
  # res <- res1 <- as.data.frame(lsoda( x0, times, derivSesam3P, parms = parms))
  # #plot(R~time, res)
  # xE <- unlist(tail(res,1))
  # #xE["IN"] <- 0.04  # decrease IN pool # does not help filled up by mircobial tvr
  #
  # increase L, no L input any more
  parms$iL <- 0
  x0P = x0 #xE[1:length(x0) + 1]
  x0P["B"] <- .3
  x0P["L"] <- 100
  x0P["LN"] <- 100/cn0
  x0P["R"] <- 1200
  x0P["RN"] <- 1200/parms$cnB
  x0P["IN"] <- 1
  timesP <- c(0,exp(seq(log(0.002),log(10), length.out = 200)))
  res <- resP <- as.data.frame(lsoda( x0P, timesP, derivSesam3P, parms = parms))
  .tmp.f <- function(){
    derivSesam3P(0, x0P, parms)
    res <- resP <- as.data.frame(lsoda( x0P, c(0,0.1), derivSesam3P, parms = parms))
    xE <- unlist(resP[nrow(resP),1+seq_along(x0)])
    derivSesam3P(0, xE, parms)
    plot(alphaR ~ time, res); abline(h=1)
    tail(filter(res, alphaR < 1),1)
  }
  res
}
# x0 <- xpd$x0; parms<-parms0; cn0=50
resCN50 <- res <- simLPulse(x0, parms0, cn0 = 50)
resCN70 <- simLPulse(x0, parms0, cn0 = 70)
resCN90 <- simLPulse(x0, parms0, cn0 = 90)
resRelCN50  <- simLPulse(x0, parmsRelative, cn0 = 50)
resRelCN70  <- simLPulse(x0, parmsRelative, cn0 = 70)
resRelCN90  <- simLPulse(x0, parmsRelative, cn0 = 90)
# x0 <- xpd$x0; parms<-parmsOptimal; cn0=50
resOptCN50  <- simLPulse(x0, parmsOptimal, cn0 = 50)
resOptCN70  <- simLPulse(x0, parmsOptimal, cn0 = 70)
resOptCN90  <- simLPulse(x0, parmsOptimal, cn0 = 90)
#
resAlpha30CN30  <- simLPulse(x0FixedAlpha, parmsFixedAlpha, cn0 = 50)
resAlpha30CN70  <- simLPulse(x0FixedAlpha, parmsFixedAlpha, cn0 = 70)
resAlpha30CN90  <- simLPulse(x0FixedAlpha, parmsFixedAlpha, cn0 = 90)

.tmp.f <- function(){
  # x0 <- x0FixedAlpha; parms <- parmsFixedAlpha; cn0=50
  # x0Alpha10 <- x0Steady; x0Alpha10["alpha"] <- 0.1
  # resAlpha10  <- simLPulse(x0Alpha10, parmsFixedAlpha)
  # x0Alpha90 <- x0Steady; x0Alpha90["alpha"] <- 0.6
  # resAlpha90  <- simLPulse(x0Alpha90, parmsFixedAlpha)
}
resP <- rbind(
  cbind(scen = "Derivative", cn0 = 50, resCN50)
  ,cbind(scen = "Derivative", cn0 = 70, resCN70)
  ,cbind(scen = "Derivative", cn0 = 90, resCN90)
  ,cbind(scen = "Relative", cn0 = 50, resRelCN50)
  ,cbind(scen = "Relative", cn0 = 70, resRelCN70)
  ,cbind(scen = "Relative", cn0 = 90, resRelCN90)
  ,cbind(scen = "Optimal", cn0 = 50, resOptCN50)
  ,cbind(scen = "Optimal", cn0 = 70, resOptCN70)
  ,cbind(scen = "Optimal", cn0 = 90, resOptCN90)
  ,cbind(scen = "Fixed", cn0 = 50, resAlpha30CN30)
  ,cbind(scen = "Fixed", cn0 = 70, resAlpha30CN70)
  ,cbind(scen = "Fixed", cn0 = 90, resAlpha30CN90)
)
```
```{r eval=FALSE}
names(resP)
unique(resP$scen)
df = filter(resP, cn0==50 & scen=="adaptable")
plot(synB ~ time, df)
plot(iL ~ time, df)
str(parms0)
```


```{r}
vars <- c("L","R","alphaR","B", "cueDef", "cueSyn") 
vars <- c("limN","alphaR","B","cueSyn","R","L") 
#vars <- c("L","R","alpha","B", "IN","cueDef", "cueSyn") 
#vars <- c("L","R","alpha","B","RL","cnRL") 
#vars <- c("alpha","B", "cnL", "PhiNB","PhiNTotal","immoNPot", "CsynBC","CsynBN") 
dsp0 <- resP %>% filter(.data$time < 3) %>%  
  #rename(uptakeC = uC, respTotal = resp) %>% 
  mutate(tvrBPred = 0.0) %>% 
  cueSesam() %>% 
  mutate(cn0 = as.factor(cn0)) %>% 
  mutate(S = R + L, cnS = (R + L)/(RN + LN)) %>% 
  mutate(relS = S/S[1], relL = L/L[1], fracDecomp = 1 - relL) 
  #select_(~time, ~scen, ~L, ~R, ~RL) %>% gather_("Pool","value", c("L","R","RL"))
dsp <- dsp0 %>% select(all_of(c("time","relS","relL","fracDecomp","scen","cn0", vars))) %>% gather("Pool","value", vars)
p2f <- dsp %>% 
  mutate(scen = fct_relevel(.data$scen, c("Optimal","Derivative","Relative"))) %>% 
  ggplot(aes(x = time, y = value, col = cn0, linetype = cn0)) + geom_line(linewidth = 1) +
  facet_grid(Pool ~ scen,scales = "free") + 
  xlab("Time (yr)") + 
  #xlab("Substrate (proportion of initial)") + scale_x_reverse() +
  #xlab("Fraction C decomposed") +
  #ylab(expression(Carbon~stock~(gCm^{-2}))) + #labs(linetype = "Substrate pool") +
  themeDefault +
  #scale_colour_discrete(drop = TRUE,limits = levels(dsp$Allocation)) +
  theme(legend.position = c(0.95,1.04), legend.justification = c(1,1)) +
  theme(panel.border = element_rect(colour = "black", fill = NA)) +
  theme(strip.background = element_blank()) +			
  theme(panel.grid.major.x = element_line(colour = "grey75")) +
  theme(axis.title.y = element_blank()) +
  c()
#print(p2f + colScale)
print(p2f)
#ggsave("Sesam_OptimMethods_Pulse_time.png", width=4.8, height=3.8)
```

In the optimal scenario, $\alpha_R$ goes to lower values also for intermediate
initial substrate C:N ratio and much sooner goes to high values in for 
high initial substrate C:N ratio.

In both the optimal and the derivative scenario, $\alpha_R$ goes to zero when
N is not limiting any more, resulting in the accumulation of a larger fraction
of the substrate ending up in the residue pool.
This is because with lower biomass, the return does not strongly decrease with 
community allocation any more and allocating to the enzymes with lower potential
return is not as beneficial any more.
Hence, it results in a higher time-integrated CUE with larger differences
between the initial litter C:N ratios.

Initially there is more N in the labile decomposition in absolute terms despite
its low relative low N contents, i.e. high C:N ratio. Hence, microbial community
shifts towards only L-degrading with high overflow respiration and low CUE.

After the initial shift towards growing on the labile substrate,
CUE is recovered to higher values sooner with flexible allocation due to
increased allocation to N-rich residue pool.

Biomass:
```{r}
vars <- c("B","respO","CsynBN","cueSyn") 
dsp0 <- resP %>% filter(.data$time < 1) %>% cueSesam() 
dsp <- dsp0 %>% select(all_of(c("time", "scen", "cn0", vars))) %>% gather("Pool","value", vars)
p2f <- dsp %>% 
  mutate(scen = fct_relevel(.data$scen, c("Optimal","Derivative","Relative"))) %>% 
  ggplot(aes(x = time, y = value, col = scen, linetype = scen)) + geom_line(linewidth = 1) +
  facet_grid(Pool ~ cn0,scales = "free") + 
  xlab("Time (yr)") + 
  #xlab("Substrate (proportion of initial)") + scale_x_reverse() +
  #xlab("Fraction C decomposed") +
  #ylab(expression(Carbon~stock~(gCm^{-2}))) + #labs(linetype = "Substrate pool") +
  themeDefault +
  #scale_colour_discrete(drop = TRUE,limits = levels(dsp$Allocation)) +
  theme(legend.position = c(0.95,1.04), legend.justification = c(1,1)) +
  theme(panel.border = element_rect(colour = "black", fill = NA)) +
  theme(strip.background = element_blank()) +			
  theme(panel.grid.major.x = element_line(colour = "grey75")) +
  theme(axis.title.y = element_blank()) +
  c()
#print(p2f + colScale)
print(p2f)
#ggsave("Sesam_OptimMethods_Pulse_approaches.png", width=4.8, height=3.8)


```
Why is biomass increasing less with optimal compared to relative scenario 
in non-extreme cn0 scenarios?
Avoids the large initial overflow respiration.
The optimal scenario trades in slighly higher biomass synthesis for a large
overflow respiration.

```{r}
#vars2 <- c("cueDef", "cueSyn") 
vars2 <- c("cueDef") 
dsp2 <- dsp0 %>% select(all_of(c("time","relS","relL","fracDecomp","scen","cn0", vars2))) %>% gather("Pool","value", vars2)
p2 <- dsp2 %>% 
  mutate(scen = fct_relevel(.data$scen, c("Optimal","Derivative","Relative"))) %>% 
  ggplot(aes(x = fracDecomp, y = value, col = cn0, linetype = cn0)) + 
#p2 <- ggplot( dsp2, aes(x = fracDecomp, y = value, col = cn0, linetype = cn0)) + 
  geom_line(linewidth = 1) +
  facet_grid(~ scen) + #,scales = "free") + 
  #xlab("Time (yr)") + 
  #xlab("Substrate (proportion of initial)") + scale_x_reverse() +
  xlab("Fraction C decomposed") +
  ylab("Carbon use efficiency") +
  expand_limits(y = 0.1) +
  #themeDefault +
  theme_classic(base_size = 18) + # presentation
  #scale_colour_discrete(drop = TRUE,limits = levels(dsp$Allocation)) +
  #theme(legend.position = c(0.95,1.04), legend.justification = c(1,1)) +
  theme(legend.position = c(0.95,0.05), legend.justification = c(1,0)) +
  theme(panel.border = element_rect(colour = "black", fill = NA)) +
  theme(strip.background = element_blank()) +			
  #theme(panel.grid.major.x = element_line(colour = "grey75")) +
  c()
#print(p2f + colScale)
print(p2)
```



```{r}
#vars2 <- c("cueDef", "cueSyn") 
vars2 <- c("cueDef") 
dsp2 <- dsp0 %>% select(all_of(c("time","relS","relL","fracDecomp","scen","cn0", vars2))) %>% gather("Pool","value", vars2)
p2 <- dsp2 %>% filter(time <= 3) %>% 
  ggplot( aes(x = time, y = value, col = cn0, linetype = cn0)) + 
#p2 <- ggplot( dsp2, aes(x = fracDecomp, y = value, col = cn0, linetype = cn0)) + 
  geom_line(linewidth = 1) +
  facet_grid(~ scen) + #,scales = "free") + 
  #xlab("Time (yr)") + 
  #xlab("Substrate (proportion of initial)") + scale_x_reverse() +
  xlab("Time (yr)") +
  ylab("Carbon use efficiency") +
  #themeDefault +
  theme_classic(base_size = 18) + # presentation
  #scale_colour_discrete(drop = TRUE,limits = levels(dsp$Allocation)) +
  #theme(legend.position = c(0.95,1.04), legend.justification = c(1,1)) +
  theme(legend.position = c(0.95,0.05), legend.justification = c(1,0)) +
  theme(panel.border = element_rect(colour = "black", fill = NA)) +
  theme(strip.background = element_blank()) +			
  #theme(panel.grid.major.x = element_line(colour = "grey75")) +
  c()
#print(p2f + colScale)
print(p2 + expand_limits(y = 0.1))
```


```{r simPulse, fig.height=2, fig.width=3.27}
p2p <- p2 + 
  expand_limits(y = 0.1) +
  theme_classic(base_size = 10) +
  theme(legend.position = c(0.98,0.02), legend.justification = c(1,0)) +
  theme(panel.border = element_rect(colour = "black", fill = NA)) +
  theme(strip.background = element_blank()) +	
  theme(legend.title = element_blank()) +
  c()  
print(p2p)
```

```{r}
# Compare the initial derivative at cn90

```

