---
title: "SESAM CUE in substrate pulse"
output: 
  pdf_document:
    keep_tex: true
  html_notebook: default
---

## Purpose
Kaiser14 showed emergent CUE stabilized during substrate pulse by
microbial competition.
Here we simulate a similar scenario using Sesam3a

```{r, echo=FALSE, results='hide', message=FALSE}
library(ggplot2)
library(grid)   #unit
#library(reshape)  # melt
library(RColorBrewer) # brewer.pal
library(tidyr)
library(dplyr)
library(knitr)
library(forcats)

myColors <- brewer.pal(5,"Dark2")	
myLinetypes <- c("solid","dashed","dotdash", "dotted","longdash","twpdash")
#adjust color default discrete color scale
scale_colour_discrete <- function(...) scale_colour_manual(...,values=myColors)

.simpleCap <- function(s) {
    paste(toupper(substring(s, 1, 1)), substring(s, 2),  sep = "")
}
colScale <- scale_colour_manual(name = "Variant",values = myColors)
.treatments <- structure(c(1,0.5), names = c("Litter input pulse","No Litter input"))
sizeScale <- scale_size_manual(name = "Treatment",values = .treatments)

.approaches <- c("Optimal","Derivative","Relative","Fixed")
scale_color_approach <- scale_colour_manual(name = "Approach",values = setNames(myColors[seq_along(.approaches)],.approaches) )
scale_linetype_approach <- scale_linetype_manual(name = "Approach",values = setNames(myLinetypes[seq_along(.approaches)],.approaches) )

adjust_facets <- function(theme){
  theme +
    theme(panel.border = element_rect(colour = "black", fill = NA)) +
    theme(strip.background = element_blank()) 
}
themePaper <- (theme_classic(base_size = 10) + theme(axis.title = element_text(size = 9))) %>% 
  adjust_facets()
themePres <- theme_classic(base_size = 18) %>% 
  adjust_facets()
theme_set(themePres) # also adjust fig.with in opts_chunk below to paper Wide

#figWidthPage <- 18/2.54; figHeightPage <- figWidthPage/1.6 
figWidthPage <- 12/2.54; figHeightPage <- figWidthPage/1.6 
#, fig.height=4.43, fig.width=7.1
#figWidthWide <- 12/2.54; figHeightWide <- figWidthWide/1.6
figWidthWide <- 3.8; figHeightWide <- 5
#, fig.height=2.95, fig.width=4.72
figWidthCol <- 8.3/2.54; figHeightCol <- figWidthCol/1.6
#, fig.height=2.04, fig.width=3.27
opts_chunk$set(
   # out.extra='style="display:block; margin: auto"'
    #, fig.align="center"
    #, fig.width=4.6, fig.height=3.2
    #fig.width = 6, fig.height = 3.75 #goldener Schnitt 1.6
    fig.width = figWidthWide, fig.height = figHeightWide #goldener Schnitt 1.6 #GI single column
    #fig.width = figWidthPage, fig.height = figHeightPage #goldener Schnitt 1.6 #GI single column
    #, dev.args = list(pointsize = 11)
    , dev = c('pdf','png')
    , echo = FALSE
    )
knit_hooks$set(spar  =  function(before, options, envir) {
    if (before) {
        par( las = 1 )                   #also y axis labels horizontal
        par(mar = c(2.0,3.3,0,0) + 0.3 )  #margins
        par(tck = 0.02 )                          #axe-tick length inside plots             
        par(mgp = c(1.1,0.2,0) )  #positioning of axis title, axis labels, axis
    }
})


#themeUsed <- c(themeUsed, scale_colour_brewer(type="qual",palette="Dark2"))
replace_labels <- function(labels, labelmap) {
  sapply(labels, function(label){
    if (is.null(labelmap[[label]])) label else labelmap[[label]]
  })
}


```


### Setup parameters and drivers
Mass units are $gC/m^2$ or $gN/m^2$. Time unit is $yr$.
```{r}
require(sesam, quietly = TRUE)

parmsModelN <- list(
  cnB = 11
  ,cnE = 3.1     # Sterner02: Protein (Fig. 2.2.), high N investment (low P) need 
    # to have low C/N for CO2 increase scenario
  ,kN = 60       ##<< /yr enzyme turnover 60 times a year, each 6 days -> fast 
    # priming pulses
  ,km = 0.05     ##<< enzyme half-saturation constant, in magnitude of enzymes, 
    # determined by kN
  ,kNB = 0.8     ##<< amount of recycling enzyme turnover by biomass (
    # added to assimilable, i.e. uptake instead of R)
  #,kR = 1/(10)        ##<< 1/(x years) # to demonstrate changes on short time scale
  ,kR = 1/(20)        ##<< 1/(x years) # to demonstrate changes on short time scale
  ,kL = 5           ##<< 1/(x years)   # formerly 1 year
  ,aE = 0.001*365   ##<< C biomass allocated to enzymes 1/day /microbial biomass
  ,m = 0.005*365    ##<< maintenance respiration rate   1/day /microbial biomass, 
  ## Bogedom Fig. 1
  ,tau = 1/60*365  ##<< biomass turnover rate (12 days)
  ,eps = 0.5      ##<< carbon use efficiency for growth respiration
  #,epsTvr = 0.3   ##<< carbon use efficiency of microbial tvr (part by predators 
  ,epsTvr = 0.45   ##<< carbon use efficiency of microbial tvr (part by predators 
  ## which respire and corresponding amount of N must be mineralized)
  #,iBN = 0.38 * 10.57 #0.0110068*365   ##<< potential immobilization flux rate 
  ,iBN = 0.4 ##<< potential immobilization flux rate 
  ## (immoNPot = iBN IN)
  ,lN = 0.96       #0.00262647*365     ##<< leaching rate of mineralN lN IN
  ,nuN = 0.9     # microbial N use efficiency accounting for apparent 
  ## minceralization of N during uptake in heterogeneous soils
  , isEnzymeMassFlux = TRUE  ##<< compute enzyme mass fluxes
  ,kINPlant = 10.57 #0.0289652*365     ##<< plant uptake rate first order of IN
)

parmsModelN <- within(parmsModelN, {
  kmN <- km*kN     # Sesam2 only uses a lumped parameter instead of km and kN
  kmR <- kmL <- km  # seam2 can accept parameters differing by substrate
  cnER <- cnEL <- cnE 
  kNR <- kNL <- kN    
  #isFixedI <- TRUE  # in this demonstration keep inorganic pool fixed
})
```


For simplicity model drivers currently are prescribed as constant parameters.
```{r}
driversN <- within(list(
  iR = 0          ##<< input of residue litter
  ,iL = 400         # g/m2 input per year (half NPP)
  ,cnIR = 7       ##<< C/N ratio of the residue litter input
  ,cnIL = 30      ##<< C/N ratio of the labile litter input: 
  ## N poor substrate, here no inorganic N supply, 
  ## currently as absolute flux
  ,iIN = 0         ##<< input of mineral N
),{
  plantNUp <- iL/cnIL   ##<< pland uptake of organic N  
})
```

The initial pools must be specified but model will be run into steady state.
```{r}
x0N <- c( #aE = 0.001*365
        B = 17                     ##<< microbial biomass 
        #,ER  = 2*parms0$km                  ##<< total enzyme pool
        #,EL  = 4*parms0$km                   ##<< total enzyme pool
        ,R = 1100                ##<< N rich substrate
        ,RN = 1100/driversN$cnIR   ##<< N rich substrate N pool
        ,L = 10#100                 ##<< N poor substrate
        ,LN = 10/driversN$cnIL #100/parms0$cnIL    ##<< N poor substrate N pool
        ,IN =  0.04                ##<< inorganic pool gN/m2
        , alpha = 0.5            ##<< microbial community partitioning to R [0,1]
)
```


```{r}
extend_CNP_setting <- function(x0, parms, drivers) {
  driversP <- within(drivers,{
    cpIR <- 40
    cpIL <- 40*3
    iIP <- 0.001  # very small
    plantPUp <- iL/cpIL   ##<< pland uptake of organic N  
  })
  parmsP <- within(parms,{
    lP <- 0.001  # very small to not run into P limitation  
    nuP <- 0.0     # microbial P use efficiency: no direct P uptake
     cpE <- 1e5
     cpB <- 1e8 # for not running into P limitation, here assume no P required in biomass
     iBP <- 0.38 * 10.57 # start with same as N
     kIPPlant <- 0.0 #kINPlant  # plant uptake rate of P equals that of N
     e_P <- 0.0 # no plant P enzyme production
     kLP <- kL
     kRP <- kR
  })
  x0P <- c( 
        x0[c("B","R","RN")]
        ,RP = unname(x0["R"]/driversP$cpIR)
        ,x0[c("L","LN")]
        ,LP = unname(x0["L"]/driversP$cpIL)    ##<< N poor substrate N pool
        ,x0[c("IN")]
        ,IP =  150 # prevent P limitation          ##<< inorganic pool gN/m2
        ,alphaL=1- unname(x0["alpha"])
        ,alphaR=unname(x0["alpha"])
  )
  list(x0=x0P, parms=parmsP, drivers=driversP)
}
```

```{r}
xpd <- extend_CNP_setting(x0N, parmsModelN, driversN)
x0 <- xpd$x0
parms0 <- c(xpd$parms, xpd$drivers)

parms <- parms0 <- within( parms0, {
  # change inorganic pool only by mobilization/immobilization flux
  kINPlant <- 0
  plantNUpAbs <- iL/cnIL
  lN <- 0.96 # to not let IN increase to infinity
  lP <- 1e-2
  iIN <- 0
  iIP <- 0
  #isFixedI <- TRUE
})

# x0FixedAlpha <- x0
# x0FixedAlpha["alphaR"] <- 0.3
# x0FixedAlpha["alphaL"] <- 1-x0FixedAlpha["alphaR"]
parmsFixedAlpha <- within(parms0, isFixedAlpha <- TRUE)
parmsRelative <- within(parms0,isRelativeAlpha<-TRUE)
parmsOptimal <- within(parms0,isOptimalAlpha<-TRUE)
```

```{r}
get_storage_parameters <- function(p, iL = 1) {
  # only some minor dust input and negligible leaching
  p$iL <- iL
  p$iIN <- 0
  p$iIP <- 0
  p$lN <- 1e-2
  p$lP <- 1e-2
  p
}
```


First simulate a steady state microbial biomass before storage with field
parameters, and then storage for two years with low L input.
```{r}
times_fresh <- c(0,exp(seq(log(0.002),log(2000), length.out = 200)))
times_storage <- c(0,exp(seq(log(0.002),log(2), length.out = 200)))
res <- as.data.frame(lsoda( x0, times_fresh, derivSesam3P, parms = parms0))
  x0f <- unlist(res[nrow(res),1+seq_along(x0)])
  res <- as.data.frame(lsoda( x0f, times_storage, derivSesam3P, parms = get_storage_parameters(parms0)))
  x0_deriv <- unlist(res[nrow(res),1+seq_along(x0)])
res <- as.data.frame(lsoda( x0, times_fresh, derivSesam3P, parms = parmsRelative))
  x0f <- unlist(res[nrow(res),1+seq_along(x0)])
  res <- as.data.frame(lsoda( x0f, times_storage, derivSesam3P, parms = get_storage_parameters(parmsRelative)))
  x0_rel <- unlist(res[nrow(res),1+seq_along(x0)])
res <- as.data.frame(lsoda( x0, times_fresh, derivSesam3P, parms = parmsOptimal))
  x0f <- unlist(res[nrow(res),1+seq_along(x0)])
  res <- as.data.frame(lsoda( x0f, times_storage, derivSesam3P, parms = get_storage_parameters(parmsOptimal)))
  x0_opt <- unlist(res[nrow(res),1+seq_along(x0)])
res <- as.data.frame(lsoda( x0, times_fresh, derivSesam3P, parms = parmsFixedAlpha))
  x0f <- unlist(res[nrow(res),1+seq_along(x0)])
  res <- as.data.frame(lsoda( x0f, times_storage, derivSesam3P, parms = get_storage_parameters(parmsFixedAlpha)))
  x0_fix <- unlist(res[nrow(res),1+seq_along(x0)])
.tmp.f <- function(){
  plot(alphaR ~ time, res); abline(h=1)
  plot(alphaL ~ time, res); abline(h=1)
  plot(limN ~ time, res); abline(h=1)
  plot(cnL ~ time, res); abline(h=1)
  plot(cnR ~ time, res); abline(h=1)
  plot(B ~ time, res); abline(h=1)
  plot(R ~ time, res); abline(h=1)
  plot(L ~ time, res); abline(h=1)
  plot(limC ~ time, res); abline(h=1)
  plot(IN ~ time, res); abline(h=1)
  plot(IP ~ time, res); abline(h=1)
}
```


```{r}
simLPulse <- function(
  x0, parms, cn0=75, 
  #amend=500
  amend=100
  ){
  # # spinup to steady state at low substrate inputs
  # times <- seq(0,2000, length.out = 2)
  # #times <- c(0,exp(seq(log(0.02),log(2000), length.out=100)))
  # #setwd("../..");loadPkg()
  # res <- res1 <- as.data.frame(lsoda( x0, times, derivSesam3P, parms = parms))
  # #plot(R~time, res)
  # xE <- unlist(tail(res,1))
  # #xE["IN"] <- 0.04  # decrease IN pool # does not help filled up by mircobial tvr
  #
  # increase L, no L input any more
  parms$iL <- 0
  x0P = x0 #xE[1:length(x0) + 1]
  x0P["L"] <- x0["L"] + amend
  x0P["LN"] <- x0["LN"] + amend/cn0
  x0P["IN"] <- 1  # moisten with DI water
  x0P["IP"] <- 1
  timesP <- c(0,exp(seq(log(0.002),log(10), length.out = 200)))
  res <- resP <- as.data.frame(lsoda( x0P, timesP, derivSesam3P, parms = parms))
  .tmp.f <- function(){
    derivSesam3P(0, x0P, parms)
    res <- resP <- as.data.frame(lsoda( x0P, c(0,0.1), derivSesam3P, parms = parms))
    xE <- unlist(resP[nrow(resP),1+seq_along(x0)])
    derivSesam3P(0, xE, parms)
    plot(alphaR ~ time, res); abline(h=1)
    tail(filter(res, alphaR < 1),1)
  }
  res
}
# x0 <- xpd$x0; parms<-parms0; cn0=50
resCN50 <- res <- simLPulse(x0_deriv, get_storage_parameters(parms0), cn0 = 50)
resCN70 <- simLPulse(x0_deriv, get_storage_parameters(parms0), cn0 = 70)
resCN90 <- simLPulse(x0_deriv, get_storage_parameters(parms0), cn0 = 90)
resRelCN50  <- simLPulse(x0_rel, get_storage_parameters(parmsRelative), cn0 = 50)
resRelCN70  <- simLPulse(x0_rel, get_storage_parameters(parmsRelative), cn0 = 70)
resRelCN90  <- simLPulse(x0_rel, get_storage_parameters(parmsRelative), cn0 = 90)
# x0 <- xpd$x0; parms<-parmsOptimal; cn0=50
resOptCN50  <- simLPulse(x0_opt, get_storage_parameters(parmsOptimal), cn0 = 50)
resOptCN70  <- simLPulse(x0_opt, get_storage_parameters(parmsOptimal), cn0 = 70)
resOptCN90  <- simLPulse(x0_opt, get_storage_parameters(parmsOptimal), cn0 = 90)
#
resAlpha30CN30  <- simLPulse(x0_fix, get_storage_parameters(parmsFixedAlpha), cn0 = 50)
resAlpha30CN70  <- simLPulse(x0_fix, get_storage_parameters(parmsFixedAlpha), cn0 = 70)
resAlpha30CN90  <- simLPulse(x0_fix, get_storage_parameters(parmsFixedAlpha), cn0 = 90)

.tmp.f <- function(){
  # x0 <- x0FixedAlpha; parms <- parmsFixedAlpha; cn0=50
  # x0Alpha10 <- x0Steady; x0Alpha10["alpha"] <- 0.1
  # resAlpha10  <- simLPulse(x0Alpha10, parmsFixedAlpha)
  # x0Alpha90 <- x0Steady; x0Alpha90["alpha"] <- 0.6
  # resAlpha90  <- simLPulse(x0Alpha90, parmsFixedAlpha)
}
resP <- rbind(
  cbind(scen = "Derivative", cn0 = 50, resCN50)
  ,cbind(scen = "Derivative", cn0 = 70, resCN70)
  ,cbind(scen = "Derivative", cn0 = 90, resCN90)
  ,cbind(scen = "Relative", cn0 = 50, resRelCN50)
  ,cbind(scen = "Relative", cn0 = 70, resRelCN70)
  ,cbind(scen = "Relative", cn0 = 90, resRelCN90)
  ,cbind(scen = "Optimal", cn0 = 50, resOptCN50)
  ,cbind(scen = "Optimal", cn0 = 70, resOptCN70)
  ,cbind(scen = "Optimal", cn0 = 90, resOptCN90)
  ,cbind(scen = "Fixed", cn0 = 50, resAlpha30CN30)
  ,cbind(scen = "Fixed", cn0 = 70, resAlpha30CN70)
  ,cbind(scen = "Fixed", cn0 = 90, resAlpha30CN90)
)
```
```{r eval=FALSE}
names(resP)
unique(resP$scen)
df = filter(resP, cn0==50 & scen=="adaptable")
plot(synB ~ time, df)
plot(iL ~ time, df)
str(parms0)
```


```{r}
vars <- c("L","R","alphaR","B", "cueDef", "cueSyn") 
vars <- c("limN","alphaR","B","cueDef","R_norm","L") 
#vars <- c("L","R","alpha","B", "IN","cueDef", "cueSyn") 
#vars <- c("L","R","alpha","B","RL","cnRL") 
#vars <- c("alpha","B", "cnL", "PhiNB","PhiNTotal","immoNPot", "CsynBC","CsynBN") 
dsp0 <- resP %>% filter(.data$time < 3) %>%  
  #rename(uptakeC = uC, respTotal = resp) %>% 
  mutate(tvrBPred = 0.0) %>% 
  cueSesam() %>% 
  group_by(cn0,scen) %>% mutate(R_norm = R/R[1]) %>% ungroup() %>% 
  mutate(cn0 = as.factor(cn0))  
  #select_(~time, ~scen, ~L, ~R, ~RL) %>% gather_("Pool","value", c("L","R","RL"))
dsp <- dsp0 %>% select(all_of(c("time","scen","cn0", vars))) %>% gather("Pool","value", vars)
p2f <- dsp %>% 
  mutate(scen = fct_relevel(.data$scen, c("Optimal","Derivative","Relative"))) %>% 
  ggplot(aes(x = time, y = value, col = cn0, linetype = cn0)) + geom_line(linewidth = 1) +
  facet_grid(Pool ~ scen,scales = "free") + 
  xlab("Time (yr)") + 
  #xlab("Substrate (proportion of initial)") + scale_x_reverse() +
  #xlab("Fraction C decomposed") +
  #ylab(expression(Carbon~stock~(gCm^{-2}))) + #labs(linetype = "Substrate pool") +
  #themeDefault +
  #scale_colour_discrete(drop = TRUE,limits = levels(dsp$Allocation)) +
  themePaper +
  theme(legend.position = c(0.99,0.99), legend.justification = c(1,1)) +
  theme(legend.title = element_blank()) +
  theme(panel.grid.major.x = element_line(colour = "grey75")) +
  theme(axis.title.y = element_blank()) +
  c()
#print(p2f + colScale)
print(p2f)
#ggsave("Sesam_OptimMethods_Pulse_time.png", width=4.8, height=3.8)
```

In the optimal scenario, $\alpha_R$ goes to lower values also for intermediate
initial substrate C:N ratio and much sooner goes to high values in for 
high initial substrate C:N ratio.

In both the optimal and the derivative scenario, $\alpha_R$ goes to zero when
N is not limiting any more, resulting in the accumulation of a larger fraction
of the substrate ending up in the residue pool.
This is because with lower biomass, the return does not strongly decrease with 
community allocation any more and allocating to the enzymes with lower potential
return is not as beneficial any more.
Hence, it results in a higher time-integrated CUE with larger differences
between the initial litter C:N ratios.

Initially there is more N in the labile decomposition in absolute terms despite
its low relative low N contents, i.e. high C:N ratio. Hence, microbial community
shifts towards only L-degrading with high overflow respiration and low CUE.

After the initial shift towards growing on the labile substrate,
CUE is recovered to higher values sooner with flexible allocation due to
increased allocation to N-rich residue pool.

Biomass:
```{r}
vars <- c("B","synB","alphaR","respO","cueDef","limN","L") 
dsp0 <- resP %>% filter(.data$time < 1) %>% cueSesam() 
dsp <- dsp0 %>% select(all_of(c("time", "scen", "cn0", vars))) %>%
  gather("Pool","value", vars)
p2f <- dsp %>% 
  mutate(Approach = fct_relevel(.data$scen, c("Optimal","Derivative","Relative"))) %>% 
  mutate(Pool = fct_relevel(.data$Pool, vars)) %>% 
  ggplot(aes(x = time, y = value, col = Approach, linetype = Approach)) + geom_line(linewidth = 1) +
  facet_grid(Pool ~ cn0,scales = "free") + 
  scale_color_approach +
  scale_linetype_approach +
  xlab("Time (yr)") + 
  #xlab("Substrate (proportion of initial)") + scale_x_reverse() +
  #xlab("Fraction C decomposed") +
  #ylab(expression(Carbon~stock~(gCm^{-2}))) + #labs(linetype = "Substrate pool") +
  #scale_colour_discrete(drop = TRUE,limits = levels(dsp$Allocation)) +
  #theme(legend.position = c(0.95,1.04), legend.justification = c(1,1)) +
  themePaper +
  theme(panel.grid.major.x = element_line(colour = "grey75")) +
  theme(axis.title.y = element_blank()) +
  c()
#print(p2f + colScale)
print(p2f)
#ggsave("Sesam_OptimMethods_Pulse_approaches.png", width=4.8, height=3.8)


```

Biomass growing better in optimal approach.
However, partially due to starting at
higher levels obtained in steady-state simulation before.

Community adaptation leads to less overflow respiration and higher CUE during 
phase of N limitation.
However, N-limitation only induced if initial inorganic N levels are set low
assuming diluting during re-wetting of the soil. 
Over time, inorganic N levels increase, because
of decomposition an no loss from the system during incubation or during storage.
This is in contrast to field, where N is lost by plant uptake and leaching.

Note however, that for extremely large amendmends, of e.g. 500 gC/m^2,
also optimally allocating community shifts towards labile degrading shares
because return is much higher compared to residue levels.
This then leads to even higher overflow respiration compared to the fixed
and relative approaches.

### paper plot
```{r}
vars <- c("synB","alphaR","respO") 
labelmap = list(synB = "syn[B]", alphaR="alpha[R]", respO="resp[O]")
dsp0 <- resP %>% filter(.data$time <= 0.5) %>% cueSesam() 
dsp <- dsp0 %>% select(all_of(c("time", "scen", "cn0", vars))) %>%
  gather("Pool","value", vars)
p2f <- dsp %>% 
  filter(cn0 == 50) %>% 
  mutate(Approach = fct_relevel(.data$scen, c("Optimal","Derivative","Relative"))) %>% 
  mutate(Pool = fct_relevel(.data$Pool, vars)) %>% 
  mutate(Pool = fct_relabel(Pool, ~replace_labels(.,labelmap))) %>% 
  ggplot(aes(x = time, y = value, col = Approach, linetype = Approach)) + geom_line() +
  facet_grid(Pool ~ .,scales = "free",labeller = label_parsed) + 
  scale_color_approach +
  scale_linetype_approach +
  xlab("Time (yr)") + 
  #xlab("Substrate (proportion of initial)") + scale_x_reverse() +
  #xlab("Fraction C decomposed") +
  #ylab(expression(Carbon~stock~(gCm^{-2}))) + #labs(linetype = "Substrate pool") +
  #scale_colour_discrete(drop = TRUE,limits = levels(dsp$Allocation)) +
  #theme(legend.position = c(0.95,1.04), legend.justification = c(1,1)) +
  themePaper +
  theme(panel.grid.major.x = element_line(colour = "grey75")) +
  theme(axis.title.y = element_blank()) +
  c()
#print(p2f + colScale)
print(p2f)
#ggsave("Sesam_OptimMethods_Pulse_approaches.png", width=4.8, height=3.8)
```
```{r eval=FALSE, include=FALSE}
p3b <- p2f +
  #theme(legend.position = c(0.93,1.15), legend.justification = c(1,1))
  c()
ggsave("pulse_approaches.pdf", p3b, width = figWidthCol, height = figHeightCol) 
```


```{r}
#vars2 <- c("cueDef", "cueSyn") 
vars2 <- c("cueDef") 
dsp2 <- resP %>% 
  cueSesam() %>% 
  mutate(S = R + L, cnS = (R + L)/(RN + LN)) %>% 
  mutate(relS = S/S[1], relL = L/L[1], fracDecomp = 1 - relL) %>% 
  mutate(cn0 = as.factor(cn0)) %>% 
  select(all_of(c("time","relS","relL","fracDecomp","scen","cn0", vars2))) %>% gather("Pool","value", vars2)

# theme_set(themeUsed) 

p2 <- dsp2 %>% 
  mutate(scen = fct_relevel(.data$scen, c("Optimal","Derivative","Relative"))) %>% 
  filter(scen %in% c("Optimal","Relative","Fixed")) %>% 
  ggplot(aes(x = fracDecomp, y = value, col = cn0, linetype = cn0)) + 
#p2 <- ggplot( dsp2, aes(x = fracDecomp, y = value, col = cn0, linetype = cn0)) + 
  geom_line(linewidth = 1) +
  facet_grid(~ scen) + #,scales = "free") + 
  #xlab("Time (yr)") + 
  #xlab("Substrate (proportion of initial)") + scale_x_reverse() +
  xlab("Fraction C decomposed") +
  ylab("Carbon use efficiency") +
  expand_limits(y = 0.3) +
  #themePres + # presentation
  guides(color=guide_legend(title = "C:N amendment")) +
  guides(linetype=guide_legend(title = "C:N amendment")) +
  #themeDefault +
  # scale_colour_discrete(drop = TRUE,limits = levels(dsp$Allocation)) +
  #theme(legend.position = c(0.95,1.04), legend.justification = c(1,1)) +
  theme(legend.position = c(0.05,0.05), legend.justification = c(0,0)) +
  #theme(panel.border = element_rect(colour = "black", fill = NA)) +
  #theme(strip.background = element_blank()) +			
  #theme(panel.grid.major.x = element_line(colour = "grey75")) +
  c()
#print(p2f + colScale)
print(p2)
```
```{r eval=FALSE, include=FALSE}
p3b <- p2 +
  #theme(legend.position = c(0.93,1.15), legend.justification = c(1,1)) +
  themePaper +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  theme(legend.position = "bottom") +
  guides(linetype = guide_legend(title = "C:N", nrow = 1, byrow = TRUE)) +
  guides(color = guide_legend(title = "C:N", nrow = 1, byrow = TRUE)) +
  #facet_grid(scen ~ .) + #,scales = "free") + 
  #theme(legend.title = element_blank()) +
  c()
ggsave("pulse_cn.pdf", p3b, width = figWidthCol, height = figHeightCol) 
```



```{r simPulse, fig.height=2.2, fig.width=3.27}
p2p <- p2
p2p$layers[2] <- NULL # remove expand_limits
p2p <- p2p +
  themePaper +
  #theme(legend.position = c(0.02,0.02), legend.justification = c(0,0)) +
  theme(legend.position = "bottom") +
  theme(panel.border = element_rect(colour = "black", fill = NA)) +
  theme(strip.background = element_blank()) +	
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  #theme(legend.title = element_blank()) +
  c()  
print(p2p)
```
## Inspect at low biomass

```{r}
vars <- c("B","synB","alphaR","cueDef","L","R") 
vars <- c("B","alphaR","cueDef","L","R") 
vars <- c("B","alphaR","L") 
dsp0 <- resP %>% filter(between(.data$time,1.1,5)) %>% cueSesam() 
dsp <- dsp0 %>% select(all_of(c("time", "scen", "cn0", vars))) %>%
  gather("Pool","value", vars) %>% 
  mutate(Pool = fct_relevel(Pool, vars)) %>% 
  mutate(Pool = fct_relabel(Pool, ~replace_labels(.,labelmap)))
  
p2f <- dsp %>% 
  mutate(Approach = fct_relevel(.data$scen, c("Optimal","Derivative","Relative"))) %>% 
  filter(scen %in% c("Optimal","Relative")) %>% 
  filter(cn0 == 50) %>% 
  #mutate(Pool = fct_relevel(.data$Pool, vars)) %>% 
  ggplot(aes(x = time, y = value, col = Approach, linetype = Approach)) + geom_line(linewidth = 1) +
  #facet_grid(Pool ~ cn0,scales = "free") + 
  facet_grid(Pool~.,scales = "free",labeller = label_parsed) + 
  xlab("Time (yr)") + 
  scale_color_approach +
  scale_linetype_approach +
  #xlab("Substrate (proportion of initial)") + scale_x_reverse() +
  #xlab("Fraction C decomposed") +
  #ylab(expression(Carbon~stock~(gCm^{-2}))) + #labs(linetype = "Substrate pool") +
  #theme(legend.position = c(0.95,1.04), legend.justification = c(1,1)) +
  #theme(panel.grid.major.x = element_line(colour = "grey75")) +
  theme(axis.title.y = element_blank()) +
  c()
#print(p2f + colScale)
print(p2f)
#ggsave("Sesam_OptimMethods_Pulse_approaches.png", width=4.8, height=3.8)
```
Optimal approach does not allocate to L degrading enzymes any more.
Yielding higher biomass and higher R decomposition.

```{r eval=FALSE, include=FALSE}
p3b <- p2f +
  themePaper +
  theme(axis.title.y = element_blank()) +
  #theme(legend.position = c(0.93,1.15), legend.justification = c(1,1))
  c()
ggsave("pulse_lowB.pdf", p3b, width = figWidthCol, height = figHeightCol) 
```

```{r}
# Compare the initial derivative at cn90

```

